### 2024-03-18 scripts/examples/manuscript/index.qmd
fix(scripts): comment examples authors Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Notebook template"
toc: true
number-sections: true
highlight-style: pygments
csl: bibstyle.csl
format:
  nature-pdf:
    natbibstyle: sn-nature
    classoption: 
      - lineno
    cite-method: citeproc
    keep-tex: true
  html:
    mainfont: Latin Modern Roman
    code-fold: true
    html-math-method: katex
  docx: default
execute: 
  cache: true
  keep-ipynb: true
author:
  - name: Pyrovelocity Authors
  # - name: Qian Qin
  #   affiliations:
  #     - name: Broad Institute of MIT and Harvard
  #       department: Broad Institute of MIT and Harvard
  #       id: 1
  #       city: Cambridge
  #       state: MA
  #   email: qqin@broadinstitute.org
  #   attributes:
  #     equal-contributor: true
  # - name: Cameron Smith
  #   affiliations:
  #     - ref: 1
  #     - name: Harvard Medical School
  #       id: 2
  #       department: Department of Pathology
  #       city: Boston
  #       state: MA
  #     - name: Massachusetts General Hospital
  #       id: 3
  #       department: Molecular Pathology Unit, Center for Cancer Research
  #       city: Boston
  #       state: MA
  #   email: cameron.smith@mgh.harvard.edu
  #   attributes:
  #     equal-contributor: true
  # - name: Luca Pinello
  #   affiliations:
  #     - ref: 1
  #     - ref: 2
  #     - ref: 3
  #   email: lpinello@mgh.harvard.edu
  #   attributes:
  #     corresponding: true
abstract: |
  The abstract serves both as a general introduction to the topic and as a
  brief, non-technical summary of the main results and their implications.
  Authors are advised to check the author instructions for the journal they are
  submitting to for word limits and if structural elements like subheadings,
  citations, or equations are permitted.
keywords: [template, demo]
bibliography: references.bib
jupyter:
  jupytext:
    cell_metadata_filter: all
    cell_metadata_json: true
    notebook_metadata_filter: all
    text_representation:
      extension: .qmd
      format_name: quarto
      format_version: '1.0'
      jupytext_version: 1.16.0
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
---

# Summary {#sec-summary}

Use this file as a template for creating new notebooks that can be rendered to
pdf, html, and ipynb from the common source of truth in this markdown file. You
can include citations using the [default pandoc citation syntax][pandoc-citation]
`[@key01; @key02]` [@La_Manno2018-lj; @Qin2022-ls].

# Dependencies {#sec-dependencies}

Copies of this template will require the following dependencies:

- [quarto-cli][quarto-pypi] or [quarto][quarto-conda]
- [jupyter-cache][jupyter-cache-pypi]

Its other dependencies are not specific to the purpose of the template.

# Usage {#sec-usage}

Install the \LaTeX template with:

``` bash
quarto add cameronraysmith/nature@natbib
```

This will add an [_extensions/](./_extensions/) subdirectory containing
the template extension.

Render the notebook with:

``` bash
quarto render index.qmd --to all -P TEST_MODE:False
```

With the current configuration in the YAML header
this should produce

- [index.pdf](./index.pdf)
- [index.html](./index.html)
- [index.ipynb](./index.ipynb)

together with the 
[index_files](./index_files/) and 
[.jupyter_cache](./.jupyter_cache/) 
subdirectories that contain the quarto output and the cache of the jupyter
notebook execution. What follows are just a few example computations to
illustrate notebook content. At least @sec-dependencies and @sec-usage can be
deleted from any instance of this notebook.

# Setup {#sec-setup}

```{python}
#| execution: {iopub.execute_input: '2024-02-28T00:03:22.418885Z', iopub.status.busy: '2024-02-28T00:03:22.418672Z', iopub.status.idle: '2024-02-28T00:03:22.700860Z', shell.execute_reply: '2024-02-28T00:03:22.700498Z'}
#| lines_to_next_cell: 2
#| slideshow: {slide_type: fragment}
#| tags: []
import os

import matplotlib.pyplot as plt
import numpy as np
```

## Parameters {#subsec-setup-parameters}

The following cell defines parameters to be exposed to quarto or papermill
execution using the `tags: [parameters]` in the cell below

```{python}
#| execution: {iopub.execute_input: '2024-02-28T00:03:22.715524Z', iopub.status.busy: '2024-02-28T00:03:22.715310Z', iopub.status.idle: '2024-02-28T00:03:22.718434Z', shell.execute_reply: '2024-02-28T00:03:22.717807Z'}
#| slideshow: {slide_type: fragment}
#| tags: [parameters]
TEST_MODE = True
print(f"Test mode: {TEST_MODE}") 
```

Parameters injected by quarto or papermill should appear above.

## Functions {#subsec-setup-functions}

Define functions to be used in the notebook.

```{python}
#| execution: {iopub.execute_input: '2024-02-28T00:03:22.744628Z', iopub.status.busy: '2024-02-28T00:03:22.744439Z', iopub.status.idle: '2024-02-28T00:03:22.746661Z', shell.execute_reply: '2024-02-28T00:03:22.746335Z'}
#| slideshow: {slide_type: fragment}
#| tags: []
def plot_random_numbers(size=10):
    plt.plot(np.random.rand(size))
    plt.show()
```

This cell checks if the notebook is running in test mode. If it is, it will reduce the size of the data to be processed.

```{python}
#| execution: {iopub.execute_input: '2024-02-28T00:03:22.756996Z', iopub.status.busy: '2024-02-28T00:03:22.756828Z', iopub.status.idle: '2024-02-28T00:03:22.759487Z', shell.execute_reply: '2024-02-28T00:03:22.759002Z'}
#| slideshow: {slide_type: fragment}
#| tags: []
print(f"Test mode: {TEST_MODE}")
size = 5 if TEST_MODE else 10
print(f"Size: {size}")
```

# Results {#sec-results}

The results of the analysis are shown in @fig-random-numbers.

```{python}
#| label: fig-random-numbers 
#| fig-cap: "Plot of samples from the uniform distribution."
#| execution: {iopub.execute_input: '2024-02-28T00:03:22.775090Z', iopub.status.busy: '2024-02-28T00:03:22.774911Z', iopub.status.idle: '2024-02-28T00:03:22.854889Z', shell.execute_reply: '2024-02-28T00:03:22.854428Z'}
#| slideshow: {slide_type: fragment}
#| tags: []
plot_random_numbers(size)
```

# Postprocessing {#sec-postprocessing}

The notebook can also be converted to ipynb using

``` bash
quarto convert index.qmd
```

or vice-versa to markdown via

``` bash
quarto convert index.ipynb
```

figures should already be present in the [index_files/](./index_files/).
You can always rerender the notebook to generate them

``` bash
quarto render index.qmd --to all
```

If you are concerned that the cache me be invalidated without being detected
you can override it with

``` bash
quarto render index.qmd --to all --cache-refresh
```

# References

:::{#refs}
:::

<!-- links -->
[pandoc-citation]: https://pandoc.org/MANUAL.html#citation-syntax
[quarto-pypi]: https://pypi.org/project/quarto-cli/
[quarto-conda]: https://anaconda.org/conda-forge/quarto
[jupyter-cache-pypi]: https://pypi.org/project/jupyter-cache/