### 2024-03-24 nbs/guides/installation.qmd
feat(nbs): init installation Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Installation
---

## PyPI

## Anaconda