### 2024-03-24 nbs/index.qmd
fix(nbs): remove gh pages badge Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Pyro-Velocity
description: "Probabilistic modeling of RNA velocity"
format:
  html:
    toc: false

about:
  id: information
  template: jolla
  image: _static/logo.png
  links: 
    - icon: download
      href: guides/installation.qmd
    - icon: book
      href: tutorials/introduction.qmd
    - icon: github
      text: GitHub
      href: https://github.com/pinellolab/pyrovelocity
---

::: {#information}
:::

<div style="margin-left: auto; margin-right: auto; max-width: 560px; overflow-x: auto;">

|         |                                                                                                                                                                        |
| ------- | ---------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| CI/CD   | [![CID][cid-badge]][cid-link] [![CML][cml-badge]][cml-link] [![pre-commit.ci status][precommit-badge]][precommit-link]                                                 |
| Docs    | [![Cloudflare Pages][cloudflare-badge]][cloudflare-link] [![Preprint][preprint-badge]][preprint-link]                 |
| Package | [![PyPI - Version][pypi-badge]][pypi-link] [![Anaconda badge][anaconda-badge]][anaconda-link] [![Docker image][docker-badge]][docker-link]                             |
| Meta    | [![codecov][codecov-badge]][codecov-link] [![code style][black-badge]][black-link] [![License - MIT][license-badge]][license-link] [![Tuple][tuple-badge]][tuple-link] |

[cid-badge]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cid.yaml/badge.svg?branch=master
[cid-link]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cid.yaml
[cml-badge]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cml.yml/badge.svg
[cml-link]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cml.yml
[precommit-badge]: https://results.pre-commit.ci/badge/github/pinellolab/pyrovelocity/main.svg
[precommit-link]: https://results.pre-commit.ci/latest/github/pinellolab/pyrovelocity/master
[cloudflare-badge]: https://img.shields.io/badge/Docs-pages-F38020?style=flat&logo=cloudflare&labelColor=black
[cloudflare-link]: https://docs.pyrovelocity.net
[preprint-badge]: https://img.shields.io/badge/doi-10.1101/2022.09.12.507691v2-B31B1B
[preprint-link]: https://doi.org/10.1101/2022.09.12.507691
[pypi-badge]: https://img.shields.io/pypi/v/pyrovelocity.svg?logo=pypi&label=PyPI&logoColor=gold
[pypi-link]: https://pypi.org/project/pyrovelocity/
[anaconda-badge]: https://anaconda.org/conda-forge/pyrovelocity/badges/version.svg
[anaconda-link]: https://anaconda.org/conda-forge/pyrovelocity
[docker-badge]: https://img.shields.io/badge/docker-image-blue?logo=docker
[docker-link]: https://github.com/pinellolab/pyrovelocity/pkgs/container/pyrovelocity
[codecov-badge]: https://codecov.io/gh/pinellolab/pyrovelocity/branch/main/graph/badge.svg
[codecov-link]: https://codecov.io/gh/pinellolab/pyrovelocity
[black-badge]: https://img.shields.io/badge/code%20style-black-000000.svg
[black-link]: https://github.com/psf/black
[license-badge]: https://img.shields.io/badge/license-AGPL%203-purple
[license-link]: https://spdx.org/licenses/
[tuple-badge]: https://img.shields.io/badge/Tuple%20❤️%20OSS-5A67D8?logo=tuple
[tuple-link]: https://tuple.app/github-badge

</div>