### 2024-03-29 nbs/index.qmd
docs(readme): update badges
- add pyro, hydra-zen, and scvi-tools
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Pyro-Velocity
description: "𝒫robabilistic modeling of RNA velocity ⬱"
format:
  html:
    toc: false

about:
  id: information
  template: jolla
  image: _static/logo.png
  links: 
    - icon: download
      href: guides/installation.qmd
    - icon: book
      href: tutorials/introduction.qmd
    - icon: github
      text: GitHub
      href: https://github.com/pinellolab/pyrovelocity
---

::: {#information}
:::

<div style="margin-left: auto; margin-right: auto; max-width: 520px; overflow-x: auto;">

|         |                                                                                                                                                           |
| ------- | --------------------------------------------------------------------------------------------------------------------------------------------------------- |
| CI/CD   | [![CID][cid-badge]][cid-link] [![codecov][codecov-badge]][codecov-link] [![pre-commit.ci status][precommit-badge]][precommit-link]                        |
| Docs    | [![Cloudflare Pages][cloudflare-badge]][cloudflare-link] [![Preprint][preprint-badge]][preprint-link]                                                     |
| Package | [![PyPI - Version][pypi-badge]][pypi-link] [![Conda-forge badge][conda-forge-badge]][anaconda-link] [![Docker image][docker-badge]][docker-link]          |
| Meta    | [![pyro-ppl-badge]][pyro-ppl-link] [![flyte-badge]][flyte-link] [![hydra-zen-badge]][hydra-zen-link] [![scvi-tools-badge]][scvi-tools-link] [![code style][black-badge]][black-link] [![License - MIT][license-badge]][license-link] [![Tuple][tuple-badge]][tuple-link] |

[cid-badge]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cid.yaml/badge.svg?branch=master
[cid-link]: https://github.com/pinellolab/pyrovelocity/actions/workflows/cid.yaml
[precommit-badge]: https://results.pre-commit.ci/badge/github/pinellolab/pyrovelocity/main.svg
[precommit-link]: https://results.pre-commit.ci/latest/github/pinellolab/pyrovelocity/main
[flyte-badge]: https://img.shields.io/badge/Flyte-carrier-gray.svg?color=7552A2&logo=data:image/svg%2bxml;base64,PHN2ZyBpZD0iTGF5ZXJfMSIgZGF0YS1uYW1lPSJMYXllciAxIiB4bWxucz0iaHR0cDovL3d3dy53My5vcmcvMjAwMC9zdmciIHhtbG5zOnhsaW5rPSJodHRwOi8vd3d3LnczLm9yZy8xOTk5L3hsaW5rIiB2aWV3Qm94PSIwIDAgMzYwIDMzMy40ODM3OSI+PGRlZnM+PHN0eWxlPi5jbHMtMXtmaWxsOnVybCgjbGluZWFyLWdyYWRpZW50KTt9PC9zdHlsZT48bGluZWFyR3JhZGllbnQgaWQ9ImxpbmVhci1ncmFkaWVudCIgeDE9IjYuMTkzNjUiIHkxPSIxNjcuMzQ3NDUiIHgyPSIzNTIiIHkyPSIxNjcuMzQ3NDUiIGdyYWRpZW50VW5pdHM9InVzZXJTcGFjZU9uVXNlIj48c3RvcCBvZmZzZXQ9IjAiIHN0b3AtY29sb3I9IiNlNWM5ZDMiLz48c3RvcCBvZmZzZXQ9IjAuMzMiIHN0b3AtY29sb3I9IiNiMzg1ZGYiLz48c3RvcCBvZmZzZXQ9IjEiIHN0b3AtY29sb3I9IiM2ZjJhZWYiLz48L2xpbmVhckdyYWRpZW50PjwvZGVmcz48cGF0aCBjbGFzcz0iY2xzLTEiIGQ9Ik05MC45NjQ3NSwxNzAuODEzMzRjNy42Mjc3Ny0xOC4xMjcsMTEuNTEwNDctMzYuNzc2NTUsMTUuMDA2NC01NS40OTI4OSw1Ljk2MDM3LTMxLjkxMDIsOS45OS02NC4wMzIzLDguNjM2MjEtOTYuNTk0NzUtLjI1NzIzLTYuMTg3NDIsMS40ODQ5NS04LjQ4NzYzLDguMTQ5My03Ljg1ODIzLDE4LjA0NDY1LDEuNzA0MjUsMzUuOTM3NDYsMy43NjY1Nyw1Mi4xNTksMTIuNjUyODYsNC4yOTQyNSwyLjM1MjQxLDguMDYyNzIsMS4wODk0OSwxMi4xNjgzLS4wNjgxLDMxLjA0ODU5LTguNzU0NCw2Mi41MzI3Mi0xNC42OTE0Niw5NC45NjgyNS0xNC4wMjQ1NiwyMS4zMTQuNDM4Miw0MS45ODQyNSwzLjg3ODkxLDYxLjY0Mzc5LDEyLjcwMjc2LDYuNDI3MjksMi44ODQ3OSw4LjgwODY5LDYuOTg2NDEsOC4yMTY1NywxMy44MS0xLjU5MzM4LDE4LjM2MDg0LTguOTMzNDUsMzQuODE3NTEtMTYuNTU1NzEsNTEuMTgwMzgtMi4wODM3NSw0LjQ3MzEzLTUuMTEzMiwxLjkwMDYzLTcuNjg0MTcuMjc2ODYtMjUuODcwMzMtMTYuMzM4ODYtNTMuNjUxLTI4LjcwMDYzLTgyLjI2NC0zOS4yNTUxNi0yMC4zMDkwNi03LjQ5MTQ2LTQwLjg0NzY4LTE0LjMwOTczLTYzLjA1ODIzLTE3LjcxMjkzLDYuODUzODgsNi43ODE4MiwxMy40NjQwNywxMy44MzI2NiwyMC42MDM1NSwyMC4yOTkxMiwzMS42MDk3OSwyOC42Myw2NC41NTE4NSw1NS40OTQsMTAyLjYyNDMsNzUuMzE3NjQsNC4yMDA1MiwyLjE4NzEyLDUuMjE5NDYsNC40Mjk0MiwyLjMzNzMyLDguNTIyMTMtMTAuNTgzNzQsMTUuMDI5MDgtMjEuNDEwODUsMjkuODczMzEtMzcuMjkwNjQsMzkuODIzMTItNC43MjI4MywyLjk1OTE4LTUuODUyNDYsNy4xNjgzOC03LjAyMjgzLDExLjg0MjItOS4xMDQzMywzNi4zNTkyOS0yMi4zMzIsNzAuOTU1ODEtNDQuMDI1MzEsMTAxLjg2NTA3YTE0MS4zMjU0NSwxNDEuMzI1NDUsMCwwLDEtMzIuNzkzMiwzMy44NDUyMWMtNS44MDIwOSw0LjE5ODExLTEwLjk2OTg0LDQuNjkyLTE2Ljg1MDguMjcxOC0xNC4xOTA0Ny0xMC42NjU4OS0yNC44NjAxOS0yNC40MDA5My0zNC43NTE4NC0zOC44NzctMi43MTgzNy0zLjk3ODMzLS43ODg4Ni02LjE1NjEyLDIuNzA4NC03Ljk1NSwyOC41MTE4My0xNC42NjUyNyw1My44NDQ5NS0zMy45MTk2Nyw3OC4yNjQzNi01NC40NTM3YTQ1Mi44NjksNDUyLjg2OSwwLDAsMCw0Mi41MjQwOS00MC4wNmMtMi44MDMzNy0yLjY3MTQxLTUuMjA3ODItLjcwNTI5LTcuMzA0OC0uMTQ2MTctNDcuMDUyLDEyLjU0NS05Mi4zNzI1NCwyOS4zODkyMS0xMzQuMDQwODIsNTUuMDEwOS01LjY5NTE1LDMuNTAxOTMtOC43MjM0NiwzLjA2NjE2LTExLjQ5OTg1LTMuMzM5MzctNy4wOTI2OC0xNi4zNjM2Ny0xMy42NjczNC0zMi44Njc1MS0xNC43NzQxOS01MC44NjkzNi0uMzYyNjgtNS44OTgxNC0zLjk2ODM2LTkuMDc1NTUtNy41NjI2NC0xMi41Njg0Ni0yNS42NTEyMy0yNC45MjcxNy00Ny45NjEtNTIuMzc0ODMtNjMuNDg0ODYtODQuODQzN0MxMi40MjgyOSw2OC4yNTUsNi44ODkxOSw1MS44MDAxNCw2LjIxMDg5LDMzLjkzMzgxYy0uMjY5LTcuMDg2MDYsMi42MDM2OC0xMS4wMzkyNSw5LjAwMTQ3LTEzLjM0NDlhMTY4LjI0NDkxLDE2OC4yNDQ5MSwwLDAsMSw0Ni4xNTYtOS42ODQxNGM2LjMyMDE1LS40MjE2Miw3Ljg3MzEsMS44MDQzMyw3LjcwNzY4LDcuNTA2NjgtMS4xNjg0MSw0MC4yODA2LDQuOTg5MjMsNzkuNzU1LDEzLjU3NjQ2LDExOC45MjMsMi40NjE1MiwxMS4yMjc0MSw1LjUyNDg2LDIyLjMyMjk0LDguMzEyMjcsMzMuNDc4ODRaTTE3OS40MTM2OSwzOS41MDhDMTY5LjA5Niw5Mi45NjE2MSwxMzYuNjc2ODUsMTMyLjkwNDIxLDEwMC4xMzIxMywxNzAuNTI4MmM0OS44OTE3Ny0yMC4zNzEsOTkuNjI4NTctMTEuMjM1NywxNDkuODI1NDMtMS4zMjA3NkMyMTEuNDQzODcsMTMzLjc1OTY5LDE5My45ODkyNiw4Ny43MTM2MywxNzkuNDEzNjksMzkuNTA4WiIvPjwvc3ZnPg==
[flyte-link]: https://docs.flyte.org
[hydra-zen-link]: https://mit-ll-responsible-ai.github.io/hydra-zen/
[hydra-zen-badge]: https://img.shields.io/badge/Hydra-zen-gray.svg?color=193966&logo=data:image/svg%2bxml;base64,PD94bWwgdmVyc2lvbj0iMS4wIiBlbmNvZGluZz0idXRmLTgiPz4KPCEtLSBHZW5lcmF0b3I6IEFkb2JlIElsbHVzdHJhdG9yIDI1LjQuMSwgU1ZHIEV4cG9ydCBQbHVnLUluIC4gU1ZHIFZlcnNpb246IDYuMDAgQnVpbGQgMCkgIC0tPgo8c3ZnIHZlcnNpb249IjEuMSIgaWQ9IkxheWVyXzEiIHhtbG5zPSJodHRwOi8vd3d3LnczLm9yZy8yMDAwL3N2ZyIgeG1sbnM6eGxpbms9Imh0dHA6Ly93d3cudzMub3JnLzE5OTkveGxpbmsiIHg9IjBweCIgeT0iMHB4IgoJIHZpZXdCb3g9IjAgMCA1MDAgNTAwIiBzdHlsZT0iZW5hYmxlLWJhY2tncm91bmQ6bmV3IDAgMCA1MDAgNTAwOyIgeG1sOnNwYWNlPSJwcmVzZXJ2ZSI+CjxzdHlsZSB0eXBlPSJ0ZXh0L2NzcyI+Cgkuc3Qwe2ZpbGw6IzE5Mzk2Njt9Cgkuc3Qxe2ZpbGw6I0ZGRkZGRjt9Cjwvc3R5bGU+CjxwYXRoIGNsYXNzPSJzdDAiIGQ9Ik00NDkuMywwSDUwLjdDMjIuOCwwLDAsMjIuOCwwLDUwLjd2Mzk4LjVDMCw0NzcuMiwyMi44LDUwMCw1MC43LDUwMGgzOTguNWMyNy45LDAsNTAuNy0yMi44LDUwLjctNTAuN1Y1MC43CglDNTAwLDIyLjgsNDc3LjIsMCw0NDkuMywweiIvPgo8Zz4KCTxwYXRoIGNsYXNzPSJzdDEiIGQ9Ik0zOC4yLDM5Mi4xYzAtMi45LDIuMy01LjIsNS4yLTUuMmMzLDAsNS4zLDIuNCw1LjMsNS4yVjQxMmMzLTQuMyw3LjItOCwxNC40LThjMTAuNCwwLDE2LjQsNywxNi40LDE3LjZ2MjQuNgoJCWMwLDMtMi4zLDUuMi01LjIsNS4yYy0zLDAtNS4zLTIuMy01LjMtNS4ydi0yMS40YzAtNy4yLTMuNi0xMS4zLTkuOS0xMS4zYy02LjEsMC0xMC40LDQuMy0xMC40LDExLjR2MjEuMmMwLDMtMi40LDUuMi01LjMsNS4yCgkJYy0yLjksMC01LjItMi4zLTUuMi01LjJWMzkyLjF6Ii8+Cgk8cGF0aCBjbGFzcz0ic3QxIiBkPSJNOTIuNyw0NjMuOWMtMS42LTAuNS0zLjMtMS44LTMuMy00LjRjMC0yLjYsMi00LjQsNC4zLTQuNGMwLjksMCwxLjQsMC4yLDEuOSwwLjNjMS4yLDAuMywyLjIsMC42LDMuNywwLjYKCQljMy4xLDAsNC44LTEuMyw2LjYtNC45bDAuMy0xbC0xNy4xLTM4LjNjLTAuMy0wLjgtMC42LTEuOC0wLjYtMi41YzAtMywyLjMtNS4yLDUuMi01LjJjMi43LDAsNC4yLDEuNiw1LjIsNGwxMi40LDMwLjdsMTEuNS0zMC43CgkJYzAuOS0yLjIsMi40LTQsNS4xLTRjMi44LDAsNS4xLDIuMiw1LjEsNWMwLDAuOS0wLjMsMi0wLjQsMi41bC0xNi45LDQwLjRjLTMuOSw5LjQtOC4xLDEyLjktMTUuNSwxMi45CgkJQzk3LjEsNDY1LjIsOTQuOSw0NjQuOCw5Mi43LDQ2My45eiIvPgoJPHBhdGggY2xhc3M9InN0MSIgZD0iTTE4Ni44LDQ0Ni4yYzAsMy0yLjQsNS4yLTUuMiw1LjJjLTMsMC01LjMtMi4zLTUuMy01LjJ2LTIuOWMtMy40LDQuOC04LjIsOC42LTE1LjcsOC42CgkJYy0xMC44LDAtMjEuNC04LjYtMjEuNC0yMy45di0wLjJjMC0xNS4zLDEwLjQtMjMuOSwyMS40LTIzLjljNy42LDAsMTIuNCwzLjgsMTUuNyw4LjF2LTIwYzAtMywyLjQtNS4yLDUuMi01LjIKCQljMywwLDUuMywyLjMsNS4zLDUuMlY0NDYuMnogTTE0OS44LDQyNy45djAuMmMwLDguOSw2LjEsMTQuOCwxMy4zLDE0LjhjNy4xLDAsMTMuNC01LjksMTMuNC0xNC44di0wLjJjMC04LjktNi4zLTE0LjgtMTMuNC0xNC44CgkJQzE1NS44LDQxMy4xLDE0OS44LDQxOC43LDE0OS44LDQyNy45eiIvPgoJPHBhdGggY2xhc3M9InN0MSIgZD0iTTIwMC4xLDQwOS43YzAtMywyLjMtNS4zLDUuMi01LjNjMywwLDUuMywyLjQsNS4zLDUuM3Y0LjdjMi40LTUuOCw3LTEwLjEsMTEuNi0xMC4xYzMuMywwLDUuMiwyLjIsNS4yLDUuMgoJCWMwLDIuOC0xLjgsNC42LTQuMiw1LjFjLTcuNSwxLjMtMTIuNyw3LjEtMTIuNywxOHYxMy42YzAsMi45LTIuNCw1LjItNS4zLDUuMmMtMi45LDAtNS4yLTIuMy01LjItNS4yVjQwOS43eiIvPgoJPHBhdGggY2xhc3M9InN0MSIgZD0iTTIzMC45LDQzNy43di0wLjJjMC0xMCw3LjgtMTQuOCwxOS0xNC44YzUuMiwwLDguOCwwLjgsMTIuNCwxLjl2LTEuMWMwLTYuNi00LTEwLTExLjQtMTAKCQljLTQsMC03LjMsMC43LTEwLjIsMS44Yy0wLjYsMC4yLTEuMSwwLjMtMS43LDAuM2MtMi40LDAtNC41LTEuOS00LjUtNC40YzAtMS45LDEuMy0zLjYsMi45LTQuMmM0LjQtMS43LDguOC0yLjcsMTQuOC0yLjcKCQljNi45LDAsMTIuMSwxLjgsMTUuMyw1LjJjMy40LDMuMyw1LDguMiw1LDE0LjJ2MjIuN2MwLDIuOS0yLjMsNS4xLTUuMiw1LjFjLTMuMSwwLTUuMi0yLjEtNS4yLTQuNXYtMS43Yy0zLjEsMy44LTcuOSw2LjctMTUsNi43CgkJQzIzOC42LDQ1MiwyMzAuOSw0NDcsMjMwLjksNDM3Ljd6IE0yNjIuNSw0MzQuNHYtMy4xYy0yLjctMS02LjMtMS44LTEwLjUtMS44Yy02LjgsMC0xMC44LDIuOS0xMC44LDcuN3YwLjJjMCw0LjUsMy45LDcsOSw3CgkJQzI1Ny4yLDQ0NC4zLDI2Mi41LDQ0MC4zLDI2Mi41LDQzNC40eiIvPgoJPHBhdGggY2xhc3M9InN0MSIgZD0iTTI4My4xLDQyNC4zYzAtMi45LDIuNC01LjIsNS4yLTUuMmgxNi40YzIuOCwwLDUuMiwyLjQsNS4yLDUuMmMwLDIuOC0yLjQsNS4yLTUuMiw1LjJoLTE2LjQKCQlDMjg1LjQsNDI5LjQsMjgzLjEsNDI3LjEsMjgzLjEsNDI0LjN6Ii8+Cgk8cGF0aCBjbGFzcz0ic3QxIiBkPSJNMzE3LjcsNDQ2Ljd2LTAuMmMwLTIsMC43LTMuMywyLjQtNS4xbDI0LjUtMjcuOWgtMjEuN2MtMi40LDAtNC40LTEuOS00LjQtNC4zYzAtMi40LDEuOS00LjQsNC40LTQuNGgzMC40CgkJYzIuOSwwLDUuMSwxLjcsNS4xLDQuMnYwLjNjMCwyLTAuNywzLjMtMi4zLDUuMWwtMjQuNiwyNy45aDIyLjdjMi40LDAsNC40LDEuOSw0LjQsNC40YzAsMi40LTEuOSw0LjMtNC40LDQuM2gtMzEuMwoJCUMzMTkuOCw0NTEsMzE3LjcsNDQ5LjIsMzE3LjcsNDQ2Ljd6Ii8+Cgk8cGF0aCBjbGFzcz0ic3QxIiBkPSJNMzg4LjcsNDUyLjFjLTEzLjQsMC0yMy43LTkuNy0yMy43LTIzLjl2LTAuMmMwLTEzLjIsOS4zLTI0LDIyLjUtMjRjMTQuNywwLDIxLjksMTIuMSwyMS45LDIyLjcKCQljMCwzLTIuMyw1LjEtNSw1LjFoLTI4LjljMS4xLDcuNSw2LjUsMTEuNywxMy4zLDExLjdjNC41LDAsNy45LTEuNiwxMC45LTRjMC44LTAuNiwxLjUtMSwyLjctMWMyLjQsMCw0LjIsMS44LDQuMiw0LjMKCQljMCwxLjMtMC42LDIuNC0xLjQsMy4yQzQwMS4yLDQ0OS43LDM5NS45LDQ1Mi4xLDM4OC43LDQ1Mi4xeiBNMzk5LjIsNDI0LjdjLTAuNy02LjgtNC43LTEyLjEtMTEuNy0xMi4xYy02LjUsMC0xMSw1LTEyLDEyLjFIMzk5LjIKCQl6Ii8+Cgk8cGF0aCBjbGFzcz0ic3QxIiBkPSJNNDIwLjQsNDA5LjdjMC0zLDIuMy01LjMsNS4yLTUuM2MzLDAsNS4zLDIuNCw1LjMsNS4zdjIuM2MzLTQuMyw3LjItOCwxNC40LThjMTAuNCwwLDE2LjQsNywxNi40LDE3LjZ2MjQuNgoJCWMwLDMtMi4zLDUuMi01LjIsNS4yYy0zLDAtNS4zLTIuMy01LjMtNS4ydi0yMS40YzAtNy4yLTMuNi0xMS4zLTkuOS0xMS4zYy02LjEsMC0xMC40LDQuMy0xMC40LDExLjR2MjEuMmMwLDMtMi40LDUuMi01LjMsNS4yCgkJYy0yLjksMC01LjItMi4zLTUuMi01LjJWNDA5Ljd6Ii8+CjwvZz4KPHBhdGggY2xhc3M9InN0MSIgZD0iTTQ3My4xLDM0MS42Yy0zLjgsMy4yLTEzLjEsMy44LTEzLjEsMy44bDMuNC0yLjFoLTNjNy01LDguNi0xNS4yLDguNi0xNS4yYy0xLjQsMi4xLTgsNi42LTEwLjcsNy45CgljLTIuNywxLjMtMTUuMSwwLjYtMTUuMSwwLjZjMy45LTEuNCw2LTMuMiw2LTMuMmgtMi41YzEuOC0xLjQsNS42LTcuNSw1LjYtNy41bC0yLjMsMC42YzMuNS01LTQuMi0xNS45LTQuMi0xNS45CgljMi41LDcuNy0zLjIsMTUuMi0zLjIsMTUuMnMtMy44LDkuMi0xMy4zLDcuM2MtOS41LTEuOC04LjEtMTEuMi02LjEtMTJjMi4xLTAuOCw0LjktNS40LDQuOS01LjRjMi4zLTYuOSwxNi4xLTYuOSwxNi4xLTYuOQoJYy03LjgtNy4xLTIwLjgtMC41LTIwLjgtMC41bDAuOC0yLjRjLTEuNywyLjEtOS44LDguMS05LjgsOC4xbDEuMS0zLjRjLTEuNiwyLjktMTAuMiwxMC42LTEwLjIsMTAuNnYtMi45CgljLTAuMyw0LjEtMTEuMyw5LjYtMTQuMiw4LjljMCwwLTQuOS0wLjUtMTAuNi02Yy01LjgtNS41LTI1LjktMTUuMy0yNS45LTE1LjNjLTE1LjctOC40LTEzLjgtMzEuMS0xMy44LTMxLjFsMCwwCgljMS45LTI0LjQtMS4yLTQ5LjMtMTAuMS03NC45YzguOC0wLjQsMTYuNCwxLjgsMTkuOSwxMi4yYzMsOC44LDguNywxMiwxOC40LDguMmMxOC42LTcuNSwzNy40LTExLjIsNTQuOSwxLjYKCWM5LjUtNy4yLDYuOC0xNS42LDQuNC0yMy44Yy0yLjQtOC4xLTkuNy02LjYtMTUuOS04LjFjLTcuNy0xLjgtMTctMy43LTIxLjktOWMtNy40LTgtMTUuNC0xMi4yLTI1LjUtMTIuOQoJYy0xNy4zLTEuMS0zMi02LjktNDIuNC0yMy4xYy0xLjMsMTMuNSw5LjcsMTguMiwxNi41LDI3Yy03LjEtMC43LTEyLjctMS0xOC4zLTEuOWMtMjYuMi00LjMtNDUuNi0xNy4zLTUzLjktNDMuNwoJYy0zLjQtMTAuNy0wLjYtMjEuOSw2LjgtMjdjMTEuNS03LjksMjMuNS01LjYsMzUuMS0wLjJDMzE5LjEsMTA0LjIsMzI4LDExMCwzMjYsMTI0Yy0xLjEsNy42LDIuMiwxMi45LDExLjIsMTMuOQoJYzIxLjUsMi4zLDQyLjEsNi44LDUzLDI4LjNjMi45LTEuNSw1LjMtMy4yLDcuMy00LjljLTQuNC0zLjYtMTMuMy0xMS4xLTE3LjMtMTVjLTAuNi0wLjYtMi4yLTEuNy00LTMuMWMtNy01LjItMTMuMy0xMC4xLTEzLjgtMTMuNAoJYy0wLjEtMC43LDAuMS0xLjQsMC41LTEuOWw0LjMsMC41Yy0zLjksMy42LDI1LjUsMjIuMiwzMi45LDMwLjNjOS40LTkuMiw2LjEtMTkuMS02LjgtMjguMWMtNi43LTQuNy0xNC40LTEwLjktMTYuNy0xOC4xCgljLTMuNS0xMS4xLTkuOC0xOC0xOC44LTI0LjNjLTYuMy00LjQtMTIuNy05LjctMTYuNS0xNi4xYy02LjYtMTEtMTEuMy0yMy4yLTE2LjYtMzQuM2MtMy45LDEzLjgsMy42LDI3LjIsNy42LDQwLjYKCWMtMjIuNi01LjEtMzAuOS0zMS4yLTU1LjktMzUuOGM4LjYsOC41LDE5LjUsMTIuNiwyMC45LDI0LjhjLTU1LjMtNi42LTYwLjktNi4zLTcyLjgsNC4yYzE1LjEtNC41LDI5LjItNi4zLDQ0LjEsMS4yCgljLTMyLjUsNS4yLTU2LjEsMjEuNC02My4xLDQyLjZjNS4zLTIuOSwxMC41LTUuOSwxNy4zLTkuN2MtMy44LDI3LjctOS44LDU0LjMsMTMuMyw3Ni44Yy0xLjYtOC40LTIuOS0xNS40LTMuNy0yMAoJYzYuNCw1LjMsMTMuNywxMS40LDIwLjgsMTcuNGMtNywzLjctMTMuOCw3LjMtMjAuNiwxMC45YzEyLjItMC43LDIzLjUtOCwzNy40LDFjLTI3LjMsNi4yLTQzLjIsMjIuOC01NC45LDQ3LjkKCWM2LjktNC44LDExLjUtOCwxNi4xLTExLjJjLTEyLjUsMjQuOS02LjEsMjguOC05LjEsNDkuMmMtNC4zLDE2LjQsMi43LTIuMy0xNC0xNS45Yy0xLjMtMS4xLTIuNy0yLTQtMi45CgljLTE2LjYtMTMuOC0xOS4xLTQwLjUtNC4xLTU0LjljLTAuNCw3LjQtMC43LDE0LjItMS4yLDIyLjVjMTUuNy0yMS43LDEyLjktNDIuNiwzLjEtNjQuM2MzLjMsMi4xLDYuNiw0LjEsMTIuNCw3LjcKCWMtOS4yLTE4LjUtMTkuNC0zMS44LTM3LjEtMzguMWM2LjYsMS44LDEzLjEsMy42LDIzLDYuM2MtMTYuMi0xNy4zLTMzLjMtMjItNTMuNy0yMS43YzcuMS00LjUsMTQuMi05LDIzLjYtMTUKCWMtMjAsMS41LTMzLjQsMTEuMy00Ni44LDIyLjFjLTEuOS0xOS02LjUtMjQuNi0yMC40LTI1LjZjOC45LDcuNyw5LjIsMTcsNC42LDI2LjJjLTQuMyw4LjktMTAuNCwxNi45LTE1LjEsMjUuNgoJYy0zLjIsNi01LjMsMTIuNi03LjcsMTljLTMsNy44LTYsMTUuNy04LjksMjMuNmMtMiw1LjQtNi4yLDExLjMtNS4yLDE2LjFjMSw0LjcsMy41LDkuMSw3LjcsMTEuM2MxLjEtNiw1LjctMTcuNCwxMS4xLTIzLjYKCWM1LTUuNyw5LjctOS43LDEwLjMtMTMuNGMwLjQtMi40LTEuMi01LjItMS4yLTUuM2w0LTEuM2MwLjYsMC40LDEuMyw0LDEuMiw2LjJjLTAuMSw0LjUtNS43LDkuNS0xMi4zLDE3CgljLTQuMyw0LjgtOC41LDE1LjItOS41LDIxLjZjMS43LDAuMywzLjcsMC40LDUuOCwwYzMuNy0xOC41LDE3LjMtMjcuNiwzMy4yLTM0LjNjOC0zLjQsMTIuNS03LjYsOC40LTE3Yy0zLjQtNy43LDAuMy0xMi42LDcuOC0xNS4zCgljMTEuNi00LjIsMjQuMiwzLjEsMjQuOCwxNi42YzAuNCwxMC0wLjksMjEuMS00LjksMzAuMWMtOS43LDIxLjktMTAuMSw0NC40LTIuMiw2My44YzAuMSwwLjIsMC4xLDAuMywwLjIsMC41CgljMC4xLDAuMywwLjIsMC42LDAuNCwwLjhjMC4zLDAuNiwwLjYsMS4zLDAuOSwxLjljMS4xLDIuNCwyLDQuNywyLjksNy4xbDAsMGMwLjIsMi41LDEsMjMtMTMuOCwzMC44YzAsMC0yMC4xLDkuOC0yNS45LDE1LjMKCWMtNS44LDUuNS0xMC42LDYtMTAuNiw2Yy0yLjksMC43LTE0LTQuNy0xNC4yLTguOWwwLDIuOWMwLDAtOC42LTcuOC0xMC4yLTEwLjZsMS4xLDMuNGMwLDAtOC4xLTYtOS44LTguMWwwLjgsMi40CgljMCwwLTEzLTYuNi0yMC44LDAuNWMwLDAsMTMuOCwwLDE2LjEsNi45YzAsMCwyLjksNC42LDQuOSw1LjRjMi4xLDAuOCwzLjQsMTAuMi02LjEsMTJjLTkuNSwxLjgtMTMuMy03LjMtMTMuMy03LjMKCXMtNS43LTcuNi0zLjItMTUuMmMwLDAtNy44LDEwLjktNC4yLDE1LjlsLTIuMy0wLjZjMCwwLDMuOCw2LjEsNS42LDcuNWgtMi41YzAsMCwyLjEsMS44LDYsMy4yYzAsMC0xMi40LDAuNy0xNS4xLTAuNgoJYy0yLjctMS4zLTkuMy01LjgtMTAuNy03LjljMCwwLDEuNiwxMC4yLDguNiwxNS4yaC0zbDMuNCwyLjFjMCwwLTkuMy0wLjYtMTMuMS0zLjhjMCwwLDUsNy4yLDExLjEsOC4xbC0yLjksMS40aDIwLjVsLTYuOSwyLjMKCWMwLDAsMjcuNSwwLjIsMzMuMy0xLjFsLTcuMSw0LjJjMCwwLDMwLjYtMC41LDQxLjYtOC45bC00LDYuMWwyOC4zLTE5LjdjLTAuNiwzLTIsOC4zLTAuNiw5LjZjMS43LTAuMiwxMC44LTE2LjYsMTUuNS0xOC44CgljMy4zLTEuNSwxMC45LTYuMiwxNy4xLDguMWwwLDBjMC4xLDAuMywwLjIsMC41LDAuMywwLjdjMi4zLDE1LjksMy44LDI1LjksMTAuMiwyNS45YzIwLjUsMCwxMTMuMywwLDEzMS42LDAKCWM0LjYtOC44LDguNy0xNy42LDEyLjEtMjYuNGMwLTAuMSwwLjEtMC4xLDAuMS0wLjJjNi4yLTE0LjQsMTMuOC05LjcsMTcuMS04LjJjNC43LDIuMiwxMy44LDE4LjUsMTUuNSwxOC44YzEuNC0xLjQsMC02LjYtMC42LTkuNgoJbDI4LjMsMTkuN2wtNC02LjFjMTEsOC41LDQxLjYsOC45LDQxLjYsOC45bC03LjEtNC4yYzUuOCwxLjQsMzMuMywxLjEsMzMuMywxLjFsLTYuOS0yLjNoMjAuNWwtMi45LTEuNAoJQzQ2OCwzNDguOCw0NzMuMSwzNDEuNiw0NzMuMSwzNDEuNnogTTM4NC4xLDE4NGwtMiwyLjdjLTUuOS00LjUtOS41LTIuMS05LjctMmwtMS41LTNDMzcxLjEsMTgxLjUsMzc3LjEsMTc4LjMsMzg0LjEsMTg0egoJIE0zNjcuMSwxMTEuMmwtMi45LDEuMmMtMS44LTQuMy04LTMuMi04LjEtMy4ybC0wLjItMy4yQzM1Ni4zLDEwNi4xLDM2NC43LDEwNS41LDM2Ny4xLDExMS4yeiBNMTAwLjcsMTU3LjhsLTMuMi0xLjIKCWMxLjUtMi42LDEuNi01LjksMTEuNC03bDAuMywzQzEwMywxNTMuMywxMDAuNSwxNTcsMTAwLjcsMTU3Ljh6IE0yOTIuMiwyODYuOGMtNi4yLTcuMi0xMi40LTE0LjItMTguNS0yMS40CgljLTE0LjMtMTctMTIuOC0zMS42LDQuNi00NS41YzMuNy0yLjksNy43LTUuNCwxMS41LThDMzA4LjUsMjMxLjYsMzA5LjMsMjYxLjUsMjkyLjIsMjg2Ljh6Ii8+Cjwvc3ZnPgo=
[cloudflare-badge]: https://img.shields.io/badge/Docs-pages-gray.svg?style=flat&logo=cloudflare&color=F26722
[cloudflare-link]: https://docs.pyrovelocity.net
[preprint-badge]: https://img.shields.io/badge/doi-10.1101/2022.09.12.507691v2-B31B1B
[preprint-link]: https://doi.org/10.1101/2022.09.12.507691
[pypi-badge]: https://img.shields.io/pypi/v/pyrovelocity.svg?logo=pypi&label=PyPI&color=F26722&logoColor=F26722
[pypi-link]: https://pypi.org/project/pyrovelocity/
[conda-forge-badge]: https://img.shields.io/conda/vn/conda-forge/pyrovelocity.svg?logo=conda-forge&label=conda-forge&color=F26722
[anaconda-link]: https://anaconda.org/conda-forge/pyrovelocity
[docker-badge]: https://img.shields.io/badge/docker-image-blue?logo=docker
[docker-link]: https://github.com/pinellolab/pyrovelocity/pkgs/container/pyrovelocity
[codecov-badge]: https://codecov.io/gh/pinellolab/pyrovelocity/branch/main/graph/badge.svg
[codecov-link]: https://codecov.io/gh/pinellolab/pyrovelocity
[black-badge]: https://img.shields.io/badge/code%20style-black-000000.svg
[black-link]: https://github.com/psf/black
[license-badge]: https://img.shields.io/badge/license-AGPL%203-gray.svg?color=F26722
[license-link]: https://spdx.org/licenses/
[pyro-ppl-link]: https://pyro.ai
[pyro-ppl-badge]: https://img.shields.io/badge/Pyro-PPL-gray.svg?color=F16722&logo=data:image/svg%2bxml;base64,PHN2ZyBpZD0iTGF5ZXJfMSIgZGF0YS1uYW1lPSJMYXllciAxIiB4bWxucz0iaHR0cDovL3d3dy53My5vcmcvMjAwMC9zdmciIHZpZXdCb3g9IjAgMCA1NDEgNDY0Ij48ZGVmcz48c3R5bGU+LmNscy0xe2ZpbGw6I2VjMWYyNDt9LmNscy0ye2ZpbGw6I2YxNjcyMjt9LmNscy0ze2ZpbGw6I2ZkY2MwODt9PC9zdHlsZT48L2RlZnM+PHRpdGxlPnB5cm8taG9yaXpvbnRhbC1jb2xvcjwvdGl0bGU+PHBhdGggY2xhc3M9ImNscy0xIiBkPSJNMTA0Ljk2MjQ0LDcuMDM1ODlxNjAuMDExNTUuMDg0MjcsMTIwLjAxNDQuMDM3ODJhODMuODQ2LDgzLjg0NiwwLDAsMSwxNy43NTk0NCwxLjg4NDIzYzIuNzEwMzMuNjE5MTUsNi4xMzkzMS43OTQzOSw2LjczOTExLDQuNjk1NTFhMTI5LjcyMDY0LDEyOS43MjA2NCwwLDAsMC0xOS42NDU1Ny4wMzY5Yy0yMi44MzU0OCwxLjM3Mi0zOS44MjYzNiwxNi41NzctMzkuOTAzNjQsMzkuNDgzNjgtLjIwNDY4LDYwLjY0MjI1LS45NDY4MSwxMjEuMzAyODkuMjU4MTUsMTgxLjkyMTczLjY3MDM4LDMzLjcyNTMxLTE0LjUzNzQ2LDU2Ljc4MDg1LTQxLjYyNTYsNzMuNTg5NjYtMjEuMTcxNjEsMTMuMTM3NDUtNDQuMjEyMTMsMTkuNjk3MzUtNjkuMjM4NjgsMTQuMDY3NDctMjcuMTQzODMtNi4xMDYxNy0zOS41OTg2Mi0yOS42NS0yOS4zMTI4Ni01NS41MDUxOSwxMC4xMzI2LTI1LjQ3MDIxLDMwLjAyNzA3LTQxLjEwMTgsNTQuNjA1LTUxLjM0MDU2LDEyLjQ4ODE2LTUuMjAyMzYsMjQuNzU3NC0xMC4zNjYzNiwzNC4xMTI3LTIwLjgxMzc5LDE3LjQxNzI0LTE5LjQ1MDYsMjQuMDQzMTItNDIuNjk0NCwyNC4zNjQzOC02OC4wOTAyMS4zMTI1OS0yNC43MTE4NC4wODgxMy00OS40MzA3OS4wNjE1NC03NC4xNDY2My0uMDIxMDctMTkuNTM0ODgtMTMuMTA1NjQtMzUuNzM2OTMtMzIuMzY5LTM4LjQwMDI5LTE3LjcxOC0yLjQ0OTcxLTM1LjA2MTYtLjUxNy01MC42MDExMyw5LjU1M0EzOS45MDUxNSwzOS45MDUxNSwwLDAsMCw3Mi40NzkyNCwzMC4xNjRjLTUuMzAwNDMsNS43NDQxOS00Ljc3OTQzLDcuOTU3LDIuMTUxNzgsMTEuNzg5ODEsMTguNTkxNTIsMTAuMjgwNTYsMjguMDM2NjUsMzEuNDMwOTMsMjMuNDk3LDUyLjYxNjM1LTQuMjMsMTkuNzQwNDYtMjIuMDU4MjksMzUuMjE3MjgtNDIuNjE0ODgsMzYuOTk0NjktMjAuNDk0MjksMS43NzItNDAuNzg2NDEtMTAuMzI1MjEtNDguNTUxMjktMjguOTQzODEtOS4xNzM1NC0yMS45OTYzNS42MDA1OS01NS42NDYsMjAuODcxNzYtNzEuODAwMDgsMTYuNjE0MjgtMTMuMjQsMzYuMDQ1MzItMTkuNDgwMjMsNTYuNjU2NzctMjIuODE3NDUiLz48cGF0aCBjbGFzcz0iY2xzLTIiIGQ9Ik0zMDYuODA3LDcuMDU4NTZjMzEuNzM1Mi4wMjEyOSw2My4zMTIwNy0uODI3MzgsOTUuMDQ3MjctLjg0ODY4LDEuNzQxLS4wMDExNiwyLjc5Mzg1LjA3NjM5LDUuMzE1NzMuMDc2MzZoNC41ODc4M2MyLjYwODQzLDIuMDY2NCw2LjE5NjM4LDIuMTAzLDguNzc3MTYsNC4xODkzLDIuOTc0MzQsMi40MDQzNSw1LjkxNCwzLjE4MjE5LDIuMjQ2NzcsNC4wNDk0OS05LjAzMzMtLjg2NjQ5LTIxLjE5NTA3LjczNzU2LTI5LjQzLDUuMDg3MkMzODEuNDc5LDI1Ljg4MzQyLDM3NS4zODgxMiwzNS43NjEsMzc0LjM4MDM0LDQ4Ljk4NDczcS0uMDU4MTQuNzYyMjUtLjA1OTg0LDEuNTI3NzdjLS4wODMyLDc3LjQ3NTkzLS4yMDcxMywxNTQuOTUyMzcuMDg0NzEsMjMyLjQyNzQ2LjA3Njk0LDIwLjQzMDE1LDE2LjU2MzIsMzUuNjQ4NzIsMzcuMTIxMTksMzUuODI1NDEsMi4yOS0uMTI1NjQsOS41NTMwNy0uNzIzNTcsMTEuODM3MDktLjUxNjI5LDQuMjYyNTUsMS4wOTYwNy0uOTQwNjQsMi42OTc2LTQuOTUyMzYsNS41MDM4Ny0zLjk1NTE0LDIuNzY2Ny04LjcyNTEzLDIuNDM0LTEzLjIyOTUyLDIuODM0MzgtMy43ODk5Mi4zMzY4Ni03LjYyNTg5LS4wNjQtMTEuNDM4NDIuMTY5YTEwNi42MTE0MSwxMDYuNjExNDEsMCwwLDEtMTUuNTg1NDUtMS4wMzc0OGMtMjAuNDc3NzIsMi42Njk0MS0zOS4yMDY0OC0zLjA5MTIxLTU2LjE5NjYyLTEzLjE4MzIyLTI0LjczMTY2LTE0LjY5MDMxLTM3LjA4MTMzLTM2LjkwNjYyLTM2LjQ0NTU5LTY1Ljk1NjgyLS4xNzE0MS0zLjI5MTc2LS40OTA0MS02LjU4MzQxLS40OTIxOC05Ljg3NTMyLS4wMzA4Ni01Ny4wMDAyOC4wNDktMTE0LjAwMDctLjA1NTY0LTE3MS4wMDA3OS0uMDQ2ODctMjUuNjA1ODktMTQuMDc2LTQ1LjA4MTQ2LTM1LjQ5MjMxLTUyLjA0OTMxLTcuMTE1NDItMy44MDc3My0xNS4xMDM4My00LjkwMi0yMi43NzY2NC02LjU2MzIzIi8+PHBhdGggY2xhc3M9ImNscy0zIiBkPSJNNTI0Ljg4OSw5Ni41OTYzNkM1MTEuODQwNDIsNDYuODYzLDQ3OS4zNzgsMTYuODQzMjksNDI3Ljk4ODEsOC4zMDEyM2MtMy4zNjQzNS0uNTU5MjEtMi45MDg1MS0uODkwNDMtOS43NzM1My0xLjYyMjQyLTEuNTIwNi0uMTYyMTItNi40NTY1Ny0uMzkyNTQtNi40NTY1Ny0uMzkyNTRoMGMxNy4yMTU0NSw4LjMwOTkyLDI3LjgzNzI4LDI3LjM3ODIsMjcuOTc3NjMsNDYuNTI4N3EuODM0NTgsMTEzLjg3MjY5LjAzMDUxLDIyNy43NDM1MmMtLjA5MTI3LDE0LjQ1ODQ3LTYuMTg3MjIsMjkuMDIxODktMTkuNzY1NjgsMzYuNTczMzRsLTI2LjI1NzE0LDkuNjI0NTdjMjAuOTUyNTEuNjYzNDUsNDEuNzEwMjkuNTM0LDYwLjk1NzU4LTkuODMwNjksMzEuNDk5ODQtMTYuOTYyNjEsNTIuMjQxMDYtNDMuMzE1MTIsNjUuNDc0NjctNzUuNzc5MTNDNTM5LjUyNzMxLDE5My42NzQ0OCw1MzcuNTc1MSwxNDQuOTQ4LDUyNC44ODksOTYuNTk2MzZaIi8+PHBhdGggY2xhc3M9ImNscy0zIiBkPSJNMzI4LjEyMjg5LDQ1OC4xNzM4OGMtMjMuOTEwNC0uMDgyOTEtNDcuODIxMjItLjExNy03MS43MzA2Mi0uMzE5MjQtMi41OTAxOC0uMDIxODUtNS42MjMyNywxLjM1NjE0LTcuNzg2NS0xLjQwOTc5LDE3LjU5MzgxLTMuMjkzODYsMjYuNzA4NzQtMTUuMzM1NzYsMzIuMDM0NjQtMzEuMTkyLDQuMDk0NTctMTIuMTkwMzcsNC45NDQ4OS0yNC44MzkyMyw0Ljk0Njg3LTM3LjU3OTgzcS4wMTAzLTY2Ljg2OTE2LjAwNzIzLTEzMy43MzgzYy4wMDEtMi43OTI1NC0uMjQ3NS01LjU4NS0uMTM2NjktOC4zNzcwOSwxLjIwNjc2LDEzLjc3NTIxLDQuODczNDgsMjYuOTk3OTIsMTIuMDk1ODMsMzguOTIzNjEsMTQuMjU0LDIzLjUzNjM1LDM2Ljc4ODYsMzMuOTQ0ODksNjIuNjMzNDIsMzguODE1MjgsNS45MzI1LDEuMTE4LDExLjk3OSwxLjYzMTIsMTcuOTcwNzcsMi40MjIzLTQuNTk4MDkuODM4NjMtMy4yMTgwNSw0LjUwNTI1LTMuMjUzLDcuMjcxMzEtLjI2MzMxLDIwLjc5ODQ5LS4zNzM1Nyw0MS41OTg4NC0uNjA5NDEsNjIuMzk3NzlhMTAxLjIxNzgxLDEwMS4yMTc4MSwwLDAsMS01LjIyMTQsMzEuMzYyNjFDMzYyLjIwMzE2LDQ0Ny4xMjMyNSwzNDkuNjYxMjgsNDU2LjY4MTM5LDMyOC4xMjI4OSw0NTguMTczODhaIi8+PC9zdmc+
[scvi-tools-link]: https://scvi-tools.org
[scvi-tools-badge]: https://img.shields.io/badge/scvi-tools-gray.svg?color=FCC64C&logo=data:image/svg%2bxml;base64,PD94bWwgdmVyc2lvbj0iMS4wIiBlbmNvZGluZz0idXRmLTgiPz4KPCEtLSBHZW5lcmF0b3I6IEFkb2JlIElsbHVzdHJhdG9yIDI0LjEuMCwgU1ZHIEV4cG9ydCBQbHVnLUluIC4gU1ZHIFZlcnNpb246IDYuMDAgQnVpbGQgMCkgIC0tPgo8c3ZnIHZlcnNpb249IjEuMSIgaWQ9IkxheWVyXzEiIHhtbG5zPSJodHRwOi8vd3d3LnczLm9yZy8yMDAwL3N2ZyIgeG1sbnM6eGxpbms9Imh0dHA6Ly93d3cudzMub3JnLzE5OTkveGxpbmsiIHg9IjBweCIgeT0iMHB4IgoJIHZpZXdCb3g9IjAgMCAyMS43IDIyLjgiIHN0eWxlPSJlbmFibGUtYmFja2dyb3VuZDpuZXcgMCAwIDIxLjcgMjIuODsiIHhtbDpzcGFjZT0icHJlc2VydmUiPgo8c3R5bGUgdHlwZT0idGV4dC9jc3MiPgoJLnN0MHtvcGFjaXR5OjAuNzc7ZmlsbDojRkJCNTE2O3N0cm9rZTojQzM4MzJCO3N0cm9rZS13aWR0aDowLjU7c3Ryb2tlLW1pdGVybGltaXQ6MTA7ZW5hYmxlLWJhY2tncm91bmQ6bmV3ICAgIDt9Cgkuc3Qxe29wYWNpdHk6MC43NTtmaWxsOiNDMzgzMkI7c3Ryb2tlOiNBNjU3MkE7c3Ryb2tlLXdpZHRoOjAuNzU7c3Ryb2tlLW1pdGVybGltaXQ6MTA7ZW5hYmxlLWJhY2tncm91bmQ6bmV3ICAgIDt9Cjwvc3R5bGU+CjxwYXRoIGNsYXNzPSJzdDAiIGQ9Ik0yMS40LDUuMWMwLjEsMS44LTEuMiwyLjEtMS43LDQuNGMtMC44LDMuOCwyLDUuNiwwLjksNy42cy0zLjUsMC40LTguMiwyLjZjLTMuNCwxLjUtMy45LDMuMS01LjksMi44CglzLTMuNC0yLjMtNC41LTRjLTAuOC0xLjMtMS45LTIuOS0xLjctNC44YzAuMS0xLjMsMC43LTEuOCwxLjYtMy4zYzIuOS00LjcsMS42LTYuNywzLjYtNy44czMuNiwwLjksNy0wLjNjMi4yLTAuOCwyLjQtMiw0LTIKCUMxOC44LDAuMiwyMS4yLDIuNywyMS40LDUuMXoiLz4KPGVsbGlwc2UgdHJhbnNmb3JtPSJtYXRyaXgoMC44MDkxIC0wLjU4NzYgMC41ODc2IDAuODA5MSAtNC4yNjIgOC44ODIxKSIgY2xhc3M9InN0MSIgY3g9IjExLjUiIGN5PSIxMSIgcng9IjQiIHJ5PSI0Ii8+Cjwvc3ZnPgo=
[tuple-badge]: https://img.shields.io/badge/Tuple%20❤️%20OSS-5A67D8?logo=tuple
[tuple-link]: https://tuple.app/github-badge

</div>