### 2024-04-06 nbs/concepts/inferenceindynamicalsystems/index.qmd
feat(nbs): update dimensional analysis Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Probabilistic inference in dynamical systems
toc: true
number-sections: true
highlight-style: pygments
csl: ../../bibstyle.csl
lightbox: auto
format:
#   nature-pdf:
#     natbibstyle: sn-nature
#     # classoption:
#     #   - lineno
#     cite-method: citeproc
#     keep-tex: true
  html:
    html-math-method: mathjax
    include-in-header: ../../mathjax.html
#     mainfont: Latin Modern Sans
#     code-fold: false
#     embed-resources: true
#   docx: default
  ipynb: default
format-links: [ipynb]
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
  # keep-ipynb: true
  # keep-md: true
author:
  - name: Pyrovelocity Team
    email: team@pyrovelocity.net
abstract: |
  This notebook provides a discussion and demonstration of one way to proceed in
  performing probabilistic inference in dynamical systems. In particular we prepare 
  and simulate a deterministic dynamical system and then perform probabilistic 
  inference on the latent variables of the system. This illustrates how to combine
  these two approaches, which is essential for understanding Pyro-Velocity.
keywords: [single-cell transcriptomics, probabilistic modeling, dynamical systems, RNA velocity]
bibliography: ../../references.bib
jupyter:
  jupytext:
    cell_metadata_filter: all
    cell_metadata_json: true
    notebook_metadata_filter: all
    text_representation:
      extension: .qmd
      format_name: quarto
      format_version: 1.0
      jupytext_version: 1.16.0
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
  language_info:
    name: python
  rise:
    scroll: true
    theme: black
  toc-autonumbering: true
  toc-showcode: false
  toc-showmarkdowntxt: false
---

## Model

### Description

Here we describe the model we will use to demonstrate probabilistic 
inference in dynamical systems. We begin with a candidate 
[effective theory](https://en.wikipedia.org/wiki/Effective_theory) 
[@Wells2012-qf;@Alon2019-ah] of gene transcription and splicing following 
[@Zeisel2011-nz;@La_Manno2018-lj;@Bergen2020-pj] for the model and 
[@Cantwell2002-dv] for the presentation of the analysis. In particular we show in 
@tbl-transcription-splicing-degradation the variables and parameters of the model
together with their units and rough order of magnitude estimates for their values 
[@Phillips2012-mi].

| Symbol   | Description                 | Units                         | $O(-)$ Estimate | Discussion                                                                          |
|----------|-----------------------------|-------------------------------|-----------------|-------------------------------------------------------------------------------------|
| $u$      | Concentration of pre-mRNA   | molecules/cell                | $10^0 - 10^4$   | Wide range reflecting low to high gene expression levels.                           |
| $s$      | Concentration of mRNA       | molecules/cell                | $10^0 - 10^5$   | Similar to $u$; depends on gene expression and stability of mRNA.                   |
| $t$      | Time                        | seconds ($s$) to hours ($hr$) | $10s$ - $24hr$  | Depends on the duration of the experimental observation.                            |
| $\alpha$ | Production rate of pre-mRNA | molecules/(cell·hr)           | $1 - 10^3$      | Based on variable gene transcription rates.                                         |
| $\beta$  | Splicing rate of pre-mRNA   | $hr^{-1}$                     | $0.1 - 1$       | Pre-mRNA to mRNA splicing rates, assuming rapid processing within the nucleus.      |
| $\gamma$ | Degradation rate of mRNA    | $hr^{-1}$                     | $0.1 - 1$       | mRNA half-lives can vary widely, but many are in the range of a few hours to a day. |

: Variables and parameters of the transcription-splicing-degradation model of [@Zeisel2011-nz]. {#tbl-transcription-splicing-degradation}

Given the state variables representing concentrations of pre-mRNA, $u$, 
and mRNA, $s$, we have the following ordinary differential equations

\begin{align}
\frac{du}{dt} & = \alpha - \beta u  \label{eq-dudt} \\
\frac{ds}{dt} & = \beta u - \gamma s  \label{eq-dsdt}
\end{align}

\eqref{eq-dudt} and \eqref{eq-dsdt} are usually presented with 
the concept that the parameter $\alpha$ could better account for the external 
inputs to the regulation of transcription of the gene if it were allowed to be 
a function of time, $\alpha(t)$. We will return to this point later, but 
for now we will assume that $\alpha$ and other parameters are constant.

### Dimensional analysis

It is common to perform a dimensional analysis of the model to ensure the
units of the variables and parameters are consistent and reduce the total number
of parameters by the number of dimensions. Intuitively, this folds or maps parameters
to equivalence classes of similar dynamics.  For reference, the procedure
we outline is also commonly referred to as Buckingham's $\Pi$ theorem.

@tbl-transcription-splicing-degradation together with (\ref{eq-dudt}-\ref{eq-dsdt})
contain essentially all the information we require to get started. The dimensions of the variables and parameters together with dimensionless ones are provided in @tbl-parameter-dimensions.

| Dimensioned Parameter | Relation to Rescaling Parameters | Fundamental Units        |
|-----------------------|--------------------------------------|--------------------------|
| Production Rate ($\alpha$) | $\alpha = U_0 \beta$             | molecules/(cell·time)    |
| Splicing Rate ($\beta$)    | Reference Scale for $t^*$        | $1/\text{time}$        |
| Degradation Rate ($\gamma$)| $\gamma = \gamma^* \beta$        | $1/\text{time}$        |
| Pre-mRNA Concentration ($u$)| $u = u^* U_0$                    | molecules/cell           |
| mRNA Concentration ($s$)    | $s = s^* U_0$                    | molecules/cell           |
| Time ($t$)                  | $t = t^* / \beta$                | time                     |

: Variables and parameters of the transcription-splicing-degradation model together with their fundamental units and their relations to dimensionless parameters. {#tbl-parameter-dimensions}

Note that molecules are dimensionless numbers while cells have units associated to their volume (i.e. the cube of a distance or length $L^3$). Usually we would just write volume but we retain cell to associate to the object that determines the relevant volume in this case. The dimensionless parameters are defined in @tbl-dimensionless-parameters.

| Dimensionless Variables and Parameters | Definition                     | Intuitive Description                                      |
|-------------------------|--------------------------------|------------------------------------------------------------|
| $u^{\ast}$ , $s^{\ast}$ | $u / U_0$, $s / U_0$             | Characteristic Scale of (pre-)mRNA concentration based on the balance between production and splicing rates. |
| $t^*$                 | $\beta t$                    | Characteristic time scale relative to the splicing rate. |
| $\gamma^*$            | $\gamma / \beta$             | Relative degradation rate, comparing the degradation rate of mRNA to the splicing rate, indicating the stability or turnover rate of mRNA relative to its production/splicing. |

: Dimensionless Parameters, definitions, and intuitive Descriptions {#tbl-dimensionless-parameters}

Combining variables and parameters we have six and we have two fundamental dimensions (time and length) we can eliminate to arrive at four essential variables and parameters. To derive these, we begin declaring the equations in python using [sympy](https://www.sympy.org/) and substitute the dimensionless variables and parameters to obtain the dimensionless equations.

```{python}
# | label: transcription-splicing-degradation-equations
from sympy import (
    diff,
    Eq,
    Function,
    init_printing,
    solve,
    symbols,
)
from IPython.display import display

init_printing(use_latex=True)

alpha, beta, gamma, gamma_star = symbols("alpha beta gamma gamma_star")
U0, T0 = symbols("U0 T0")
t, t_star = symbols(
    "t t_star",
)

u = Function("u")(t)
s = Function("s")(t)
u_star = Function("u_star")(t_star)
s_star = Function("s_star")(t_star)

du_dt = Eq(diff(u, t), alpha - beta * u)
ds_dt = Eq(diff(s, t), beta * u - gamma * s)
display(du_dt)
display(ds_dt)
```

and define dimensionelss variables and parameters as

```{python}
# | label: transcription-splicing-degradation-dimensionless
U0_eq = Eq(U0, alpha / beta)
T0_eq = Eq(T0, 1 / beta)

u_dimless = Eq(u_star, u / U0)
s_dimless = Eq(s_star, s / U0)
t_dimless = Eq(t_star, beta * t)
```

Now we change the variables in the equations to the dimensionless ones

```{python}
du_dt_dimless = Eq(
    beta * diff(u_star * alpha / beta, t_star),
    du_dt.rhs.subs(
        {
            u: u_star * alpha / beta,
            s: s_star * alpha / beta,
            t: t_star / beta,
            gamma: gamma_star * beta,
        }
    ),
).simplify()
du_dt_dimless
```

and

```{python}
ds_dt_dimless = Eq(
    beta * diff(s_star * alpha / beta, t_star),
    ds_dt.rhs.subs(
        {
            u: u_star * alpha / beta,
            s: s_star * alpha / beta,
            t: t_star / beta,
            gamma: gamma_star * beta,
        }
    ),
).simplify()

ds_dt_dimless
```

We see that we correctly arrive at the equivalent dimensionless system of equations

\begin{align}
\frac{du^{\ast}}{dt^{\ast}} & = 1 - u^{\ast}  \label{eq-dustardtstar} \\
\frac{ds^{\ast}}{dt^{\ast}} & = u^{\ast} - \gamma^{\ast} s^{\ast}  \label{eq-dsstardtstar}
\end{align}

in \eqref{eq-dustardtstar} and \eqref{eq-dsstardtstar}, which contains the $6 - 2 = 4$ variables and parameters: $u^{\ast}, s^{\ast}, t^{\ast}, \gamma^{\ast}$. It is trivial to solve this system analytically, but in more general cases this will not be possible. We can now proceed to simulate the system to characterize the possible solution sets and then perform probabilistic inference on the latent variables.