### 2024-04-10 nbs/concepts/inferenceindynamicalsystems/index.qmd
feat(nbs): add inference model description Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Probabilistic inference in dynamical systems
toc: true
number-sections: true
highlight-style: pygments
csl: ../../bibstyle.csl
lightbox: auto
format:
#   nature-pdf:
#     natbibstyle: sn-nature
#     # classoption:
#     #   - lineno
#     cite-method: citeproc
#     keep-tex: true
  html:
    html-math-method: mathjax
    include-in-header: ../../mathjax.html
#     mainfont: Latin Modern Sans
#     code-fold: false
#     embed-resources: true
#   docx: default
  ipynb: default
format-links: [ipynb]
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
  # keep-ipynb: true
  # keep-md: true
author:
  - name: Pyrovelocity Team
    email: team@pyrovelocity.net
abstract: |
  This notebook provides a discussion and demonstration of one way to proceed in
  performing probabilistic inference in dynamical systems. In particular we prepare 
  and simulate a deterministic dynamical system and then perform probabilistic 
  inference on the latent variables of the system. This illustrates how to combine
  these two approaches, which is essential for understanding Pyro-Velocity.
keywords: [single-cell transcriptomics, probabilistic modeling, dynamical systems, RNA velocity]
bibliography: ../../references.bib
jupyter:
  jupytext:
    cell_metadata_filter: all
    cell_metadata_json: true
    notebook_metadata_filter: all
    text_representation:
      extension: .qmd
      format_name: quarto
      format_version: 1.0
      jupytext_version: 1.16.0
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
  language_info:
    name: python
  rise:
    scroll: true
    theme: black
  toc-autonumbering: true
  toc-showcode: false
  toc-showmarkdowntxt: false
---

## Model

### Description

Here we describe the model we will use to demonstrate probabilistic 
inference in dynamical systems. We begin with a candidate 
[effective theory](https://en.wikipedia.org/wiki/Effective_theory) 
[@Wells2012-qf;@Alon2019-ah] of gene transcription and splicing following 
[@Zeisel2011-nz;@La_Manno2018-lj;@Bergen2020-pj] for the model and 
[@Cantwell2002-dv] for the presentation of the analysis. In particular we show in 
@tbl-transcription-splicing-degradation the variables and parameters of the model
together with their units and rough order of magnitude estimates for their values 
[@Phillips2012-mi].

| Symbol   | Description                 | Units                         | $O(-)$ Estimate  | Note                                                                          |
| -------- | --------------------------- | ----------------------------- | ---------------- | ----------------------------------------------------------------------------- |
| $u$      | Number of pre-mRNA          | molecules/cell                | $10^0 - 10^4$    | Wide range accounts for low to high gene expression levels.                   |
| $s$      | Number of mRNA              | molecules/cell                | $10^0 - 10^5$    | Similar to $u$; depends on gene expression and stability of mRNA.             |
| $t$      | Time                        | seconds ($s$) to hours ($hr$) | $5s$ - $48hr$    | Depends on the duration of the experimental observation.                      |
| $\alpha$ | Production rate of pre-mRNA | molecules/(cell·hr)           | $10^0 - 10^3$    | Many transcripts are produced at rates in the range or 1 to 1000 per hour.    |
| $\beta$  | Splicing rate of pre-mRNA   | $hr^{-1}$                     | $10^{-1} - 10^2$ | Many Pre-mRNA to mRNA splicing rates are in the range of 1 minute to 6 hours. |
| $\gamma$ | Degradation rate of mRNA    | $hr^{-1}$                     | $10^{-2} - 10^0$ | Many mRNA half-lives are in the range of a half-hour to a day.                |

: Variables and parameters of the transcription-splicing-degradation model of [@Zeisel2011-nz] with order of magnitude estimates for ranges based on [@Phillips2012-mi] or references therein. {#tbl-transcription-splicing-degradation}

Given state variables representing concentrations of pre-mRNA, $u$, 
and mRNA, $s$, we have the following ordinary differential equations 
taken from [@Zeisel2011-nz] (please refer again to 
@tbl-transcription-splicing-degradation for the meaning of the variables 
and parameters),

\begin{align}
\frac{du}{dt} & = \alpha - \beta u  \label{eq-dudt}, \\
\frac{ds}{dt} & = \beta u - \gamma s  \label{eq-dsdt},
\end{align}

which proposes a mean-field model for the dynamics of transcription and 
splicing on a continuous state space. It is important to note this is only
one among an eventual ensemble of models that, with appropriate analysis
and inference, may be organized into a hierarchy of such models according
to their domains of validity relative to the scales involved in and resolution 
of observations of the relevant phenomenon [@Wells2012-qf].
Of course, in the context of single gene transcription without explicit modeling
of interactions, this ensemble of models would include the various adaptations and
extensions of the so-called random telegraph model and the experimental 
justification for its consideration 
[@Ham2020-ij;@Cao2020-xv;@Bohrer2021-qk].
As an example of a minimal extension, \eqref{eq-dudt} and \eqref{eq-dsdt} are 
usually presented with the concept that the parameter $\alpha$ could better 
account for the external inputs to the regulation of transcription of the gene 
if it were allowed to be a function of time, $\alpha(t)$. However, because this
complicates or eliminates analytical tractability, and thus maximum likelihood
inference, approximation of this function may be piecewise constant 
[@Bergen2020-pj]. We will return to 
this point later, but, for the purpose of simply illustrating the relationsips 
among analyzing, simulating, and performing inference upon the parameters of a 
dynamical system, we will assume that $\alpha$ and other parameters are 
constant values.

### Dimensional analysis

It is common to perform a dimensional analysis of dynamical models to ensure the
units of the variables and parameters are consistent and reduce the total number
of parameters by the number of dimensions of the associated dilation group 
symmetry [@Pegoraro1973-tt]. Intuitively, this maps sets of parameters
to associated equivalence classes of similar dynamics. For reference, in addition to 
[dimensional analysis](https://en.wikipedia.org/wiki/Dimensional_analysis) and 
[nondimensionalization](https://en.wikipedia.org/wiki/Nondimensionalization), 
the procedure we make use of depends on what is frequently 
referred to as the 
[Buckingham $\Pi$ theorem](https://en.wikipedia.org/wiki/Buckingham_%CF%80_theorem) 
[@Cantwell2002-dv].

@tbl-transcription-splicing-degradation together with (\ref{eq-dudt}-\ref{eq-dsdt})
contain essentially all the information we require to get started. 
The dimensions of the variables and parameters together with dimensionless 
ones are provided in @tbl-parameter-dimensions.

| Dimensioned Parameter        | Relation to Rescaling Parameters | Fundamental Units     |
| ---------------------------- | -------------------------------- | --------------------- |
| Production Rate ($\alpha$)   | $\alpha = U_0 \beta$             | molecules/(cell·time) |
| Splicing Rate ($\beta$)      | Reference Scale for $t^*$        | $1/\text{time}$       |
| Degradation Rate ($\gamma$)  | $\gamma = \gamma^* \beta$        | $1/\text{time}$       |
| Pre-mRNA Concentration ($u$) | $u = u^* U_0$                    | molecules/cell        |
| mRNA Concentration ($s$)     | $s = s^* U_0$                    | molecules/cell        |
| Time ($t$)                   | $t = t^* / \beta$                | time                  |

: Variables and parameters of the transcription-splicing-degradation model 
together with their fundamental units and their relations to dimensionless 
parameters. {#tbl-parameter-dimensions}

Note that molecules are dimensionless numbers while cells have units associated 
to their volume (i.e. the cube of a distance or length $L^3$). Usually we would 
just write volume but we retain "cell" to associate to the object that determines 
the relevant volume in this case. The dimensionless parameters are defined in 
@tbl-dimensionless-parameters.

| Dimensionless Variables and Parameters | Definition           | Description                                                                                                                                                                    |
| -------------------------------------- | -------------------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ |
| $u^{\ast}$ , $s^{\ast}$                | $u / U_0$, $s / U_0$ | Characteristic scale of (pre-)mRNA concentration based on the balance between production and splicing rates.                                                                   |
| $t^*$                                  | $\beta t$            | Characteristic time scale relative to the splicing rate.                                                                                                                       |
| $\gamma^*$                             | $\gamma / \beta$     | Relative degradation rate, comparing the degradation rate of mRNA to the splicing rate, indicating the stability or turnover rate of mRNA relative to its production/splicing. |

: Dimensionless Parameters, definitions, and intuitive Descriptions 
{#tbl-dimensionless-parameters}

Given our previously defined parameter range for the splicing rate $\beta$, 
we can interpret the dimensionless parameter $t^{\ast}$ according to @tbl-dimensionless-time-interpretation.

| $\beta$ (hr$^{-1}$) | 1 unit of $t^*$ (hr) | 10 units of $t^*$ (hr) | 100 units of $t^*$ (hr) | 1 unit of $t^{\ast}$ | 10 units of $t^{\ast}$ | 100 units of $t^{\ast}$ |
| ------------------- | -------------------- | ---------------------- | ----------------------- | -------------------- | ---------------------- | ----------------------- |
| $10^{-1}$           | 10                   | 100                    | 1000                    | 1                    | 10                     | 100                     |
| $10^0$              | 1                    | 10                     | 100                     | 1                    | 10                     | 100                     |
| $10^1$              | 0.1                  | 1                      | 10                      | 1                    | 10                     | 100                     |
| $10^2$              | 0.01                 | 0.1                    | 1                       | 1                    | 10                     | 100                     |

: This table interprets the dimensionless time $t^{\ast}$ across various
splicing rates $\beta$ (hr$^{-1}$), demonstrating how real time and expected
splicing events scale with changes in $\beta$. Each unit of $t^{\ast}$
represents a normalized measure of time relative to the splicing rate, with the
final three columns indicating the **expected number of splicing events per
molecule of pre-mRNA** for 1, 10, and 100 units of $t^{\ast}$ is indeed the
meaning of time units for this scaling rule. These columns confirm that the
dimensionless time directly reflects the expected splicing activity, endowing
otherwise meaningless or difficult to interpret numbers into a universally
applicable framework based on scales intrinsic to a given instance of the model.
{#tbl-dimensionless-time-interpretation}

Combining variables and parameters gives six total. We have two fundamental 
dimensions (time and length) we can eliminate to arrive at four essential 
variables and parameters. To derive these, we begin declaring the equations 
in python using [sympy](https://www.sympy.org/) and substitute the 
dimensionless variables and parameters to obtain the dimensionless equations.

```{python}
#| label: transcription-splicing-degradation-equations
#| code-fold: true
from sympy import (
    diff,
    Eq,
    Function,
    init_printing,
    solve,
    symbols,
)
from IPython.display import display

init_printing(use_latex=True)

alpha, beta, gamma, gamma_star = symbols("alpha beta gamma gamma_star")
U0, T0 = symbols("U0 T0")
t, t_star = symbols(
    "t t_star",
)

u = Function("u")(t)
s = Function("s")(t)
u_star = Function("u_star")(t_star)
s_star = Function("s_star")(t_star)

du_dt = Eq(diff(u, t), alpha - beta * u)
ds_dt = Eq(diff(s, t), beta * u - gamma * s)
display(du_dt)
display(ds_dt)
```

and define dimensionless variables and parameters as

```{python}
# | label: transcription-splicing-degradation-dimensionless
U0_eq = Eq(U0, alpha / beta)
T0_eq = Eq(T0, 1 / beta)

u_dimless = Eq(u_star, u / U0)
s_dimless = Eq(s_star, s / U0)
t_dimless = Eq(t_star, beta * t)
```

Now we change the variables in the equations to the dimensionless ones

```{python}
#| label: symbolic-tsd-dudt-dimensionless-equation
#| code-fold: true
du_dt_dimless = Eq(
    beta * diff(u_star * alpha / beta, t_star),
    du_dt.rhs.subs(
        {
            u: u_star * alpha / beta,
            s: s_star * alpha / beta,
            t: t_star / beta,
            gamma: gamma_star * beta,
        }
    ),
).simplify()
du_dt_dimless
```

and

```{python}
#| label: symbolic-tsd-dsdt-dimensionless-equation
# | code-fold: true
ds_dt_dimless = Eq(
    beta * diff(s_star * alpha / beta, t_star),
    ds_dt.rhs.subs(
        {
            u: u_star * alpha / beta,
            s: s_star * alpha / beta,
            t: t_star / beta,
            gamma: gamma_star * beta,
        }
    ),
).simplify()

ds_dt_dimless
```

We see that we correctly arrive at the equivalent dimensionless system of 
equations

\begin{align}
\frac{du^{\ast}}{dt^{\ast}} & = 1 - u^{\ast}  \label{eq-dustardtstar}, \\
\frac{ds^{\ast}}{dt^{\ast}} & = u^{\ast} - \gamma^{\ast} s^{\ast},  
  \label{eq-dsstardtstar}
\end{align}

in \eqref{eq-dustardtstar} and \eqref{eq-dsstardtstar}, which contains the 
$6 - 2 = 4$ variables and parameters: $u^{\ast}, s^{\ast}, t^{\ast}, \gamma^{\ast}$.

We'll solve the dimensionless system analytically, focusing on what can be 
understood about the system's dynamics.
Given specific initial conditions $u^{\ast}(0) = u^{\ast}_0$ and 
$s^{\ast}(0) = s^{\ast}_0$, which we restore to arbitrary $t^{\ast}_0$ 
at the end of the derivation, the first equation

$$
\frac{du^{\ast}}{dt^{\ast}} = 1 - u^{\ast},
$$

can be solved using separation of variables

$$
u^{\ast}(t^{\ast}) = C_1 e^{-t^{\ast}} + 1.
$$

$C_1$ is of course the constant determined by the initial conditions, which can 
be shown to be $C_1 = u^{\ast}_0 - 1,$ giving us

$$
u^{\ast}(t^{\ast}) = (u^{\ast}_0 - 1) e^{-t^{\ast}} + 1.
$$

This equation describes how the concentration of pre-mRNA, in units of the 
characteristic concentration scale, $U_0 = \alpha/\beta$, evolves over the 
timescale characterized by the inverse of the splicing rate, $1/\beta$. 
The term $e^{-t^{\ast}}$ indicates that any deviation of $u^{\ast}$ from 
its steady state $1$ decays exponentially over time.

Now, we substitute the expression for $u^{\ast}(t^{\ast})$ into the 
differential equation for mRNA concentration

$$
\frac{ds^{\ast}}{dt^{\ast}} = u^{\ast} - \gamma^{\ast} s^{\ast}
$$

which gives

$$
\frac{ds^{\ast}}{dt^{\ast}} = [(u^{\ast}_0 - 1) e^{-t^{\ast}} + 1] - 
  \gamma^{\ast} s^{\ast}
$$

This equation is a non-homogeneous first-order linear ODE for $s^{\ast}$. 
To solve it, we can use an integrating factor $e^{\gamma^{\ast} t^{\ast}}$, 
which gives

$$
e^{\gamma^{\ast} t^{\ast}} \frac{ds^{\ast}}{dt^{\ast}} + 
  \gamma^{\ast} e^{\gamma^{\ast} t^{\ast}} s^{\ast} = 
  (u^{\ast}_0 - 1) e^{(\gamma^{\ast}-1) t^{\ast}} + 
  e^{\gamma^{\ast} t^{\ast}}
$$

Integrating both sides with respect to $t^{\ast}$ yields $s^{\ast}(t^{\ast})$

$$
s^{\ast}(t^{\ast}) = C_2 e^{-\gamma^{\ast} t^{\ast}} + 
  \frac{u^{\ast}_0 - 1}{\gamma^{\ast} - 1} e^{-t^{\ast}} + 
  \frac{1}{\gamma^{\ast}}
$$

where $C_2$ is another integration constant determined by initial conditions. 
Incorporating the initial condition  $s^{\ast}(0) = s^{\ast}_0$ into the 
solution for $s^{\ast}(t^{\ast})$ to determine $C_2$

$$
C_2 = s^{\ast}_0 - \frac{u^{\ast}_0 - 1}{\gamma^{\ast} - 1} - 
  \frac{1}{\gamma^{\ast}}.
$$

With $C_2$ determined, we define

$$
\xi = \frac{u^{\ast}_0 - 1}{\gamma^{\ast} - 1}.
$$

The complete solutions for the dimensionless system for $\gamma^{\ast} \neq 1$ 
are then given by

\begin{align}
u^{\ast}(t^{\ast}) & = \underbrace{1}_{\text{Steady State}} + 
  \underbrace{(u^{\ast}_0 - 1) e^{-t^{\ast}}}_{\text{Transient Component}}, 
  \label{eq-ustar-dimless} \\
s^{\ast}(t^{\ast}) & = \underbrace{\frac{1}{\gamma^{\ast}}}_{\text{Steady State}} + 
  \underbrace{\left( s^{\ast}_0 - \xi - \frac{1}{\gamma^{\ast}} \right) 
  e^{-\gamma^{\ast} t^{\ast}} + \xi e^{-t^{\ast}}}_{\text{Transient Components}}.  
  \label{eq-sstar-dimless}
\end{align}

For the case $\gamma^{\ast} = 1$, we can solve \eqref{eq-dustardtstar} and 
\eqref{eq-dsstardtstar} to find the solution for $u^{\ast}$ is unchanged
and $s^{\ast}$ is then given by

$$
s^{\ast}(t^{\ast}) = 1 + (s^{\ast}_0 - 1) e^{-t^{\ast}} + 
  (u^{\ast}_0 - 1) t^{\ast} e^{-t^{\ast}}.
$$

Note we can restore an arbitrary initial timepoint putting 
$t^{\ast} \rightarrow t^{\ast} - t^{\ast}_0$.

The dimensionless **pre-mRNA dynamics** $u^{\ast}(t^{\ast})$
given in \eqref{eq-ustar-dimless} 
settles to its steady state value of 1 at rate 1, regardless of the initial 
concentration, reflecting the balance between transcription and splicing 
rates. Note again that this does not refer to an arbitrary value of $1$ but 
rather represents the equivalence class of ratios $\alpha/\beta$.

The dimensionless **mRNA dynamics** $s^{\ast}(t^{\ast})$ 
given in \eqref{eq-sstar-dimless} reveals a 
slightly more complicated balance of forces. The first transient term shows 
an exponential decay influenced by $\gamma^{\ast}$,the degradation rate 
relative to the splicing rate, when $\gamma^{\ast}$ is not precisely $1$. 
The second transient term reflects how the initial deviation in pre-mRNA 
concentration from its steady state affects mRNA levels over time. The steady 
state of $s^{\ast}$ simply reflects the balance of the splicing rate, which is 
the only source of mRNA production in this model, and the degradation rate 
also given, just as in the case of the pre-mRNA, in units of the transcription 
rate relative to the splicing rate.

The analytical solutions to the dimensionless system reveal how the pre-mRNA 
and mRNA concentrations evolve over time towards their steady states. The 
exponential terms highlight the system's inherent timescales: $u^{\ast}$ relaxes 
to its steady state independently, while $s^{\ast}$’s dynamics are coupled to 
both $u^{\ast}$ and its own degradation rate. While this system is extremely
simplified relative to well-known biology, as a stylized model it still
indicates the interplay among transcription, splicing, and degradation rates.

While it is trivial to solve this system analytically, in more general cases this
 will not be possible. 
So we will proceed to simulate the system to characterize the possible solution 
sets and then perform probabilistic inference on the latent variables.

## Data

If we are given samples of pre-mRNA and mRNA counts along with
the sampling time of each cell, we can describe the data set 
$\mathcal{D}$ consisting of pre-mRNA and mRNA counts for a number 
of genes $G$ across a number of cells $N$

\begin{align}
\mathcal{D} = \left\{ (t_j, u_{ij}, s_{ij}) \mid 
  i \in \{1, \ldots, G\}, j \in \{1, \ldots, N\} \right\} \label{eq-dataset},
\end{align}

where

- $t_j$ represents the sampling time for cell $j$,
- $u_{ij}$ represents the count of pre-mRNA for gene $i$ in cell $j$,
- $s_{ij}$ represents the count of mRNA for the same gene $i$ in the same cell $j$,
- $G$ is the total number of genes in the study,
- $N$ is the total number of cells sampled.

Roughly speaking, the inference problem involves

1. **Normalization**: Transforming the observed counts $u_{ij}$ and $s_{ij}$ to 
   their dimensionless counterparts $u^{\ast}_{ij}$ and $s^{\ast}_{ij}$ using 
   a reasonable concentration scale $U_0$.
2. **Model Fitting**: Applying statistical methods to estimate the latent 
   variables 
   $\Theta = \left( \gamma^{\ast}, u^{\ast}_0, s^{\ast}_0 \right)$ 
   that best fit the observed data.
3. **Evaluation**: Assessing the fit of the model and the estimated parameters'
   biological plausibility and consistency across different genes and cells.

We will return later to the alternative version of this problem where $t_j$ 
are not observed and must be inferred along with the latent variables 
$\Theta = \left( \gamma^{\ast}, u^{\ast}_0, s^{\ast}_0, t^{\ast} \right)$.

This representation of the dataset and the associated objective of 
statistical learning provides a solid foundation for applying probabilistic
modeling techniques to calibrate an inference procedure and evaluate the 
plausibility of this model in such a manner that we can eventually compare 
multiple candidate models to one another.

## Simulation

We will primarily focus on simulating the system, since this will generalize
to more complicated models. However, we will confirm our simulations for this
first and simplest model recapitulate its analytical solution derived above.

As a simple example, we will simulate the system for a single set of initial
conditions and parameters.

```{python}
# | label: transcription-splicing-degradation-simulation
import diffrax
from pyrovelocity.models import solve_transcription_splicing_model
from pyrovelocity.models import solve_transcription_splicing_model_analytical
from pyrovelocity.logging import configure_logging
from jax import numpy as jnp

logger = configure_logging("nbs")

ts0 = jnp.linspace(0.1, 4.0, 40)
ts1 = jnp.linspace(4.0 + (10.0 - 4.0) / 20, 10.0, 20)
ts = jnp.concatenate([ts0, ts1])
initial_state = jnp.array([0.1, 0.1])
params = (1.00,)
colormap_name = "RdBu"

solution = solve_transcription_splicing_model(
    ts,
    initial_state,
    params,
)
analytical_solution = solve_transcription_splicing_model_analytical(
    ts,
    initial_state,
    params,
)
analytical_simulation_error = diffrax.Solution(
    t0=ts[0],
    t1=ts[-1],
    ts=ts,
    ys=analytical_solution.ys - solution.ys,
    interpolation=None,
    stats={},
    result=diffrax.RESULTS.successful,
    solver_state=None,
    controller_state=None,
    made_jump=None,
)

logger.info(
    f"\nTrajectory preview times:\n{solution.ts[:3]}\n"
    f"\nSimulated trajectory values:\n{solution.ys[:5]}\n"
    f"\nAnalytical trajectory values:\n{analytical_solution.ys[:5]}\n\n"
    f"\nTrajectory preview times:\n{solution.ts[-3:]}\n"
    f"\nSimulated trajectory values:\n{solution.ys[-3:]}\n"
    f"\nAnalytical trajectory values:\n{analytical_solution.ys[-3:]}\n\n"
)
```

We can visualize the results of the simulation by plotting the pre-mRNA and mRNA
concentrations over time as shown in 
@fig-transcription-splicing-degradation-plot-simulation-trajectory-1,

```{python}
# | label: fig-transcription-splicing-degradation-plot-simulation-trajectory-1
# | code-fold: true
# | fig-cap: Simulated trajectories of pre-mRNA and mRNA concentrations over time for the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_trajectories

plot_deterministic_simulation_trajectories(
    solution=solution,
    title_prefix="TSD Model Simulated",
    colormap_name=colormap_name,
)
```

We see that if we plot the analytical solution, we get the same result as shown 
in @fig-transcription-splicing-degradation-plot-analytical-trajectory-1 to 
within the expected error of the numerical simulation ($O(10^{-5} - 10^{-8})$).

```{python}
# | label: fig-transcription-splicing-degradation-plot-analytical-trajectory-1
# | code-fold: true
# | fig-cap: Absolute error in analytical vs simulated trajectories of pre-mRNA and mRNA concentrations over time for the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_trajectories

plot_deterministic_simulation_trajectories(
    solution=analytical_simulation_error,
    title_prefix="TSD Model Analytical-Simulation Error",
    colormap_name=colormap_name,
)
```

The same simulated trajectories from 
@fig-transcription-splicing-degradation-plot-simulation-trajectory-1 are shown 
in the phase space given by $u^{\ast} \otimes s^{\ast}$ in 
@fig-transcription-splicing-degradation-plot-simulation-phase-portrait-1.

```{python}
# | label: fig-transcription-splicing-degradation-plot-simulation-phase-portrait-1
# | code-fold: true
# | fig-cap: Phase portrait of the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_phase_portrait

plot_deterministic_simulation_phase_portrait(
    solution=solution,
    title_prefix="TSD Model",
    colormap_name=colormap_name,
)
```

## Inference

Now that we have illustrated how to simulate a prototypical system, we can
proceed to define a probabilistic model and perform inference on its latent
variables given data like \eqref{eq-dataset}. In general we should consider multiple
models with different assumptions and levels of complexity, but for now we will
focus on one of the simpler models to illustrate the process. In particular, in
most cases, when performing inference on the latent variables of a dynamical
system, we assume that the observation time points are fixed or otherwise
observed. We will begin with this assumption and then proceed to discuss how to
relax it.

In the case where we can indeed estimate analytical solutions to the system, we
can consider a model with the structure proposed in 
(\ref{eq-init-conds-priors-gene}--\ref{eq-s-obs-cell}) where, given a number of 
genes $G$ and cells $N$, for each
$(i,j) \in \{1, \ldots, G\} \otimes \{1, \ldots, N\}$ we have

\begin{align}
    u^{\ast}_{0i}, s^{\ast}_{0i} &\sim 
      \text{LogNormal}(\mu_{0}, \sigma_{0}^2), 
      \label{eq-init-conds-priors-gene} \\

    \gamma^{\ast}_i &\sim 
      \text{LogNormal}(\mu_{\gamma}, \sigma_{\gamma}^2),
      \label{eq-gamma-prior-gene} \\

    \sigma_{ui}, \sigma_{si} &\sim 
      \text{HalfNormal}(\mu_{\sigma}, \sigma_{\sigma}^2),
      \label{eq-noise-std-priors-gene} \\

    u^{\ast}_{ij} &= 
      1 + (u^{\ast}_{0i} - 1) \cdot e^{-t^{\ast}_j},
      \label{eq-u-star-model-cell} \\

    s^{\ast}_{ij} &= 
      \begin{cases}
        \frac{1}{\gamma^{\ast}_i} + 
        \left( s^{\ast}_{0i} - \xi_i - \frac{1}{\gamma^{\ast}_i} \right) \cdot e^{-\gamma^{\ast}_i t^{\ast}_j} + 
        \xi_i \cdot e^{-t^{\ast}_j},& \gamma^{\ast}_i \neq 1 \\

        1 + (s^{\ast}_{0i} - 1) e^{-t^{\ast}_j} + 
          (u^{\ast}_{0i} - 1) t^{\ast}_j e^{-t^{\ast}_j},& \gamma^{\ast}_i = 1 \\
      \end{cases}, \label{eq-s-star-model-cell} \\

    \hat{u}^{\ast}_{ij} &\sim \text{Normal}(u^{\ast}_{ij}, \sigma_{ui}^2), 
    \label{eq-u-obs-cell} \\

    \hat{s}^{\ast}_{ij} &\sim \text{Normal}(s^{\ast}_{ij}, \sigma_{si}^2). 
    \label{eq-s-obs-cell}
\end{align}

The model description in (\ref{eq-init-conds-priors-gene}--\ref{eq-s-obs-cell}) 
can be associated to the plate diagram in 
@fig-transcription-splicing-degradation-graphical-model-1.

```{python}
# | label: fig-transcription-splicing-degradation-graphical-model-1
# | code-fold: true
# | fig-cap: Graphical model for the transcription-splicing-degradation model.
import daft
import matplotlib.pyplot as mpl

mpl.rcParams["font.family"] = "serif"
mpl.rcParams["font.size"] = 16
mpl.rcParams["text.usetex"] = True

pgm = daft.PGM(line_width=1.2)

# hyperparameters
pgm.add_node("mu_init", r"$\mu_{0}$", 0.5, 6, fixed=True)
pgm.add_node("sigma_init", r"$\sigma_{0}^2$", 1.5, 6, fixed=True)
pgm.add_node("mu_gamma", r"$\mu_{\gamma}$", 2.5, 6, fixed=True)
pgm.add_node("sigma_gamma", r"$\sigma_{\gamma}^2$", 3.5, 6, fixed=True)
pgm.add_node("mu_sigma", r"$\mu_{\sigma}$", 4.5, 6, fixed=True)
pgm.add_node("sigma_sigma", r"$\sigma_{\sigma}^2$", 5.5, 6, fixed=True)

# latent variables for gene-specific parameters
pgm.add_node("u_star_0i", r"$u^{\ast}_{0i}$", 1, 5)
pgm.add_node("s_star_0i", r"$s^{\ast}_{0i}$", 2, 5)
pgm.add_node("gamma_star_i", r"$\gamma^{\ast}_i$", 3, 5)
pgm.add_node("sigma_ui", r"$\sigma_{ui}$", 4, 5)
pgm.add_node("sigma_si", r"$\sigma_{si}$", 5, 5)

# latent variables for cell-specific outcomes
pgm.add_node(
    "u_star_ij",
    r"$u^{\ast}_{ij}$",
    2,
    4,
    scale=1.0,
    shape="rectangle",
)
pgm.add_node(
    "s_star_ij",
    r"$s^{\ast}_{ij}$",
    4,
    4,
    scale=1.0,
    shape="rectangle",
)

# observed data
pgm.add_node(
    "t_star_j",
    r"$t^{\ast}_j$",
    6.0,
    3.25,
    observed=True,
    shape="rectangle",
)
pgm.add_node(
    "u_obs_ij",
    r"$\hat{u}^{\ast}_{ij}$",
    2,
    2.5,
    scale=1.0,
    observed=True,
)
pgm.add_node(
    "s_obs_ij",
    r"$\hat{s}^{\ast}_{ij}$",
    4,
    2.5,
    scale=1.0,
    observed=True,
)

# edges
edge_params = {"head_length": 0.3, "head_width": 0.25, "lw": 0.7}
pgm.add_edge("mu_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("mu_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("mu_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("sigma_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_si", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_si", plot_params=edge_params)

pgm.add_edge("u_star_0i", "u_star_ij", plot_params=edge_params)
pgm.add_edge("s_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("u_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("gamma_star_i", "s_star_ij", plot_params=edge_params)

pgm.add_edge("u_star_ij", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("s_star_ij", "s_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_ui", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_si", "s_obs_ij", plot_params=edge_params)

pgm.add_edge("t_star_j", "u_star_ij", plot_params=edge_params)
pgm.add_edge("t_star_j", "s_star_ij", plot_params=edge_params)

# plates
pgm.add_plate([0.5, 1.2, 5, 4.4], label=r"$i \in \{1, \ldots, G\}$", shift=-0.1)
pgm.add_plate(
    [1.0, 1.8, 5.5, 2.75], label=r"$j \in \{1, \ldots, N\}$", shift=-0.1
)

pgm.render()
```

Of course, in general, we will not have access to the analytical solutions, so
we will need to simulate the system and then perform inference on the latent
variables as suggested in (\ref{eq-init-conds-priors-num}--\ref{eq-s-obs-num}).

\begin{align}
    u^{\ast}_0, s^{\ast}_0 &\sim \text{LogNormal}(\mu_{0}, \sigma_{0}^2) \label{eq-init-conds-priors-num}, \\
    \gamma^{\ast} &\sim \text{LogNormal}(\mu_{\gamma}, \sigma_{\gamma}^2) \label{eq-gamma-prior-num}, \\
    \sigma_u, \sigma_s &\sim \text{HalfNormal}(\mu_{\sigma}, \sigma_{\sigma}^2) \label{eq-noise-std-priors-num}, \\
    (u^{\ast}, s^{\ast}) &= \text{NumericalSolver}\left(\frac{du^{\ast}}{dt^{\ast}}, \frac{ds^{\ast}}{dt^{\ast}}, u^{\ast}_0, s^{\ast}_0, \gamma^{\ast}, t^{\ast}\right) \label{eq-numerical-solution}, \\
    \hat{u}^{\ast} &\sim \text{Normal}(u^{\ast}, \sigma_u^2) \label{eq-u-obs-num}, \\
    \hat{s}^{\ast} &\sim \text{Normal}(s^{\ast}, \sigma_s^2) \label{eq-s-obs-num}.
\end{align}

Where $\frac{du^{\ast}}{dt^{\ast}}$ and $\frac{ds^{\ast}}{dt^{\ast}}$, 
are given by (\ref{eq-dustardtstar}-\ref{eq-dsstardtstar}) and noting that 
we suppress the indices from 
(\ref{eq-init-conds-priors-gene}--\ref{eq-s-obs-cell}) for brevity.