### 2024-04-25 nbs/concepts/rrv/index.qmd
chore(nbs): init rrv Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: \'Renormalizing\' effective theories for gene expression velocities over differentiation states
title-slide-attributes: 
  # credit: https://github.com/ddatsko/3d-ising-model/
  data-background-image: https://user-images.githubusercontent.com/35429810/116011574-72395b00-a62e-11eb-847b-26cf5034d64c.gif
  data-background-position: center
  data-background-size: contain
  data-background-opacity: "0.4"
toc: true
number-sections: true
highlight-style: pygments
csl: ../../bibstyle.csl
lightbox: auto
bibliography: ../../references.bib
format: 
  revealjs:
    slide-number: true
    width: 1600
    height: 900
    logo: "../../_static/logo.png"
    theme: ["theme/custom.scss", night]
editor: source
filters:
  - include-code-files
---

# Simulation ⇄ Inference {
    data-background-image="https://upload.wikimedia.org/wikipedia/commons/b/b2/Ising_Criticality2.gif"
    data-background-size="contain"
    data-background-opacity="0.3"
  }

## Simulation

- Simulation is the process of generating data from a model [@Wells2012-qf;@Alon2019-ah].
- Inside the model, we may generate (intermediate) data from other models.
- We can generate samples of our observables by executing all nested models.

## Inference

- Inference is the process of estimating the parameters of a model given data.
- This is done by computing the posterior distribution of the parameters given the data.
- Inside the model, we may estimate the parameters of other models.

# References