### 2024-04-26 nbs/concepts/rrv/index.qmd
feat(nbs): add inference plots to rrv Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: \'Renormalizing\' effective theories for gene expression velocities over differentiation states
title-slide-attributes: 
  # credit: https://github.com/ddatsko/3d-ising-model/
  data-background-image: https://user-images.githubusercontent.com/35429810/116011574-72395b00-a62e-11eb-847b-26cf5034d64c.gif
  data-background-position: center
  data-background-size: contain
  data-background-opacity: "0.4"
toc: true
toc-depth: 1
number-sections: true
highlight-style: pygments
csl: ../../bibstyle.csl
lightbox: auto
bibliography: ../../references.bib
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
format: 
  revealjs:
    slide-number: true
    width: 1600
    height: 900
    logo: "../../_static/logo.png"
    theme: ["theme/custom.scss", night]
editor: source
filters:
  - include-code-files
---

# Splice census from bulk to single-cell RNA-seq

## Simplistic cartoon of gene expression

![](https://www.nature.com/scitable/content/ne0000/ne0000/ne0000/ne0000/14711098/U2CP3-1_SynthesisDegredati_ksm.jpg){.r-stretch fig-align="center"}

[@Alberts2014-kt]

## Splicing mechanism

![](https://media.springernature.com/full/springer-static/image/art%3A10.1038%2Fs41580-022-00545-z/MediaObjects/41580_2022_545_Fig1_HTML.png){fig-align="center" height="760"}

[@Marasco2023-yb]

## (Pre‐)mRNA response profile classification

![](https://www.embopress.org/cms/10.1038/msb.2011.62/asset/329a576a-84a0-47c0-9bfc-125e6ab740b6/assets/graphic/msb201162-fig-0001-m.jpg){fig-align="center"}

[@Zeisel2011-nz]

## (Pre‐)mRNA stimulus‐induced response profiles

![](https://www.embopress.org/cms/10.1038/msb.2011.62/asset/042557d4-dbc2-472a-a685-94666dcb90a1/assets/graphic/msb201162-fig-0003-m.jpg){fig-align="center" height="720"}

[@Zeisel2011-nz]

## Single-cell RNA velocity

![](https://www.embopress.org/cms/10.15252/msb.202110282/asset/36cba1b7-8d35-491f-b2bd-a543103be834/assets/graphic/msb202110282-fig-0001-m.png){fig-align="center" height="720"}

[@La_Manno2018-lj;@Bergen2021-qz]

# Deterministic modeling of gene expression

## Variables {.center}

| Symbol   | Description                 | Units                         | $O(-)$ Estimate  | Note                                                                          |
| -------- | --------------------------- | ----------------------------- | ---------------- | ----------------------------------------------------------------------------- |
| $u$      | Number of pre-mRNA          | molecules/cell                | $10^0 - 10^4$    | Wide range accounts for low to high gene expression levels.                   |
| $s$      | Number of mRNA              | molecules/cell                | $10^0 - 10^5$    | Similar to $u$; depends on gene expression and stability of mRNA.             |
| $t$      | Time                        | seconds ($s$) to hours ($hr$) | $5s$ - $48hr$    | Depends on the duration of the experimental observation.                      |

## Parameters {.center}

| Symbol   | Description                 | Units                         | $O(-)$ Estimate  | Note                                                                          |
| -------- | --------------------------- | ----------------------------- | ---------------- | ----------------------------------------------------------------------------- |
| $\alpha$ | Production rate of pre-mRNA | molecules/(cell·hr)           | $10^0 - 10^3$    | Many transcripts are produced at rates in the range or 1 to 1000 per hour.    |
| $\beta$  | Splicing rate of pre-mRNA   | $hr^{-1}$                     | $10^{-1} - 10^2$ | Many Pre-mRNA to mRNA splicing rates are in the range of 1 minute to 6 hours. |
| $\gamma$ | Degradation rate of mRNA    | $hr^{-1}$                     | $10^{-2} - 10^0$ | Many mRNA half-lives are in the range of a half-hour to a day.                |

## Dynamics {.center}

Given state variables representing concentrations of pre-mRNA, $u$, 
and mRNA, $s$, we have the following ordinary differential equations 
taken from [@Zeisel2011-nz],

\begin{align}
\frac{du}{dt} & = \alpha - \beta u  \label{eq-dudt}, \\
\frac{ds}{dt} & = \beta u - \gamma s  \label{eq-dsdt},
\end{align}

These are 
usually presented with the concept that the parameter $\alpha$ could better 
account for the external inputs to the regulation of transcription of the gene 
if it were allowed to be a function of time, $\alpha(t)$.

## Dimensional analysis

| Dimensioned Parameter        | Relation to Rescaling Parameters | Fundamental Units     |
| ---------------------------- | -------------------------------- | --------------------- |
| Production Rate ($\alpha$)   | $\alpha = U_0 \beta$             | molecules/(cell·time) |
| Splicing Rate ($\beta$)      | Reference Scale for $t^*$        | $1/\text{time}$       |
| Degradation Rate ($\gamma$)  | $\gamma = \gamma^* \beta$        | $1/\text{time}$       |
| Pre-mRNA Concentration ($u$) | $u = u^* U_0$                    | molecules/cell        |
| mRNA Concentration ($s$)     | $s = s^* U_0$                    | molecules/cell        |
| Time ($t$)                   | $t = t^* / \beta$                | time                  |

## Dimensionless variables and parameters

| Dimensionless Variables and Parameters | Definition           | Description                                                                                                                                                                    |
| -------------------------------------- | -------------------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ |
| $u^{\ast}$ , $s^{\ast}$                | $u / U_0$, $s / U_0$ | Characteristic scale of (pre-)mRNA concentration based on the balance between production and splicing rates.                                                                   |
| $t^*$                                  | $\beta t$            | Characteristic time scale relative to the splicing rate.                                                                                                                       |
| $\gamma^*$                             | $\gamma / \beta$     | Relative degradation rate, comparing the degradation rate of mRNA to the splicing rate, indicating the stability or turnover rate of mRNA relative to its production/splicing. |

## Interpreting dimensionless parameters

| $\beta$ (hr$^{-1}$) | 1 unit of $t^*$ (hr) | 10 units of $t^*$ (hr) | 100 units of $t^*$ (hr) | 1 unit of $t^{\ast}$ | 10 units of $t^{\ast}$ | 100 units of $t^{\ast}$ |
| ------------------- | -------------------- | ---------------------- | ----------------------- | -------------------- | ---------------------- | ----------------------- |
| $10^{-1}$           | 10                   | 100                    | 1000                    | 1                    | 10                     | 100                     |
| $10^0$              | 1                    | 10                     | 100                     | 1                    | 10                     | 100                     |
| $10^1$              | 0.1                  | 1                      | 10                      | 1                    | 10                     | 100                     |
| $10^2$              | 0.01                 | 0.1                    | 1                       | 1                    | 10                     | 100                     |

## Dimensionless dynamics

The system of equations redefined in terms of $U_0 = \frac{\alpha}{\beta}$ and $\beta$ as the reference scale for time,

| Dimensionless | Definition           |
| -------------------------------------- | -------------------- |
| $u^{\ast}$ , $s^{\ast}$                | $u / U_0$, $s / U_0$ |
| $t^*$                                  | $\beta t$            |
| $\gamma^*$                             | $\gamma / \beta$     |

is given by

\begin{align}
\frac{du^{\ast}}{dt^{\ast}} & = 1 - u^{\ast}  \label{eq-dustardtstar}, \\
\frac{ds^{\ast}}{dt^{\ast}} & = u^{\ast} - \gamma^{\ast} s^{\ast},  
  \label{eq-dsstardtstar}
\end{align}

containing the 
$6 - 2 = 4$ variables and parameters: $u^{\ast}, s^{\ast}, t^{\ast}, \gamma^{\ast}$.

[@Zeisel2011-nz;@Cantwell2002-dv]

## Analytical solution

The complete solutions for the dimensionless system for $\gamma^{\ast} \neq 1$ and 
$\xi = \frac{u^{\ast}_0 - 1}{\gamma^{\ast} - 1}$
are then given by

\begin{align}
u^{\ast}(t^{\ast}) & = \underbrace{1}_{\text{Steady State}} + 
  \underbrace{(u^{\ast}_0 - 1) e^{-t^{\ast}}}_{\text{Transient Component}}, 
  \label{eq-ustar-dimless} \\
s^{\ast}(t^{\ast}) & = \underbrace{\frac{1}{\gamma^{\ast}}}_{\text{Steady State}} + 
  \underbrace{\left( s^{\ast}_0 - \xi - \frac{1}{\gamma^{\ast}} \right) 
  e^{-\gamma^{\ast} t^{\ast}} + \xi e^{-t^{\ast}}}_{\text{Transient Components}}.  
  \label{eq-sstar-dimless}
\end{align}

For the case $\gamma^{\ast} = 1$, we find $u^{\ast}$ is unchanged
and $s^{\ast}$ is then given by

$$
s^{\ast}(t^{\ast}) = 1 + (s^{\ast}_0 - 1) e^{-t^{\ast}} + 
  (u^{\ast}_0 - 1) t^{\ast} e^{-t^{\ast}}.
$$

Note we can restore an arbitrary initial timepoint putting 
$t^{\ast} \rightarrow \Delta t = t^{\ast} - t^{\ast}_0$.

## Simulation

```{python}
# | label: transcription-splicing-degradation-simulation
# | code-fold: true
import diffrax
from pyrovelocity.models import solve_transcription_splicing_model
from pyrovelocity.models import solve_transcription_splicing_model_analytical
from pyrovelocity.logging import configure_logging
from jax import numpy as jnp

logger = configure_logging("nbs")

ts0 = jnp.linspace(0.1, 4.0, 40)
ts1 = jnp.linspace(4.0 + (10.0 - 4.0) / 20, 10.0, 20)
ts = jnp.concatenate([ts0, ts1])
initial_state = jnp.array([0.1, 0.1])
params = (1.00,)
colormap_name = "RdBu"

solution = solve_transcription_splicing_model(
    ts,
    initial_state,
    params,
)
analytical_solution = solve_transcription_splicing_model_analytical(
    ts,
    initial_state,
    params,
)
analytical_simulation_error = diffrax.Solution(
    t0=ts[0],
    t1=ts[-1],
    ts=ts,
    ys=analytical_solution.ys - solution.ys,
    interpolation=None,
    stats={},
    result=diffrax.RESULTS.successful,
    solver_state=None,
    controller_state=None,
    made_jump=None,
)

logger.info(
    f"\nTrajectory preview times:\n{solution.ts[:3]}\n"
    f"\nSimulated trajectory values:\n{solution.ys[:3]}\n"
    f"\nAnalytical trajectory values:\n{analytical_solution.ys[:3]}\n\n"
    f"\nTrajectory preview times:\n{solution.ts[-3:]}\n"
    f"\nSimulated trajectory values:\n{solution.ys[-3:]}\n"
    f"\nAnalytical trajectory values:\n{analytical_solution.ys[-3:]}\n\n"
)
```

## Visualization {.center}

```{python}
# | label: fig-transcription-splicing-degradation-plot-simulation-trajectory-1
# | code-fold: true
# | fig-cap: Simulated trajectories of pre-mRNA and mRNA concentrations over time for the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_trajectories

plot_deterministic_simulation_trajectories(
    solution=solution,
    title_prefix="TSD Model Simulated",
    colormap_name=colormap_name,
)
```

## Error analysis {.center}

```{python}
# | label: fig-transcription-splicing-degradation-plot-analytical-trajectory-1
# | code-fold: true
# | fig-cap: Absolute error in analytical vs simulated trajectories of pre-mRNA and mRNA concentrations over time for the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_trajectories

plot_deterministic_simulation_trajectories(
    solution=analytical_simulation_error,
    title_prefix="Analytical-Simulation Error",
    colormap_name=colormap_name,
)
```

## Phase space {.center}

```{python}
# | label: fig-transcription-splicing-degradation-plot-simulation-phase-portrait-1
# | code-fold: true
# | fig-cap: Phase portrait of the transcription-splicing-degradation model.
from pyrovelocity.plots import plot_deterministic_simulation_phase_portrait

plot_deterministic_simulation_phase_portrait(
    solution=solution,
    title_prefix="TSD Model",
    colormap_name=colormap_name,
)
```

# Probabilistic inference in dynamical systems

## Data description

If we are given samples of pre-mRNA and mRNA counts along with
the sampling time of each cell, we can describe the data set 
$\mathcal{D}$ consisting of pre-mRNA and mRNA counts for a number 
of genes $G$ across a number of cells $N$

\begin{align}
\mathcal{D} = \left\{ (t_j, u_{ij}, s_{ij}) \mid 
  i \in \{1, \ldots, G\}, j \in \{1, \ldots, N\} \right\} \label{eq-dataset},
\end{align}

where

- $t_j$ represents the sampling time for cell $j$,
- $u_{ij}$ represents the count of pre-mRNA for gene $i$ in cell $j$,
- $s_{ij}$ represents the count of mRNA for the same gene $i$ in the same cell $j$,
- $G$ is the total number of genes in the study,
- $N$ is the total number of cells sampled.

If we were able to observe each cell $j$ at multiple timepoints, we could add
another index $k \in {1, ..., K_j}$.

## Inference problem

Roughly speaking, the inference problem involves

1. **Normalization**: Transforming the observed counts $u_{ij}$ and $s_{ij}$ to 
   their dimensionless counterparts $u^{\ast}_{ij}$ and $s^{\ast}_{ij}$ using 
   a reasonable concentration scale $U_0$.
2. **Model Fitting**: Applying statistical methods to estimate the latent 
   variables 
   $\Theta = \left( \gamma^{\ast}, u^{\ast}_0, s^{\ast}_0 \right)$ 
   that best fit the observed data.
3. **Evaluation**: Assessing the fit of the model and the estimated parameters'
   biological plausibility and consistency across different genes and cells.

We will return later to the alternative version of this problem where $t_j$ 
are not observed and must be inferred along with the latent variables 
$\Theta = \left( \gamma^{\ast}, u^{\ast}_0, s^{\ast}_0, t^{\ast} \right)$.

## Model formulation

given a number of 
genes $G$ and cells $N$ that are measured at $K_j$ distinct times, then for each
$(i,j,k) \in \{1, \ldots, G\} \otimes \{1, \ldots, N\} \otimes \{1, \ldots, K_j\}$ 
we have

\begin{align}
    u^{\ast}_{0i}, s^{\ast}_{0i} &\sim 
      \text{LogNormal}(\mu_{0}, \sigma_{0}^2), 
      \label{eq-init-conds-priors-gene} \\

    \gamma^{\ast}_i &\sim 
      \text{LogNormal}(\mu_{\gamma}, \sigma_{\gamma}^2),
      \label{eq-gamma-prior-gene} \\

    \sigma_{ui}, \sigma_{si} &\sim 
      \text{HalfNormal}(\mu_{\sigma}, \sigma_{\sigma}^2),
      \label{eq-noise-std-priors-gene} \\

    {u^{\ast}}^{k}_{ij} &= 
      1 + (u^{\ast}_{0i} - 1) \cdot e^{-{t^{\ast}}^k_j},
      \label{eq-u-star-model-cell} \\

    {s^{\ast}}^{k}_{ij} &= 
      \begin{cases}
        \frac{1}{\gamma^{\ast}_i} + 
        \left( s^{\ast}_{0i} - \xi_i - \frac{1}{\gamma^{\ast}_i} \right) \cdot e^{-\gamma^{\ast}_i {t^{\ast}}^k_j} + 
        \xi_i \cdot e^{-{t^{\ast}}^k_j},& \gamma^{\ast}_i \neq 1 \\

        1 + (s^{\ast}_{0i} - 1) e^{-{t^{\ast}}^k_j} + 
          (u^{\ast}_{0i} - 1) {t^{\ast}}^k_j e^{-{t^{\ast}}^k_j},& \gamma^{\ast}_i = 1 \\
      \end{cases}, \label{eq-s-star-model-cell} \\

    \hat{u}^{\ast}{}^k_{ij} &\sim \text{Normal}({u^{\ast}}^k_{ij}, \sigma_{ui}^2), 
    \label{eq-u-obs-cell} \\

    \hat{s}^{\ast}{}^k_{ij} &\sim \text{Normal}({s^{\ast}}^k_{ij}, \sigma_{si}^2). 
    \label{eq-s-obs-cell}
\end{align}

## Graphical model

```{python}
# | label: fig-transcription-splicing-degradation-graphical-model-1
# | code-fold: true
# | fig-cap: Graphical representation of the transcription-splicing-degradation inference model.
import daft
import matplotlib.pyplot as plt

plt.rcParams["font.family"] = "serif"
plt.rcParams["font.size"] = 16
plt.rcParams["text.usetex"] = True

pgm = daft.PGM(line_width=1.2, dpi=85)

# hyperparameters
pgm.add_node("mu_init", r"$\mu_{0}$", 0.5, 6, fixed=True)
pgm.add_node("sigma_init", r"$\sigma_{0}^2$", 1.5, 6, fixed=True)
pgm.add_node("mu_gamma", r"$\mu_{\gamma}$", 2.5, 6, fixed=True)
pgm.add_node("sigma_gamma", r"$\sigma_{\gamma}^2$", 3.5, 6, fixed=True)
pgm.add_node("mu_sigma", r"$\mu_{\sigma}$", 4.5, 6, fixed=True)
pgm.add_node("sigma_sigma", r"$\sigma_{\sigma}^2$", 5.5, 6, fixed=True)

# latent variables for gene-specific parameters
pgm.add_node("u_star_0i", r"$u^{\ast}_{0i}$", 1, 5)
pgm.add_node("s_star_0i", r"$s^{\ast}_{0i}$", 2, 5)
pgm.add_node("gamma_star_i", r"$\gamma^{\ast}_i$", 3, 5)
pgm.add_node("sigma_ui", r"$\sigma_{ui}$", 4, 5)
pgm.add_node("sigma_si", r"$\sigma_{si}$", 5, 5)

# latent variables for cell-specific outcomes
pgm.add_node(
    "u_star_ij",
    r"${u^{\ast}}^k_{ij}$",
    2,
    3.8,
    scale=1.0,
    shape="rectangle",
)
pgm.add_node(
    "s_star_ij",
    r"${s^{\ast}}^k_{ij}$",
    4,
    3.8,
    scale=1.0,
    shape="rectangle",
)

# observed data
pgm.add_node(
    "t_star_j",
    r"${t^{\ast}}^k_j$",
    5.9,
    3.1,
    observed=True,
    shape="rectangle",
)
pgm.add_node(
    "u_obs_ij",
    r"$\hat{u}^{\ast}{}^{k}_{ij}$",
    2,
    2.4,
    scale=1.0,
    observed=True,
)
pgm.add_node(
    "s_obs_ij",
    r"$\hat{s}^{\ast}{}^{k}_{ij}$",
    4,
    2.4,
    scale=1.0,
    observed=True,
)

# edges
edge_params = {"head_length": 0.3, "head_width": 0.25, "lw": 0.7}
pgm.add_edge("mu_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("mu_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("mu_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("sigma_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_si", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_si", plot_params=edge_params)

pgm.add_edge("u_star_0i", "u_star_ij", plot_params=edge_params)
pgm.add_edge("s_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("u_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("gamma_star_i", "s_star_ij", plot_params=edge_params)

pgm.add_edge("u_star_ij", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("s_star_ij", "s_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_ui", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_si", "s_obs_ij", plot_params=edge_params)

pgm.add_edge("t_star_j", "u_star_ij", plot_params=edge_params)
pgm.add_edge("t_star_j", "s_star_ij", plot_params=edge_params)

# plates
pgm.add_plate(
    [0.4, 1.0, 5, 4.5],
    label=r"$i \in \{1, \ldots, G\}$",
    shift=-0.1,
    fontsize=12,
)
pgm.add_plate(
    [0.8, 1.4, 5.9, 3.2],
    label=r"$j \in \{1, \ldots, N\}$",
    shift=-0.1,
    fontsize=12,
)
pgm.add_plate(
    [1.2, 1.8, 5.2, 2.5],
    label=r"$k \in \{1, \ldots, K_j\}$",
    shift=-0.1,
    fontsize=12,
)

pgm.render()
```

## Relaxing the assumption of analytical tractability

In general, we will not have access to the analytical solutions, so
we will need to simulate the system and then perform inference on the latent
variables

\begin{align}
    u^{\ast}_0, s^{\ast}_0 &\sim \text{LogNormal}(\mu_{0}, \sigma_{0}^2) \label{eq-init-conds-priors-num}, \\
    \gamma^{\ast} &\sim \text{LogNormal}(\mu_{\gamma}, \sigma_{\gamma}^2) \label{eq-gamma-prior-num}, \\
    \sigma_u, \sigma_s &\sim \text{HalfNormal}(\mu_{\sigma}, \sigma_{\sigma}^2) \label{eq-noise-std-priors-num}, \\
    (u^{\ast}, s^{\ast}) &= \text{NumericalSolver}\left(\frac{du^{\ast}}{dt^{\ast}}, \frac{ds^{\ast}}{dt^{\ast}}, u^{\ast}_0, s^{\ast}_0, \gamma^{\ast}, t^{\ast}\right) \label{eq-numerical-solution}, \\
    \hat{u}^{\ast} &\sim \text{Normal}(u^{\ast}, \sigma_u^2) \label{eq-u-obs-num}, \\
    \hat{s}^{\ast} &\sim \text{Normal}(s^{\ast}, \sigma_s^2) \label{eq-s-obs-num}.
\end{align}

Where $\frac{du^{\ast}}{dt^{\ast}}$ and $\frac{ds^{\ast}}{dt^{\ast}}$ 
are given and
we suppress the indices from the previous model for brevity.

## Relaxing the assumption of access to time series data

If we only observe one time point for each cell, 
then the model reduces to the graphical model in 
@fig-transcription-splicing-degradation-graphical-model-2.

```{python}
# | label: fig-transcription-splicing-degradation-graphical-model-2
# | code-fold: true
# | fig-cap: Graphical model for the transcription-splicing-degradation model observed at a single time point.
import daft
import matplotlib.pyplot as plt

plt.rcParams["font.family"] = "serif"
plt.rcParams["font.size"] = 16
plt.rcParams["text.usetex"] = True

pgm = daft.PGM(line_width=1.2, dpi=80)

# hyperparameters
pgm.add_node("mu_init", r"$\mu_{0}$", 0.5, 6, fixed=True)
pgm.add_node("sigma_init", r"$\sigma_{0}^2$", 1.5, 6, fixed=True)
pgm.add_node("mu_gamma", r"$\mu_{\gamma}$", 2.5, 6, fixed=True)
pgm.add_node("sigma_gamma", r"$\sigma_{\gamma}^2$", 3.5, 6, fixed=True)
pgm.add_node("mu_sigma", r"$\mu_{\sigma}$", 4.5, 6, fixed=True)
pgm.add_node("sigma_sigma", r"$\sigma_{\sigma}^2$", 5.5, 6, fixed=True)

# latent variables for gene-specific parameters
pgm.add_node("u_star_0i", r"$u^{\ast}_{0i}$", 1, 5)
pgm.add_node("s_star_0i", r"$s^{\ast}_{0i}$", 2, 5)
pgm.add_node("gamma_star_i", r"$\gamma^{\ast}_i$", 3, 5)
pgm.add_node("sigma_ui", r"$\sigma_{ui}$", 4, 5)
pgm.add_node("sigma_si", r"$\sigma_{si}$", 5, 5)

# latent variables for cell-specific outcomes
pgm.add_node(
    "u_star_ij",
    r"$u^{\ast}_{ij}$",
    2,
    4,
    scale=1.0,
    shape="rectangle",
)
pgm.add_node(
    "s_star_ij",
    r"$s^{\ast}_{ij}$",
    4,
    4,
    scale=1.0,
    shape="rectangle",
)

# observed data
pgm.add_node(
    "t_star_j",
    r"$t^{\ast}_j$",
    6.0,
    3.25,
    observed=True,
    shape="rectangle",
)
pgm.add_node(
    "u_obs_ij",
    r"$\hat{u}^{\ast}_{ij}$",
    2,
    2.5,
    scale=1.0,
    observed=True,
)
pgm.add_node(
    "s_obs_ij",
    r"$\hat{s}^{\ast}_{ij}$",
    4,
    2.5,
    scale=1.0,
    observed=True,
)

# edges
edge_params = {"head_length": 0.3, "head_width": 0.25, "lw": 0.7}
pgm.add_edge("mu_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "u_star_0i", plot_params=edge_params)
pgm.add_edge("mu_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("sigma_init", "s_star_0i", plot_params=edge_params)
pgm.add_edge("mu_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("sigma_gamma", "gamma_star_i", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_ui", plot_params=edge_params)
pgm.add_edge("mu_sigma", "sigma_si", plot_params=edge_params)
pgm.add_edge("sigma_sigma", "sigma_si", plot_params=edge_params)

pgm.add_edge("u_star_0i", "u_star_ij", plot_params=edge_params)
pgm.add_edge("s_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("u_star_0i", "s_star_ij", plot_params=edge_params)
pgm.add_edge("gamma_star_i", "s_star_ij", plot_params=edge_params)

pgm.add_edge("u_star_ij", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("s_star_ij", "s_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_ui", "u_obs_ij", plot_params=edge_params)
pgm.add_edge("sigma_si", "s_obs_ij", plot_params=edge_params)

pgm.add_edge("t_star_j", "u_star_ij", plot_params=edge_params)
pgm.add_edge("t_star_j", "s_star_ij", plot_params=edge_params)

# plates
pgm.add_plate(
    [0.5, 1.2, 5, 4.4],
    label=r"$i \in \{1, \ldots, G\}$",
    shift=-0.1,
    fontsize=12,
)
pgm.add_plate(
    [1.0, 1.8, 5.5, 2.75],
    label=r"$j \in \{1, \ldots, N\}$",
    shift=-0.1,
    fontsize=12,
)

pgm.render()
```

## HMC Inference

Now we can set up some sample data and perform inference on the latent variables

```{python}
# | label: fig-generate-test-data-for-deterministic-model-inference
# | code-fold: true
# | fig-cap: Test data
# | output: false

from pyrovelocity.models._deterministic_inference import (
    generate_posterior_inference_data,
    generate_prior_inference_data,
    generate_test_data_for_deterministic_model_inference,
    plot_sample_phase_portraits,
    plot_sample_trajectories,
    plot_sample_trajectories_with_percentiles,
    print_inference_data_structure,
)

(
    times,
    data,
    num_genes,
    num_cells,
    num_timepoints,
    num_modalities,
) = generate_test_data_for_deterministic_model_inference(
    num_genes=1,
    num_cells=20,
    num_timepoints=1,
    num_modalities=2,
    noise_levels=(0.001, 0.001),
)

num_chains = 1
num_samples = 10
num_warmup = 10

idata_prior = generate_prior_inference_data(
    times=times,
    data=data,
    num_chains=num_chains,
    num_samples=num_samples,
    num_genes=num_genes,
    num_cells=num_cells,
    num_timepoints=num_timepoints,
    num_modalities=num_modalities,
)

idata_posterior = generate_posterior_inference_data(
    times=times,
    data=data,
    num_chains=num_chains,
    num_samples=num_samples,
    num_genes=num_genes,
    num_cells=num_cells,
    num_timepoints=num_timepoints,
    num_modalities=num_modalities,
    num_warmup=num_warmup,
)
```

```{python}
# | label: show-inference-data-structure
idata_posterior
```

## Prior distributions {.center}

```{python}
# | label: fig-transcription-splicing-degradation-prior
# | code-fold: true
# | fig-cap:
# |   - Initial concentratiosn
# |   - Degradation vs splicing rate
# |   - Observational noise levels
# | layout-ncol: 3
import arviz as az
import shutil

light_gray = "#bcbcbc"
variables = ["initial_conditions", "gamma", "sigma"]
with plt.style.context("pyrovelocity.styles.common"):
    if not shutil.which("latex"):
        plt.rc("text", usetex=False)
    for var in variables:
        az.plot_posterior(
            idata_prior,
            var_names=[var],
            group="prior",
            kind="hist",
            color=light_gray,
            round_to=2,
        )
```

## Posterior distributions {.center}

```{python}
# | label: fig-transcription-splicing-degradation-posterior
# | code-fold: true
# | fig-cap:
# |   - Initial concentratiosn
# |   - Degradation vs splicing rate
# |   - Observational noise levels
# | layout-ncol: 3
with plt.style.context("pyrovelocity.styles.common"):
    if not shutil.which("latex"):
        plt.rc("text", usetex=False)
    for var in variables:
        az.plot_posterior(
            idata_posterior,
            var_names=[var],
            group="posterior",
            kind="hist",
            color=light_gray,
            round_to=2,
        )
```

## Posterior predictive trajectories

```{python}
# | label: fig-transcription-splicing-degradation-posterior-predictive-trajectories
# | code-fold: true
# | fig-cap: Posterior predictive trajectories for the transcription-splicing-degradation model.
# | warnings: false
with plt.style.context("pyrovelocity.styles.common"):
    if not shutil.which("latex"):
        plt.rc("text", usetex=False)
    figs = plot_sample_trajectories(
        idata=idata_posterior,
    )
    for fig in figs:
        fig.show()
```

## Posterior predictive phase portraits {.center}

```{python}
# | label: fig-transcription-splicing-degradation-posterior-predictive-phase-portraits
# | code-fold: true
# | fig-cap: Posterior predictive phase portraits for the transcription-splicing-degradation model.
with plt.style.context("pyrovelocity.styles.common"):
    if not shutil.which("latex"):
        plt.rc("text", usetex=False)
    figs = plot_sample_phase_portraits(
        idata=idata_posterior,
        colormap_name="RdBu",
    )
    for fig in figs:
        fig.show()
```

## Posterior predictive percentile bands

```{python}
# | label: fig-transcription-splicing-degradation-posterior-predictive
# | code-fold: true
# | fig-cap: Posterior predictive distribution for the transcription-splicing-degradation model.
with plt.style.context("pyrovelocity.styles.common"):
    if not shutil.which("latex"):
        plt.rc("text", usetex=False)
    figs = plot_sample_trajectories_with_percentiles(
        idata=idata_posterior,
    )
    for fig in figs:
        fig.show()
```

# Coarse-graining and renormalization {
    data-background-image="https://upload.wikimedia.org/wikipedia/commons/b/b2/Ising_Criticality2.gif"
    data-background-size="contain"
    data-background-opacity="0.3"
  }

## Mathematical example

{{< video https://www.youtube.com/embed/9vFbyHNz-8g?start=591 
    width="100%"
    height="85%"
>}}

[@Fisher1998-lm]

# References

<!-- :::: {.columns}

::: {.column .fragment}

- Content on the left

:::

::: {.column .fragment}
- Content on the right

:::

:::: -->

<!-- # Simulation ⇄ Inference {
    data-background-image="https://upload.wikimedia.org/wikipedia/commons/b/b2/Ising_Criticality2.gif"
    data-background-size="contain"
    data-background-opacity="0.3"
  }

## Simulation

- Simulation is the process of generating data from a model [@Wells2012-qf;@Alon2019-ah].
- Inside the model, we may generate (intermediate) data from other models.
- We can generate samples of our observables by executing all nested models.

## Inference

- Inference is the process of estimating the parameters of a model given data.
- This is done by computing the posterior distribution of the parameters given the data.
- Inside the model, we may estimate the parameters of other models. -->