### 2024-07-15 nbs/guides/installation.qmd
docs(guides): update installation
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Installation
---

::: {#nte-blank-cells .callout-note collapse=true title="Architecture and accelerator support"}
If you would like to test pyrovelocity without requiring a GPU you should be
able to install it on any platform. However, if you would like to apply it to
data sets larger than $O(10^3)$ cells, you will need a GPU.
In this case, we can only support amd64 Linux systems with a CUDA compatible GPU.
If you are having issues scaling pyrovelocity please do not hesitate to reach out to us.
:::

The latest [release of pyrovelocity](https://github.com/pinellolab/pyrovelocity/releases)
is [available on PyPI](https://pypi.org/project/pyrovelocity/) and 
[conda-forge](https://anaconda.org/conda-forge/pyrovelocity).
The instructions below are intended for users. To setup a development 
environment, please see [Contributing](../about/contributing.qmd).
Users who are having difficulty with system-level dependencies may also
benefit from consulting the [contributing documentation](../about/contributing.qmd).

::: {.panel-tabset}

## PyPI

We recommend creating a dedicated virtual environment to install and execute pyrovelocity.
One of the easiest ways to do this is with [uv](https://github.com/astral-sh/uv/tree/0.2.24?tab=readme-ov-file#getting-started).
After following the [Getting Started instructions from the uv repository](https://github.com/astral-sh/uv/tree/0.2.24?tab=readme-ov-file#getting-started)
to install the uv CLI for your platform, you can install a supported python
version, create a virtual environment, and install pyrovelocity from a blank directory with the following commands

```bash
echo "3.10" > .python-version
uv python install
uv venv -p 3.10
uv pip install pyrovelocity
```

Of course, if you have already prepared and activated a virtual environment
with a supported python version you can also install pyrovelocity directly
from [PyPI](https://pypi.org/project/pyrovelocity/) with pip.

```bash
pip install pyrovelocity
```

## Anaconda

If you would like to install pyrovelocity from [conda-forge](https://anaconda.org/conda-forge/pyrovelocity)
we strongly recommend using [pixi](https://prefix.dev/channels/conda-forge/packages/pyrovelocity)
or [micromamba](https://mamba.readthedocs.io/en/latest/installation/micromamba-installation.html).
Installation with the `conda` command will take an unreasonable amount of time and is not recommended.

::: {.panel-tabset}

### pixi

`pixi` will not automatically respect the user limit parameters you can
review via `ulimit -a`.
In particular `ulimit -n` should be set to 10000 or greater

```bash
ulimit -n 10000
```

After following the instructions to install [pixi](https://prefix.dev/channels/conda-forge/packages/pyrovelocity),
you can install pyrovelocity with the following commands

```bash
pixi init
pixi install pyrovelocity
```

You can then run `pixi shell` to enter a shell with the environment containing pyrovelocity.

### micromamba

After following the instructions to install [micromamba](https://mamba.readthedocs.io/en/latest/installation/micromamba-installation.html),
you can create a virtual environment, activate it, and install pyrovelocity with the following commands

```bash
alias mm=micromamba
mm create -n pyrovelocity python=3.10
mm activate pyrovelocity
mm install -c conda-forge pyrovelocity
```

:::

:::