### 2024-07-15 nbs/guides/installation.qmd
docs(guides): add note on dependency versions
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Installation
---

::: {#nte-architecture .callout-note collapse=true title="Architecture and accelerator support"}
If you would like to test pyrovelocity without requiring a GPU you should be
able to install it on any platform. However, if you would like to apply it to
data sets larger than $O(10^3)$ cells, you will need a GPU.
In this case, we can only support amd64 Linux systems with a CUDA compatible GPU.
If you are having issues scaling pyrovelocity please do not hesitate to reach out to us.
:::

The latest [release of pyrovelocity](https://github.com/pinellolab/pyrovelocity/releases)
is [available on PyPI](https://pypi.org/project/pyrovelocity/) and 
[conda-forge](https://anaconda.org/conda-forge/pyrovelocity).
The instructions below are intended for users. To setup a development 
environment, please see [Contributing](../about/contributing.qmd).
Users who are having difficulty with system-level dependencies may also
benefit from consulting the [contributing documentation](../about/contributing.qmd).

::: {#nte-versions .callout-important collapse=true title="Installation or usage errors and version constraints"}
If you have any issue with errors being raised during installation or usage
please check carefully that the version your package manager has installed is equivalent
to or newer than the one listed in the
[poetry.lock](https://github.com/pinellolab/pyrovelocity/blob/main/poetry.lock)
file in the repository.
We aim to support all latest stable versions of third-party dependencies
by lifting lower bounds on version constraints in the 
[pyproject.toml](https://github.com/pinellolab/pyrovelocity/blob/main/pyproject.toml) file.
This means that we generally do not support any older versions of dependencies
but consider any version more recent than the lower bound to be supported or a bug 
in pyrovelocity if attempting to use it produces an error.
Sometimes we are forced to delay support for a new version of a dependency
if it has been released released to PyPI but not yet to [conda-forge](https://conda-forge.org/packages/).
:::

::: {.panel-tabset}

## PyPI

We recommend creating a dedicated virtual environment to install and execute pyrovelocity.
One of the easiest ways to do this is with [uv](https://github.com/astral-sh/uv/tree/0.2.24?tab=readme-ov-file#getting-started).
After following the [Getting Started instructions from the uv repository](https://github.com/astral-sh/uv/tree/0.2.24?tab=readme-ov-file#getting-started)
to install the uv CLI for your platform, you can install a supported python
version, create a virtual environment, and install pyrovelocity from a blank directory with the following commands

```bash
echo "3.10" > .python-version
uv python install
uv venv -p 3.10
uv pip install pyrovelocity
```

Of course, if you have already prepared and activated a virtual environment
with a supported python version you can also install pyrovelocity directly
from [PyPI](https://pypi.org/project/pyrovelocity/) with pip.

```bash
pip install pyrovelocity
```

## Anaconda

If you would like to install pyrovelocity from [conda-forge](https://anaconda.org/conda-forge/pyrovelocity)
we strongly recommend using [pixi](https://prefix.dev/channels/conda-forge/packages/pyrovelocity)
or [micromamba](https://mamba.readthedocs.io/en/latest/installation/micromamba-installation.html).
Installation with the `conda` command will take an unreasonable amount of time and is not recommended.

::: {.panel-tabset}

### pixi

`pixi` will not automatically respect the user limit parameters you can
review via `ulimit -a`.
In particular `ulimit -n` should be set to 10000 or greater

```bash
ulimit -n 10000
```

After following the instructions to install [pixi](https://prefix.dev/channels/conda-forge/packages/pyrovelocity),
you can install pyrovelocity with the following commands

```bash
pixi init
pixi install pyrovelocity
```

You can then run `pixi shell` to enter a shell with the environment containing pyrovelocity.

### micromamba

After following the instructions to install [micromamba](https://mamba.readthedocs.io/en/latest/installation/micromamba-installation.html),
you can create a virtual environment, activate it, and install pyrovelocity with the following commands

```bash
alias mm=micromamba
mm create -n pyrovelocity python=3.10
mm activate pyrovelocity
mm install -c conda-forge pyrovelocity
```

:::

:::