### 2024-07-23 nbs/guides/interactive.qmd
docs(guides): add interactive results review guide
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Interactive results review
---

When testing or prototyping it is helpful to review the results of a workflow
execution interactively. This is relatively straightforward when working with locally
executed workflows as demonstrated in
[Usage](/templates/user_example/user_example.qmd).
In this guide we illustrate working with the results 
of workflows that have been executed remotely. In this particular case, GitHub 
OAuth access is required to the endpoint associated to the workflow execution.
Given this access, setup the remote client,

```{python}
from flytekit.remote.remote import FlyteRemote
from flytekit.configuration import Config

remote = FlyteRemote(
    Config.for_endpoint("flyte.cluster.pyrovelocity.net"),
)
```

login to the UI and copy the URI of the workflow or task whose outputs you want to review, and provide this to the `get` method of the remote client,

```{python}
workflow_outputs = remote.get(
    "flyte://v1/pyrovelocity/development/pyrovelocity-argo-nix-bui-5c9ebf8-dev-1hk-d5805de62e8044f79d4/f26c1pjy-0-dn0-0-dn7/o"
)
```

This will not download the results but will provide a summary of the outputs.

```{python}
from omegaconf import OmegaConf
from flytekit.interaction.string_literals import literal_map_string_repr

outputs_dict = literal_map_string_repr(workflow_outputs.literals)
outputs_dictconfig = OmegaConf.create(outputs_dict)
print(OmegaConf.to_yaml(outputs_dictconfig))
```

To download the desired results use

```{python}
from pyrovelocity.io.gcs import download_blob_from_uri

download_blob_from_uri(outputs_dictconfig.o0.dataframe.path)
```