### 2024-07-28 nbs/guides/interactive/interactive.qmd
fix(nbs): flesh out interactive review
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Interactive results review
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
toc: true
number-sections: true
highlight-style: gruvbox
csl: ../../bibstyle.csl
lightbox: auto
format:
  html:
    html-math-method: katex
  ipynb: default
---

```{python}
#| label: enable-autoreload
#| code-fold: true
#| output: false
%load_ext autoreload
%autoreload 2 
from IPython.display import Image, display
```

## Get results

::: {#wrn-simulated-data .callout-warning collapse=true title="Simulated Data"}
In this notebook we illustrate how to download and review results using
unrealistic simulated data. Therefore, nothing in this notebook should be taken
to represent the results of applying the model to real data.
:::

### Setup remote connection

When testing or prototyping it is helpful to review the results of a workflow
execution interactively. This is relatively straightforward when working with locally
executed workflows as demonstrated in
[Usage](/templates/user_example/user_example.qmd).
However it is increasingly the case that the data set sizes and model training
resources will not be available or accessible locally, we would like to be able to interactively review results even when they have been computed in a remote cluster with sufficient resources.

In this guide we therefore illustrate working with the results 
of workflows that have been executed remotely. In this particular case, GitHub 
OAuth access is required to the endpoint associated to the workflow execution.
Given this access, setup the remote client,

```{python}
#| label: instantiate-remote-client
from flytekit.remote.remote import FlyteRemote
from flytekit.configuration import Config

remote = FlyteRemote(
    Config.for_endpoint("flyte.cluster.pyrovelocity.net"),
)
```

### Identify results of interest

Login to the UI and copy the URI of the workflow or task whose inputs and outputs you want to review, and provide these to the `get` method of the remote client,

```{python}
# | label: get-workflow-io
workflow_inputs = remote.get(
    "flyte://v1/pyrovelocity/development/pyrovelocity-argo-nix-bui-5c9ebf8-dev-1hk-d5805de62e8044f79d4/f26c1pjy-0-dn0/i"
)
postprocessing_outputs = remote.get(
    "flyte://v1/pyrovelocity/development/pyrovelocity-argo-nix-bui-5c9ebf8-dev-1hk-d5805de62e8044f79d4/f26c1pjy-0-dn0-0-dn6/o"
)
```

This will not download the results but will provide a summary of the outputs.

```{python}
# | label: create-outputs-dict
from omegaconf import OmegaConf
from flytekit.interaction.string_literals import literal_map_string_repr
from pyrovelocity.utils import print_config_tree

inputs_dict = literal_map_string_repr(workflow_inputs.literals)
inputs_dictconfig = OmegaConf.create(inputs_dict)
print_config_tree(inputs_dict)

outputs_dict = literal_map_string_repr(postprocessing_outputs.literals)
outputs_dictconfig = OmegaConf.create(outputs_dict)
print_config_tree(outputs_dict)
```

### Download results

Download the results you would like to review,

```{python}
#| label: download-outputs
from pyrovelocity.io.gcs import download_blob_from_uri

pyrovelocity_data = download_blob_from_uri(
    outputs_dictconfig.o0.pyrovelocity_data.path
)
postprocessed_data = download_blob_from_uri(
    outputs_dictconfig.o0.postprocessed_data.path
)
```

## Analyze results

### Load data

In this example, we load the postprocessed data from the downloaded file,

```{python}
# | label: load-postprocessed-data
import scanpy as sc
from pyrovelocity.utils import print_anndata

adata = sc.read(postprocessed_data)
print_anndata(adata)
```

as well as the posterior samples,

```{python}
# | label: load-posterior-samples
from pyrovelocity.utils import pretty_print_dict
from pyrovelocity.io import CompressedPickle

posterior_samples = CompressedPickle.load(pyrovelocity_data)
pretty_print_dict(posterior_samples)
```

### Extract results of interest

In this case, we extract the gene selection data by cooptimizing MAE and correlation between spliced expression levels and estimates of temporal ordering

```{python}
# | label: extract-gene-selection
from pyrovelocity.analysis.analyze import pareto_frontier_genes

volcano_data = posterior_samples["gene_ranking"]
number_of_marker_genes = min(
    max(int(len(volcano_data) * 0.1), 4), 6, len(volcano_data)
)
putative_marker_genes = pareto_frontier_genes(
    volcano_data, number_of_marker_genes
)
```

### Generate plots

Here we re-generate the gene selection summary plot.

```{python}
# | label: generate-gene-selection-summary-plot
# | output: false
from pyrovelocity.plots import plot_gene_selection_summary

vector_field_basis = inputs_dictconfig.preprocess_data_args.vector_field_basis
cell_state = inputs_dictconfig.preprocess_data_args.cell_state

plot_gene_selection_summary(
    adata=adata,
    posterior_samples=posterior_samples,
    basis=vector_field_basis,
    cell_state=cell_state,
    plot_name="gene_selection_summary_plot.pdf",
    selected_genes=putative_marker_genes,
    show_marginal_histograms=False,
)
```

```{python}
# | label: show-gene-selection-summary-plot
# | code-fold: true
# | output: true
display(Image(filename=f"gene_selection_summary_plot.pdf.png"))
```

Of course, you can generate any plot of interest relevant to the loaded data.
See the [plots package source](https://github.com/pinellolab/pyrovelocity/tree/main/src/pyrovelocity/plots), [plots API reference](../../reference/plots.qmd), and the guide for the [workflow summarization task](https://github.com/pinellolab/pyrovelocity/blob/main/src/pyrovelocity/tasks/summarize.py) for ideas regarding built-in or custom plots.