### 2024-08-09 nbs/guides/preprocessing/preprocessing.qmd
refactor(nbs): migrate preprocessing to subdirectory
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Preprocessing
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
toc: true
number-sections: true
highlight-style: gruvbox
csl: ../../bibstyle.csl
lightbox: auto
format:
  html:
    html-math-method: katex
  ipynb: default
---

```{python}
#| label: enable-autoreload
#| code-fold: true
#| output: false
%load_ext autoreload
%autoreload 2 
from IPython.display import Image, display
```