### 2024-08-09 nbs/guides/preprocessing/preprocessing.qmd
chore(nbs): add gitignore
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Preprocessing
execute: 
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
toc: true
number-sections: true
highlight-style: gruvbox
csl: ../../bibstyle.csl
lightbox: auto
format:
  html:
    html-math-method: katex
  ipynb: default
---

```{python}
#| label: enable-autoreload
#| code-fold: true
#| output: false
%load_ext autoreload
%autoreload 2 
from IPython.display import Image, display
```

# Download data

```{python}
# | label: download-data
import polars as pl
import altair as alt

alt.data_transformers.enable("vegafusion")

flights = pl.read_parquet(
    "https://vegafusion-datasets.s3.amazonaws.com/vega/flights_1m.parquet"
)
```

# Make plots

```{python}
# | label: make-plots
delay_hist = (
    alt.Chart(flights)
    .mark_bar()
    .encode(alt.X("delay", bin=alt.Bin(maxbins=30)), alt.Y("count()"))
)
delay_hist
```