### 2024-08-20 reproducibility/manuscript/manuscript.qmd
docs(manuscript): specify per-format affiliations
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: |
  Pyro&thinsp;-Velocity: Probabilistic RNA Velocity inference from single-cell data
toc: true
toc-title: Contents
number-sections: true
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-mathphys-num
    classoption: 
      - lineno
    cite-method: natbib
    link-citations: true
    keep-tex: true
    equal-margins: false
    include-in-header: header.tex
    affiliations:
      - name: Massachusetts General Hospital Research Institute, Charlestown, MA
        id: 1
        department: Molecular Pathology Unit
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Harvard Medical School, Charlestown, MA
        id: 2
        department: Massachusetts General Hospital Cancer Center
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Massachusetts General Hospital, Boston, MA
        id: 3
        department: Center for Regenerative Medicine
        city: Boston
        state: MA
        postal-code: 02114
      - name: Harvard University, Cambridge, MA
        id: 4
        department: Harvard Stem Cell Institute
        city: Cambridge
        state: MA
        postal-code: 02139
      - name: Cambridge, MA
        id: 5
        department: Broad Institute of MIT and Harvard
        city: Cambridge
        state: MA
        postal-code: 02142
      - name: Laboratory of Neurodevelopmental Systems Biology, Lausanne, Switzerland
        id: 6
        department: |
          Brain and Mind Institute, School of Life Sciences, 
          École Polytechnique Fédérale de Lausanne (EPFL)
        city: Lausanne
        state: Switzerland
        postal-code: 1015
      - name: Cambridge, MA
        id: 7
        department: Basis Research Institute
        city: Cambridge
        state: MA
        postal-code: 02142
  html:
    code-fold: true
    html-math-method: mathjax
    affiliations:
      - name: Molecular Pathology Unit, Massachusetts General Hospital Research Institute, Charlestown, MA
        id: 1
      - name: Massachusetts General Hospital Cancer Center, Harvard Medical School, Charlestown, MA
        id: 2
      - name: Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA
        id: 3
      - name: Harvard Stem Cell Institute, Harvard University, Cambridge, MA
        id: 4
      - name: Broad Institute of MIT and Harvard, Cambridge, MA
        id: 5
      - name: Laboratory of Neurodevelopmental Systems Biology, Brain and Mind Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne (EPFL), Lausanne, Switzerland
        id: 6
      - name: Basis Research Institute, Cambridge, MA
        id: 7
  docx:
    reference-doc: styles.docx
    cite-method: citeproc
    filters:
      - authors-block
    affiliations:
      - name: Molecular Pathology Unit, Massachusetts General Hospital Research Institute, Charlestown, MA
        id: 1
      - name: Massachusetts General Hospital Cancer Center, Harvard Medical School, Charlestown, MA
        id: 2
      - name: Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA
        id: 3
      - name: Harvard Stem Cell Institute, Harvard University, Cambridge, MA
        id: 4
      - name: Broad Institute of MIT and Harvard, Cambridge, MA
        id: 5
      - name: Laboratory of Neurodevelopmental Systems Biology, Brain and Mind Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne (EPFL), Lausanne, Switzerland
        id: 6
      - name: Basis Research Institute, Cambridge, MA
        id: 7
csl: _extensions/sciexp/nature/csl/springer-mathphys-brackets.csl
authors:
  - name: Qian Qin
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
      - ref: 5
    # attributes:
    #   equal-contributor: true
  - name: Eli Bingham
    affiliations:
      - ref: 5
      - ref: 7
  - name: Gioele La Manno
    affiliations:
      - ref: 6
    email: gioele.lamanno@epfl.ch
    attributes:
      corresponding: true
  - name: David M. Langenau
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
    email: dlangenau@mgh.harvard.edu
    attributes:
      corresponding: true
  - name: Luca Pinello
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 5
    email: lpinello@mgh.harvard.edu
    attributes:
      corresponding: true
abstract: |
  Single-cell RNA Velocity has dramatically advanced our ability to model
  cellular differentiation and cell fate decisions. However, current
  preprocessing choices and model assumptions often lead to errors in assigning
  developmental trajectories. Here, we develop, Pyro-Velocity, a Bayesian,
  generative, and multivariate RNA Velocity model to estimate the uncertainty of
  cell future states. This approach models raw sequencing counts with the
  synchronized cell time across all expressed genes to provide quantifiable and
  improved information on cell fate choices and developmental trajectory
  dynamics.
keywords: [
  RNA velocity, 
  single-cell RNA sequencing, 
  probabilistic generative modeling, 
  cell fate, 
  developmental trajectory inference
]
bibliography: bibliography.bib
crossref:
  fig-prefix: Fig.
  sec-prefix: Sec.
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Fig. S
      space-before-numbering: false
      latex-list-of-description: Supplementary Figure
---

{{< pagebreak >}}

# Main text {#sec-main}

RNA Velocity is a powerful computational framework to estimate the time
derivative of gene expression
[@La_Manno2018-lj;@Svensson2018-vk;@Qiu2022-dj;@Bergen2020-pj]. The framework
has been used to study developmental cell lineage trajectories over a broad
spectrum of biological processes, such as  haematopoiesis [@Qiu2022-dj],
pancreas development [@Bastidas-Ponce2019-lf], and neural progenitors
differentiation [@Bergen2020-pj]. Yet, recent state-of-the-art RNA Velocity
approaches such as velocyto and scVelo [@La_Manno2018-lj] have several
limitations. First, they may predict developmental trajectories that do not
exist in nature, as it has been shown in the analysis of fully mature blood
cells [@Zheng2017-bz;@Bergen2021-qz]. Second, data preprocessing steps (e.g.,
selection of top variable genes, dimensionality reduction) affect velocity
estimates that are difficult to characterize and might lead to artifactual
predictions. In addition, non-linear spliced and unspliced imputation procedures
could distort phase portrait geometry and recovered trajectories
([@fig-model]a). Third, most RNA velocity algorithms perform estimates
independently for each gene, and therefore cannot exploit the natural
interdependence of gene expression. Fourth, a unified description of the
temporal relation between cells can only be assigned based on postprocessing
steps that aggregate heuristically the gene-wise fits into a shared latent time
([@fig-model]a). Finally, current RNA velocity frameworks only output point
estimates of the velocity vector field and do not provide uncertainty
estimation. Importantly, the absence of statistical confidence estimates
increases false discoveries as it leaves users without a way to evaluate the
robustness of the methods’ outputs. To overcome these limitations, we developed
a probabilistic generative model, named *Pyro&thinsp;-Velocity*, that estimates
RNA dynamics and infers cell trajectories using Pyro [@Bingham2018-id], a robust
and efficient probabilistic programming language.

Pyro-Velocity recasts the velocity estimation problem into a latent variable
posterior probability inference (@sec-methods). The proposed model is
generative and fully Bayesian, with the different parameters considered as
latent random variables. Each parameter is assigned a prior distribution, and
the model is then conditioned on the raw data (both spliced and unspliced
molecule raw counts) to estimate posterior probability distributions and to
enable uncertainty estimation of velocity. Central to the Pyro-Velocity model is
a *shared latent time* $(t)$: a global latent variable placing each cell at a
particular instant of the modeled dynamical process. From this latent time, the
expectation of spliced and unspliced RNA levels for each gene can be computed
knowing gene-specific kinetics parameters (transcriptional $(\alpha)$, splicing
$(\beta)$, and degradation rates ($\gamma$) and two gene-specific switching
times $(t_{0}^{k})$, and gene-specific transcriptional state $k$) and
integrating the RNA velocity differential equations analogously to what proposed
in the dynamical RNA velocity model by Bergen et al.
[@Bergen2020-pj;@Li2021-qa]. However, we innovate the structure and formulation
of the latent time given that our model is set to jointly fit the entire raw
data matrix, not individual genes. We approach the task of estimating the
posterior distribution by restricting the search within a family of a
variational distribution approximating the true posterior, and taking advantage
of the automatic differentiation variational inference (ADVI) capabilities of
Pyro [@Bingham2018-id;@Kucukelbir2016-bk] ([@fig-model]a,
@sec-methods). Notably, our generative Bayesian model enables to sample
from the posterior distributions of the latent and dependent variables, allowing
uncertainty estimation of phase portraits, high dimensional velocity vector
field, and shared latent time. Around this model, we built several
visualizations to assess uncertainty and functions to recover putative cell fate
priming genes and to investigate how the denoised gene expression recapitulates
known cell-type annotations and their developmental order ([@fig-model]b,
@sec-methods). Although several RNA velocity methods
[@Gayoso2024-fz;@Gorin2022-il;@Gu2022-rh;@Cui2022-em;@Gao2022-jq] have been
recently proposed, and many of them have not been peer-reviewed, here we
showcase the advantages and innovations of Pyro-Velocity by comparing it with
the current state-of-the-art dynamical RNA velocity model proposed by Bergen et
al. [@Bergen2020-pj], and implemented in scVelo.

# Methods {#sec-methods}

## Model formulation {#sec-model}

We assume the dynamical gene expression is determined by the RNA splicing
process, and infer the unspliced and spliced gene expression level from the
differential equations proposed in velocyto [@La_Manno2018-lj] and scVelo
[@Bergen2020-pj]

\begin{align}
\frac{d u\left(\tau^{\left(k_{cg}\right)}\right)}{d \tau^{\left(k_{cg}\right)}}
  &= \alpha^{\left(k_{cg}\right)}-\beta_g u\left(\tau^{\left(k_{cg}\right)}\right), \label{eq-dudt}\\
\frac{d s\left(\tau^{\left(k_{cg}\right)}\right)}{d \tau^{\left(k_{cg}\right)}}
  &= \beta_g u\left(\tau^{\left(k_{cg}\right)}\right)-\gamma_g s\left(\tau^{\left(k_{cg}\right)}\right). \label{eq-dsdt}
\end{align}

{{< pagebreak >}}

# References

:::{#refs}
:::

\FloatBarrier
{{< pagebreak >}}

# Figures

::: {#fig-model}
![](figures/fig.pdf)

Pyro-Velocity is a fully generative Bayesian method with uncertainty estimation
of velocity vector fields and shared latent time, based on raw counts and
without ad-hoc preprocessing steps. a. Schematic comparing current RNA Velocity
and Pyro-Velocity workflows. Both methods take in input spliced and unspliced
read count tables and output RNA velocity vector fields and latent (shared)
time. Previous methods require ad-hoc preprocessing steps, including the
arbitrary selection of top variable genes, number of principal components, and
kNN-pooling to impute both spliced and unspliced expression within the
PCA-reduced spliced expression space. In addition, manually tuned ad-hoc
postprocessing steps are used to compute the final vector field and latent time
based on aggregation and smoothing of the independent velocity estimates for
each gene which can dramatically change the velocity estimates. In contrast,
Pyro-Velocity is a fully generative and probabilistic model that uses raw read
counts. Our model requires minimal preprocessing and postprocessing steps and
outputs a common shared time based on all the genes without additional
aggregation steps. It also provides uncertainty estimation based on posterior
samples. b. The main outputs of the Pyro-Velocity framework. Visualization of
single gene expression across inferred developmental cell order using rainbow
plots (right panel).
:::

::: {#fig-model-results}
![](figures/fig.pdf)

Pyro-Velocity outperforms other RNA-velocity methods and accurately predicts
cell lineage trajectories. a,b. RNA Velocity model comparisons from fully mature
peripheral blood mononuclear cells (PBMCs, t-SNE embedded plots, a) or E15.5
mouse pancreas (UMAP embedded plots, b-c).  From left to right in panels a and
b: 1. Cells colored by cell type; 2. Stream plot of velocity vector field
defined by scVelo; 3. Stream plot of the posterior mean vector field from
Pyro-Velocity; 4. Pyro-Velocity visualization of cell dynamics for 3 arbitrarily
selected single cells; 5. Single-cell vector field uncertainty where each cell
is colored based on the scaled angular standard deviation to visualize
uncertainty based on n=30 iterated, posterior samples depicted as arrows noted
throughout our analysis; 6. Averaged vector field uncertainty where each vector
is colored by the averaged vector field uncertainty. The scaled angular standard
deviation range is [0 360], where 0 corresponds to the highest confidence for a
single direction, 360 corresponds to the highest uncertainty, and no
directionality preference based on 30 posterior samples (4-6). Vectors were
projected into the UMAP space using the scVelo API based on cosine
similarity-derived transition matrix. c. UMAP embedded cells colored by the
average cell shared time across 30 posterior samples. The Spearman’s correlation
between Pyro-Velocity and Cytotrace is provided. d. Pyro-Velocity automatic gene
selection and ranking for predicting important genes involved in pancreas
development. e. Phase portraits of posterior samples average spliced and
unspliced gene expression (left), rainbow plots showing denoised spliced gene
expression (middle), and gene expression within UMAP embedded space.  Single
dots correspond to individual cells (left, right) or vertical lines (middle).
Cell type annotation coloring is the same as in panel b.
:::

::: {#fig-lineage-tracing}
![](figures/fig.pdf)

Pyro-Velocity accurately predicts cell fate choices and lineage trajectory in
human LARRY barcoded hematopoietic cells. Systematic benchmarking of cell fate
prediction methods - predicting uni-lineage cell fates (a-b, Model 1),
bifurcated lineages (c, Model 2), or all lineages together (d, Model 2). Panels
from left to right show on the same SPRING embedding: 1. Cell types; 2. Clonal
progression vector field derived based on the observed clonal barcodes
(Methods); 3. scVelo recovered vector field; 4. Pyro-Velocity single cell level
vector field uncertainty (angular standard deviation from 30 posterior samples);
5. Pyro-Velocity recovered vector field; vectors are colored by the averaged
vector field uncertainty; 6. Clonal fate potency score using Cospar
[@Wang2022-xb]; 7. scVelo recovered latent time; 8. Pyro-Velocity inference of
posterior mean shared time per cell. Cosine similarity and Spearman’s
correlation for the different methods with the clonal progression and fate
potency are reported in the titles.
:::

\FloatBarrier
\clearpage
{{< pagebreak >}}

# Supplementary Figures

::: {#suppfig-graphical-models}
![](figures/fig.pdf)

Graphical representations detail the generative process of the Pyro-Velocity
models and highlight their differences. a. Pyro-Velocity Model 1. Each rectangle
represents one data dimension (e.g., gene or cell). The arrows represent the
directed hierarchical dependency among random variables. The white circles
correspond to the modeled random variables. The probability distributions of
these variables are denoted on the right legend. The gray circles correspond to
intermediate results computed from random variables. In this model, the kinetics
random variables for the cell shared time and gene are modeled with two
constraints: 1. the shared time is strictly larger than the gene-specific
activation time $t_{0}^{(0)}$, 2. the gene expression in the activate state
starts with the initial condition $(u_{0}=0,\,s_{0}=0)$, as in previous
methods[@La_Manno2018-lj;@Bergen2020-pj] (Methods). b. Pyro-Velocity Model 2.
This figure follows the same graphical convention as in a. In this model
$t_{0}^{(0)}$ is independent from the shared time, thus allowing a time lag for
the gene activation phase. In addition, a positive starting basal gene
expression level $(u_{0},\,s_{0})$ different than $(u_{0}=0,\,s_{0}=0)$ is
allowed. (Methods). c. Schematic illustrating how the constraints proposed in
Model 1 (left) and Model 2 (right) are reflected in the modeling of individual
gene expression. The x-axis corresponds to the shared time, the y-axis
corresponds to the posterior mean of the spliced expression level, the vertical
lines with arrows show the start and end time of the observed cell sampling, t0
represents the gene-specific activation time $t_{0}^{(0)}$, the red lines
highlight the differences for the potential gene-specific time lag when
$t_{0}^{(0)} > {\min}\!⁡(t)$, and the initial spliced gene expression. The same
differences can also be illustrated for the unspliced gene expression profiles
(not shown).
:::

::: {#suppfig-model-results}
![](figures/fig.pdf)

Extended version of @fig-model-results for evaluating Pyro-Velocity (Model 1)
and scVelo on the PBMC and pancreas datasets. a. Left figure shows the scVelo
vector field using the default user options, right figure shows the evaluation
of the scVelo vector field using different combinations of user options within
Epsilon cells. b. Pyro-Velocity additional single cell level uncertainty
diagnostic plots on the fully mature PBMC dataset, from left to right: 1.
standard deviation of posterior samples of shared time, 2. Rayleigh test of
posterior samples of the vector field (the title reports false discovery rate
after p-value correction for multiple tests across cells), 3. transcriptional
state uncertainty of predicted raw read count per cell. The contour lines show
the top 10% of the most uncertain cells. c. For the pancreas dataset, the
figures show from left to right: UMAP rendering of 1. Cytotrace, 2. scVelo
latent time, 3. Posterior mean of Pyro-Velocity shared time averaged across 30
posterior samples, 4. Top 50 markers overlapping between Pyro-Velocity and
scVelo (Methods, Supplementary Table 1-2), 5. The same uncertainty diagnostic
plots per cell as in b. d. The same plot as Fig. 2d. e. Selection of the top
negatively correlated genes with the cell shared time based on the posterior
mean on denoised spliced expression and negative mean absolute error of gene
expression prediction (Methods). f. Phase portraits, rainbow plots, and UMAP
rendering of posterior mean of gene expression for the genes selected in e.
:::

::: {#suppfig-model-2-result}
![](figures/fig.pdf)

Extension of @fig-model-results, Pyro-Velocity (Model 2) outperforms scVelo on
PBMC and pancreas datasets. Same format as in @fig-model-results and
@suppfig-model-results
:::

::: {#suppfig-lineage-tracing}
![](figures/fig.pdf)

Extension of @fig-lineage-tracing, Pyro-velocity (Model 1 for uni-fate data,
Model 2 for bi-fate and multi-fate data) outperforms Cytotrace on scRNA-seq
datasets with lineage barcoding information (LARRY dataset). a. SPRING embedding
rendering of Cytotrace scores, the title reports the Spearman’s correlation
between the Cytotrace and Cospar fate potency scores, b. Pyro-Velocity
additional single cell level uncertainty diagnostic plots based on the
multi-fate LARRY dataset, from left to right: 1. Standard deviation of posterior
samples of shared time per cell, 2. Rayleigh test of posterior samples of vector
field per cell (the title reports the false discovery rate after p-value
correction for multiple tests across cells), 3. Transcriptional state
uncertainty of predicted raw read count per cell. The contour lines show the top
10% of the most uncertain cells (Methods), c. Pairwise Spearman’s correlation
between Cytotrace/Cospar fate potency scores and velocity methods for the four
different benchmarking datasets selected based on the lineage barcoding
information from LARRY data. d. SPRING embedding rendering of posterior mean of
shared time across 30 posterior samples, the title shows the Spearman’s
correlation of mean shared time with Cytotrace. e. Selection of the top
positively correlated genes with the cell shared time based on the posterior
mean of spliced expression. f. Phase portraits, rainbow plots, and UMAP
rendering with splicing gene expression level for the selected genes in e.
:::

::: {#suppfig-lineage-tracing-2}
![](figures/fig.pdf)

Extension of @fig-lineage-tracing presenting an extended comparison of
Pyro-Velocity and scVelo on scRNA-seq datasets with lineage barcoding
information (LARRY dataset). a. uni-fate Monocyte lineage analysis with
Pyro-Velocity Model 2, b. uni-fate Neutrophil lineage analysis with
Pyro-Velocity Model 2, c. Neutrophil and Monocyte bi-fate lineage analysis with
Pyro-Velocity Model 1, and d. multi-fate lineage analysis with Pyro-Velocity
Model 1. The panel descriptions are the same as in @fig-lineage-tracing
:::

![An example figure (an empty plot)](figures/fig.pdf){#suppfig-plot}

<!-- \begin{table}[h]
\caption{LaTeX caption text}\label{tbl-data}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2  & Column 3 & Column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
\end{table} -->

\FloatBarrier
{{< pagebreak >}}

# Supporting Information {.appendix}