### 2024-08-20 reproducibility/manuscript/v2.tex
chore(manuscript): update v2 diff file
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-mathphys-num,
  lineno,
  twocolumn]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\usepackage{nameref}
\usepackage{placeins}
\allowdisplaybreaks
\makeatletter
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{plain}
\@ifundefined{c@chapter}{\newfloat{suppfig}{h}{losuppfig}}{\newfloat{suppfig}{h}{losuppfig}[chapter]}
\floatname{suppfig}{Supplementary Figure}
\newcommand*\listofsuppfigs{\listof{suppfig}{List of Supplementary Figures}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Pyro -Velocity: Probabilistic RNA Velocity inference from single-cell data},
  pdfkeywords={RNA velocity, single-cell RNA sequencing, probabilistic
generative modeling, cell fate, developmental trajectory inference},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Pyro -Velocity: Probabilistic RNA Velocity inference from
single-cell data]{Pyro -Velocity: Probabilistic RNA Velocity inference
from single-cell data}

% author setup
\author[1,2,3,4,5]{\fnm{Qian} \sur{Qin}}\author[5,7]{\fnm{Eli} \sur{Bingham}}\author*[6]{\fnm{Gioele La} \sur{Manno}}\email{gioele.lamanno@epfl.ch}\author*[1,2,3,4]{\fnm{David M.} \sur{Langenau}}\email{dlangenau@mgh.harvard.edu}\author*[1,2,5]{\fnm{Luca} \sur{Pinello}}\email{lpinello@mgh.harvard.edu}
% affil setup
\affil[1]{\orgdiv{Molecular Pathology Unit}, \orgname{Massachusetts
General Hospital Research Institute, Charlestown, MA}}
\affil[2]{\orgdiv{Massachusetts General Hospital Cancer
Center}, \orgname{Harvard Medical School, Charlestown, MA}}
\affil[3]{\orgdiv{Center for Regenerative
Medicine}, \orgname{Massachusetts General Hospital, Boston, MA}}
\affil[4]{\orgdiv{Harvard Stem Cell Institute}, \orgname{Harvard
University, Cambridge, MA}}
\affil[5]{\orgdiv{Broad Institute of MIT and
Harvard}, \orgname{Cambridge, MA}}
\affil[6]{\orgdiv{Brain and Mind Institute, School of Life Sciences,
École Polytechnique Fédérale de Lausanne (EPFL)}, \orgname{Laboratory of
Neurodevelopmental Systems Biology, Lausanne, Switzerland}}
\affil[7]{\orgdiv{Basis Research Institute}, \orgname{Cambridge, MA}}

% abstract

\abstract{Single-cell RNA Velocity has dramatically advanced our ability
to model cellular differentiation and cell fate decisions. However,
current preprocessing choices and model assumptions often lead to errors
in assigning developmental trajectories. Here, we develop,
Pyro -Velocity, a Bayesian, generative, and multivariate RNA Velocity
model to estimate the uncertainty of cell future states. This approach
models raw sequencing counts with the synchronized cell time across all
expressed genes to provide quantifiable and improved information on cell
fate choices and developmental trajectory dynamics.}

% keywords
\keywords{RNA velocity,  single-cell RNA sequencing,  probabilistic
generative modeling,  cell fate,  developmental trajectory inference}

\begin{document}
\maketitle
\newpage{}

\renewcommand*\contentsname{Contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\newpage{}

\section{Main text}\label{sec-main}

RNA Velocity is a powerful computational framework to estimate the time
derivative of gene expression
\citep{La_Manno2018-lj, Svensson2018-vk, Qiu2022-dj, Bergen2020-pj}. The
framework has been used to study developmental cell lineage trajectories
over a broad spectrum of biological processes, such as haematopoiesis
\citep{Qiu2022-dj}, pancreas development \citep{Bastidas-Ponce2019-lf},
and neural progenitors differentiation \citep{Bergen2020-pj}. Yet,
recent state-of-the-art RNA Velocity approaches such as velocyto and
scVelo \citep{La_Manno2018-lj} have several limitations. First, they may
predict developmental trajectories that do not exist in nature, as it
has been shown in the analysis of fully mature blood cells
\citep{Zheng2017-bz, Bergen2021-qz}. Second, data preprocessing steps
(e.g., selection of top variable genes, dimensionality reduction) affect
velocity estimates that are difficult to characterize and might lead to
artifactual predictions. In addition, non-linear spliced and unspliced
imputation procedures could distort phase portrait geometry and
recovered trajectories (Fig.~\ref{fig-model}a). Third, most RNA velocity
algorithms perform estimates independently for each gene, and therefore
cannot exploit the natural interdependence of gene expression. Fourth, a
unified description of the temporal relation between cells can only be
assigned based on postprocessing steps that aggregate heuristically the
gene-wise fits into a shared latent time (Fig.~\ref{fig-model}a).
Finally, current RNA velocity frameworks only output point estimates of
the velocity vector field and do not provide uncertainty estimation.
Importantly, the absence of statistical confidence estimates increases
false discoveries as it leaves users without a way to evaluate the
robustness of the methods' outputs. To overcome these limitations, we
developed a probabilistic generative model, named \emph{Pyro -Velocity},
that estimates RNA dynamics and infers cell trajectories using Pyro
\citep{Bingham2018-id}, a robust and efficient probabilistic programming
language.

Pyro -Velocity recasts the velocity estimation problem into a latent
variable posterior probability inference (Sec.~\ref{sec-methods}). The
proposed model is generative and fully Bayesian, with the different
parameters considered as latent random variables. Each parameter is
assigned a prior distribution, and the model is then conditioned on the
raw data (both spliced and unspliced molecule raw counts) to estimate
posterior probability distributions and to enable uncertainty estimation
of velocity. Central to the Pyro -Velocity model is a \emph{shared
latent time} \((t)\): a global latent variable placing each cell at a
particular instant of the modeled dynamical process. From this latent
time, the expectation of spliced and unspliced RNA levels for each gene
can be computed knowing gene-specific kinetics parameters
(transcriptional \((\alpha)\), splicing \((\beta)\), and degradation
rates (\(\gamma\)) and two gene-specific switching times
\((t_{0}^{k})\), and gene-specific transcriptional state \(k\)) and
integrating the RNA velocity differential equations analogously to what
proposed in the dynamical RNA velocity model by Bergen et al.
\citep{Bergen2020-pj, Li2021-qa}. However, we innovate the structure and
formulation of the latent time given that our model is set to jointly
fit the entire raw data matrix, not individual genes. We approach the
task of estimating the posterior distribution by restricting the search
within a family of a variational distribution approximating the true
posterior, and taking advantage of the automatic differentiation
variational inference (ADVI) capabilities of Pyro
\citep{Bingham2018-id, Kucukelbir2016-bk} (Fig.~\ref{fig-model}a,
Sec.~\ref{sec-methods}). Notably, our generative Bayesian model enables
to sample from the posterior distributions of the latent and dependent
variables, allowing uncertainty estimation of phase portraits, high
dimensional velocity vector field, and shared latent time. Around this
model, we built several visualizations to assess uncertainty and
functions to recover putative cell fate priming genes and to investigate
how the denoised gene expression recapitulates known cell-type
annotations and their developmental order (Fig.~\ref{fig-model}b,
Sec.~\ref{sec-methods}). Although several RNA velocity methods
\citep{Gayoso2024-fz, Gorin2022-il, Gu2022-rh, Cui2022-em, Gao2022-jq}
have been recently proposed, and many of them have not been
peer-reviewed, here we showcase the advantages and innovations of
Pyro -Velocity by comparing it with the current state-of-the-art
dynamical RNA velocity model proposed by Bergen et al.
\citep{Bergen2020-pj}, and implemented in scVelo.

One major biological hurdle for current formulations of RNA Velocity
models has been miscalling cell trajectories in fully differentiated
cell populations that lack dynamic cell state transitions. One typical
example being human peripheral blood mononuclear cells (PBMCs)
\citep{Zheng2017-bz}. In this dataset, 65,877 cells were analyzed using
the 10x platform that comprised 11 fully differentiated, mature immune
cell types. This dataset lacked stem and progenitor cells or other
signatures of an undergoing dynamical differentiation; thus, no
consistent velocity flow should be detected. We reanalyzed this dataset
using scVelo and Pyro -Velocity using a common set of genes for a fair
comparison. Using scVelo, we reproduced the original results showing
incorrect cell trajectory assignments (Fig.~\ref{fig-model-results}a)
\citep{Bergen2021-qz}. ScVelo uncovered an errant, strong linear
trajectory across a subset of mature cell types \citep{Zheng2017-bz}
that is exasperated by the inability to confidently assess lineage
directionality or a baseline to assess significance. On the contrary,
Pyro -Velocity failed to detect high-confidence trajectories in the
mature blood cell states, consistent with what is known about the
biology underlying these cells (Fig.~\ref{fig-model-results}a, panel 2
vs.~panel 3). Importantly, the Bayesian nature of Pyro -Velocity enabled
the generation of uncertainty estimates that can be easily visualized as
likely vectors for individual cells or averaged across cells on a
constantly spaced grid on the cell embedding (\emph{averaged vector
field}).

Pyro -Velocity was also able to define well-known developmental cell
hierarchies in a scRNA-seq dataset of E15.5 mouse pancreas
(Fig.~\ref{fig-model-results}b-e, \(n=3,696\) cells comprising \(8\)
cell types (Fig.~\ref{fig-model-results}b, panel 1)), identifying cell
trajectories originating from ductal progenitor cells and culminated in
the production of mature Alpha, Beta, Delta, and Epsilon cells
\citep{Bastidas-Ponce2019-lf}. Though the scVelo and Pyro -Velocity
models have qualitatively comparable vector fields, the results
generated by scVelo were greatly influenced by gene selection,
neighboring cell number for gene expression smoothing, and cannot
provide confidence assessment of lineage directionality
(Supplementary Figure~\ref{suppfig-model-results}a). In addition,
Pyro -Velocity uncertainty visualizations uncovered that differentiated
cells had a larger vector field uncertainty than stem and progenitor
cells, as would be expected for terminally differentiated cell states
where progression and lineage trajectories would not be anticipated.
Pyro -Velocity also recovered a developmental order in close agreement
with what was estimated by Cytotrace (Spearman's
\(r=0.95, \; \text{p-value} <
10^{-10}\) between latent time and developmental
order)\citep{Gulati2020-xq} - a state-of-the-art orthogonal method used
to predict cell differentiation based on the number of expressed genes
per cell.

We next used Pyro -Velocity to unbiasedly uncover potential cell fate
determinant markers genes within the E15.5 mouse pancreas. To this end,
we first selected the top 300 genes with the best goodness of fit
(negative mean absolute errors across cells). Next, we ranked them based
on the positive (or negative) correlation between inferred shared time
and corresponding denoised spliced expression per cell
(Fig.~\ref{fig-model-results}d, Methods). This procedure successfully
recovered key developmental genes that are activated (or repressed)
during endocrinogenesis (Fig.~\ref{fig-model-results}d-e,
Supplementary Figure~\ref{suppfig-model-results}). Among the top
positively correlated genes identified by Pyro -Velocity, endocrine Beta
cells expressed insulin 1 and 2 (Ins1, Ins2) and Nnat, which is a key
regulator Beta cell insulin generation \citep{Millership2018-vd} and
Ppp1r1a which is a marker of Beta cells \citep{Jiang2013-pt}. The
ranking procedure proposed by scVelo failed to identify Ins1, Ins2,
Nnat, and Ppp1r1a as lineage-restricted genes in their top 50 ranked
gene list. The inability of scVelo to identify these key developmental
genes could be likely explained by the kNN smoothing of read count and
filtering steps (Supplementary Table 1, 2). Visual inspection of these
four top genes by Pyro -Velocity revealed a single activation phase
during cell type transitions based on phase portrait curves. We also
propose their visualization through ``rainbow plots'\,' to illustrate
denoised single gene spliced expression across shared time and to
investigate the inferred cell order based on known annotations. Further,
UMAP embedding of cells colored based on denoised gene expression
confirmed the cell fate markers' smooth and progressive activation (or
repression) towards endocrine cells (Fig.~\ref{fig-model-results}e,
Supplementary Figure~\ref{suppfig-model-results}).

Ground truth velocity is not a quantity we can easily assess. Therefore,
no robust and quantitative benchmarking of RNA velocity methods has been
proposed. However, recent single-cell assays allow recording lineage
relationships and can be used to assess the ability of different methods
in recapitulating cell fate progression. Therefore, we next
systematically compared scVelo and Pyro -Velocity ability to recover
single cell fates in Lineage And RNA RecoverY (LARRY) barcoded
hematopoietic cells. LARRY leverages unique lentiviral barcodes inserted
into 3' untranslated region (UTR) of GFP reporter that is integrated as
a single copy into parental cells. Then progeny can be clonally traced
to follow the descendent cell fates over time. LARRY has been recently
applied to in vitro differentiation of human blood and accurately
recovered expected cell fate and lineage trajectories
\citep{Weinreb2020-mv}. This dataset sampled differentiation over the
course of 2, 4, and 6 days and contained \(49,302\) cells that each
could be traced with at least one barcode. Because LARRY couples
single-cell transcriptomics and direct cell lineage tracing, this data
set provides a unique opportunity to benchmark RNA Velocity
quantitatively and to provide insights into the correct interpretation
of recovered velocity vector fields and latent cell times. To
quantitatively assess the agreement between observed and predicted
fates, we considered two metrics: 1. average cosine similarity (cs)
between vectors from the RNA Velocity vector field and the clonal
progression vector field, i.e., a vector field derived directly from the
lineage data (Sec.~\ref{sec-methods}), 2. Spearman's correlation
\((\rho)\) between cell latent (shared) time and fate potency scores
derived from Cytotrace and Cospar. Cospar, our gold standard, is a
state-of-the-art method designed explicitly for predicting fate potency
based on LARRY data, while Cytotrace leverages only transcriptomic data.
We selected cell state changes associated with monocyte and neutrophil
lineages because both lineages have the largest numbers of cells in the
scRNA sequencing data set and well-known cell fate trajectories
\citep{Petukhov2018-nr}. As expected, the Pyro -Velocity vector field
accurately predicted cell state trajectories from hematopoietic
progenitor cells into mature monocytes or neutrophils and had high
cosine similarity between the inferred vector field and clonal
progression vector field assigned by LARRY (monocytes \(cs=0.49\),
neutrophils \(cs=0.38\)). High Spearman's correlation was also seen
between shared latent time and Cospar fate potency (monocytes
\(\rho=0.60\), neutrophils \(\rho=0.74\)) or shared latent time with
Cytotrace fate potency (monocytes \(\rho=0.925\), neutrophils
\(\rho=0.925\)). By contrast, scVelo generated vector fields with
reversed order or low correlation (cosine similarity: \(-0.38\),
monocytes, cosine similarity: \(0.32\), neutrophils). Moreover, Cospar
fate potency scores were negative (\(\rho=-0.59\), monocytes) or
non-significantly correlated (\(\rho=-0.07\), neutrophils). A similar
low correlation was observed for scVelo and Cytotrace fate potency
scores (monocytes \(\rho=0.309\), neutrophils \(\rho=0.751\))
(Fig.~\ref{fig-lineage-tracing}a-b,
Supplementary Figure~\ref{suppfig-lineage-tracing}c).

Next, we evaluated the ability of the Pyro -Velocity models to recover
bi-fate or multi-fate trajectories by considering clones belonging to
monocyte and neutrophil lineages or using the entire dataset. Not
surprisingly, both methods showed their limitations, given that they are
not designed to model and capture multiple fate trajectories
(Supplementary Figure~\ref{suppfig-lineage-tracing-2}c-d). The recovered
vector fields and cell latent (shared) time show inconsistent directions
for both methods. However, Pyro -Velocity identified higher uncertainty
for some of the problematic regions. Furthermore, we discovered that
vector field and shared time estimation could be significantly improved
by extending the original kinetic model by allowing each gene a time lag
for the transcriptional activation and a baseline expression level
potentially different than 0 (Pyro -Velocity Model 2,
Sec.~\ref{sec-methods}) (Fig.~\ref{fig-lineage-tracing}c,d,
Supplementary Figure~\ref{suppfig-graphical-models}b,c). This extended
model recapitulated a vector field consistent with the clonal
progression vector field for most branches with some spurious
predictions within the intermediate cell state between stem cells and
Megakaryocytes and most terminal Neutrophils (bi-fate cosine similarity:
\(0.62\), multi-fate cosine similarity: \(0.63\)). However, the proposed
uncertainty estimation accurately flagged these subpopulations to warn
users of potential false discoveries. Notably, this extended model
significantly improved the correlation of cell shared time with the
clonal fate potency (bi-fate \(\rho=0.65\), multi-fate \(\rho=0.74\))
and also improved the correlation with Cytotrace fate potency (bi-fate
\(\rho=0.931\), multi-fate \(\rho=0.614\)). Remarkably, the shared time
per cell from Pyro -Velocity also outperformed Cytotrace for
developmental order prediction
(Supplementary Figure~\ref{suppfig-lineage-tracing}a,c). Importantly,
this extended model retained its ability to accurately identify cell
lineage trajectories and shared time for all the other datasets analyzed
above (Supplementary Figure~\ref{suppfig-model-2-result}). Taken
together, our approach using LARRY barcoded datasets established the
first quantitative benchmark of RNA velocity-based cell fate trajectory
inference and provides an exciting new framework for continued
optimization of multi-fate predictions in the future.

\section{Methods}\label{sec-methods}

\subsection{Model formulation}\label{sec-model}

We assume the dynamical gene expression is determined by the RNA
splicing process, and infer the unspliced and spliced gene expression
level from the differential equations proposed in velocyto
\citep{La_Manno2018-lj} and scVelo \citep{Bergen2020-pj}

\begin{align}
\frac{d u\left(\tau^{\left(k_{cg}\right)}\right)}{d \tau^{\left(k_{cg}\right)}}
  &= \alpha^{\left(k_{cg}\right)}-\beta_g u\left(\tau^{\left(k_{cg}\right)}\right), \label{eq-dudt}\\
\frac{d s\left(\tau^{\left(k_{cg}\right)}\right)}{d \tau^{\left(k_{cg}\right)}}
  &= \beta_g u\left(\tau^{\left(k_{cg}\right)}\right)-\gamma_g s\left(\tau^{\left(k_{cg}\right)}\right). \label{eq-dsdt}
\end{align}

\newpage{}

\onecolumn

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\FloatBarrier

\newpage{}

\section{Figures}\label{figures}

\begin{figure}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{fig-model}Pyro -Velocity is a fully generative Bayesian
method with uncertainty estimation of velocity vector fields and shared
latent time, based on raw counts and without ad-hoc preprocessing steps.
a. Schematic comparing current RNA Velocity and Pyro -Velocity
workflows. Both methods take in input spliced and unspliced read count
tables and output RNA velocity vector fields and latent (shared) time.
Previous methods require ad-hoc preprocessing steps, including the
arbitrary selection of top variable genes, number of principal
components, and kNN-pooling to impute both spliced and unspliced
expression within the PCA-reduced spliced expression space. In addition,
manually tuned ad-hoc postprocessing steps are used to compute the final
vector field and latent time based on aggregation and smoothing of the
independent velocity estimates for each gene which can dramatically
change the velocity estimates. In contrast, Pyro -Velocity is a fully
generative and probabilistic model that uses raw read counts. Our model
requires minimal preprocessing and postprocessing steps and outputs a
common shared time based on all the genes without additional aggregation
steps. It also provides uncertainty estimation based on posterior
samples. b. The main outputs of the Pyro -Velocity framework.
Visualization of single gene expression across inferred developmental
cell order using rainbow plots (right panel).}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{fig-model-results}Pyro -Velocity outperforms other
RNA-velocity methods and accurately predicts cell lineage trajectories.
a,b. RNA Velocity model comparisons from fully mature peripheral blood
mononuclear cells (PBMCs, t-SNE embedded plots, a) or E15.5 mouse
pancreas (UMAP embedded plots, b-c). From left to right in panels a and
b: 1. Cells colored by cell type; 2. Stream plot of velocity vector
field defined by scVelo; 3. Stream plot of the posterior mean vector
field from Pyro -Velocity; 4. Pyro -Velocity visualization of cell
dynamics for 3 arbitrarily selected single cells; 5. Single-cell vector
field uncertainty where each cell is colored based on the scaled angular
standard deviation to visualize uncertainty based on n=30 iterated,
posterior samples depicted as arrows noted throughout our analysis; 6.
Averaged vector field uncertainty where each vector is colored by the
averaged vector field uncertainty. The scaled angular standard deviation
range is {[}0 360{]}, where 0 corresponds to the highest confidence for
a single direction, 360 corresponds to the highest uncertainty, and no
directionality preference based on 30 posterior samples (4-6). Vectors
were projected into the UMAP space using the scVelo API based on cosine
similarity-derived transition matrix. c.~UMAP embedded cells colored by
the average cell shared time across 30 posterior samples. The Spearman's
correlation between Pyro -Velocity and Cytotrace is provided.
d.~Pyro -Velocity automatic gene selection and ranking for predicting
important genes involved in pancreas development. e. Phase portraits of
posterior samples average spliced and unspliced gene expression (left),
rainbow plots showing denoised spliced gene expression (middle), and
gene expression within UMAP embedded space. Single dots correspond to
individual cells (left, right) or vertical lines (middle). Cell type
annotation coloring is the same as in panel b.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{fig-lineage-tracing}Pyro -Velocity accurately predicts
cell fate choices and lineage trajectory in human LARRY barcoded
hematopoietic cells. Systematic benchmarking of cell fate prediction
methods - predicting uni-lineage cell fates (a-b, Model 1), bifurcated
lineages (c, Model 2), or all lineages together (d, Model 2). Panels
from left to right show on the same SPRING embedding: 1. Cell types; 2.
Clonal progression vector field derived based on the observed clonal
barcodes (Methods); 3. scVelo recovered vector field; 4. Pyro -Velocity
single cell level vector field uncertainty (angular standard deviation
from 30 posterior samples); 5. Pyro -Velocity recovered vector field;
vectors are colored by the averaged vector field uncertainty; 6. Clonal
fate potency score using Cospar \citep{Wang2022-xb}; 7. scVelo recovered
latent time; 8. Pyro -Velocity inference of posterior mean shared time
per cell. Cosine similarity and Spearman's correlation for the different
methods with the clonal progression and fate potency are reported in the
titles.}

\end{figure}%

\FloatBarrier
\clearpage

\newpage{}

\section{Supplementary Figures}\label{supplementary-figures}

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-graphical-models}Graphical representations
detail the generative process of the Pyro -Velocity models and highlight
their differences. a. Pyro -Velocity Model 1. Each rectangle represents
one data dimension (e.g., gene or cell). The arrows represent the
directed hierarchical dependency among random variables. The white
circles correspond to the modeled random variables. The probability
distributions of these variables are denoted on the right legend. The
gray circles correspond to intermediate results computed from random
variables. In this model, the kinetics random variables for the cell
shared time and gene are modeled with two constraints: 1. the shared
time is strictly larger than the gene-specific activation time
\(t_{0}^{(0)}\), 2. the gene expression in the activate state starts
with the initial condition \((u_{0}=0,\,s_{0}=0)\), as in previous
methods\citep{La_Manno2018-lj, Bergen2020-pj} (Methods). b.
Pyro -Velocity Model 2. This figure follows the same graphical
convention as in a. In this model \(t_{0}^{(0)}\) is independent from
the shared time, thus allowing a time lag for the gene activation phase.
In addition, a positive starting basal gene expression level
\((u_{0},\,s_{0})\) different than \((u_{0}=0,\,s_{0}=0)\) is allowed.
(Methods). c.~Schematic illustrating how the constraints proposed in
Model 1 (left) and Model 2 (right) are reflected in the modeling of
individual gene expression. The x-axis corresponds to the shared time,
the y-axis corresponds to the posterior mean of the spliced expression
level, the vertical lines with arrows show the start and end time of the
observed cell sampling, t0 represents the gene-specific activation time
\(t_{0}^{(0)}\), the red lines highlight the differences for the
potential gene-specific time lag when \(t_{0}^{(0)} > {\min}\!⁡(t)\), and
the initial spliced gene expression. The same differences can also be
illustrated for the unspliced gene expression profiles (not shown).}

\end{suppfig}%

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-model-results}Extended version of
Fig.~\ref{fig-model-results} for evaluating Pyro -Velocity (Model 1) and
scVelo on the PBMC and pancreas datasets. a. Left figure shows the
scVelo vector field using the default user options, right figure shows
the evaluation of the scVelo vector field using different combinations
of user options within Epsilon cells. b. Pyro -Velocity additional
single cell level uncertainty diagnostic plots on the fully mature PBMC
dataset, from left to right: 1. standard deviation of posterior samples
of shared time, 2. Rayleigh test of posterior samples of the vector
field (the title reports false discovery rate after p-value correction
for multiple tests across cells), 3. transcriptional state uncertainty
of predicted raw read count per cell. The contour lines show the top
10\% of the most uncertain cells. c.~For the pancreas dataset, the
figures show from left to right: UMAP rendering of 1. Cytotrace, 2.
scVelo latent time, 3. Posterior mean of Pyro -Velocity shared time
averaged across 30 posterior samples, 4. Top 50 markers overlapping
between Pyro -Velocity and scVelo (Methods, Supplementary Table 1-2), 5.
The same uncertainty diagnostic plots per cell as in b. d.~The same plot
as Fig. 2d. e. Selection of the top negatively correlated genes with the
cell shared time based on the posterior mean on denoised spliced
expression and negative mean absolute error of gene expression
prediction (Methods). f.~Phase portraits, rainbow plots, and UMAP
rendering of posterior mean of gene expression for the genes selected in
e.}

\end{suppfig}%

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-model-2-result}Extension of
Fig.~\ref{fig-model-results}, Pyro -Velocity (Model 2) outperforms
scVelo on PBMC and pancreas datasets. Same format as in
Fig.~\ref{fig-model-results} and
Supplementary Figure~\ref{suppfig-model-results}}

\end{suppfig}%

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-lineage-tracing}Extension of
Fig.~\ref{fig-lineage-tracing}, Pyro -Velocity (Model 1 for uni-fate
data, Model 2 for bi-fate and multi-fate data) outperforms Cytotrace on
scRNA-seq datasets with lineage barcoding information (LARRY dataset).
a. SPRING embedding rendering of Cytotrace scores, the title reports the
Spearman's correlation between the Cytotrace and Cospar fate potency
scores, b. Pyro -Velocity additional single cell level uncertainty
diagnostic plots based on the multi-fate LARRY dataset, from left to
right: 1. Standard deviation of posterior samples of shared time per
cell, 2. Rayleigh test of posterior samples of vector field per cell
(the title reports the false discovery rate after p-value correction for
multiple tests across cells), 3. Transcriptional state uncertainty of
predicted raw read count per cell. The contour lines show the top 10\%
of the most uncertain cells (Methods), c.~Pairwise Spearman's
correlation between Cytotrace/Cospar fate potency scores and velocity
methods for the four different benchmarking datasets selected based on
the lineage barcoding information from LARRY data. d.~SPRING embedding
rendering of posterior mean of shared time across 30 posterior samples,
the title shows the Spearman's correlation of mean shared time with
Cytotrace. e. Selection of the top positively correlated genes with the
cell shared time based on the posterior mean of spliced expression.
f.~Phase portraits, rainbow plots, and UMAP rendering with splicing gene
expression level for the selected genes in e.}

\end{suppfig}%

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-lineage-tracing-2}Extension of
Fig.~\ref{fig-lineage-tracing} presenting an extended comparison of
Pyro -Velocity and scVelo on scRNA-seq datasets with lineage barcoding
information (LARRY dataset). a. uni-fate Monocyte lineage analysis with
Pyro -Velocity Model 2, b. uni-fate Neutrophil lineage analysis with
Pyro -Velocity Model 2, c.~Neutrophil and Monocyte bi-fate lineage
analysis with Pyro -Velocity Model 1, and d.~multi-fate lineage analysis
with Pyro -Velocity Model 1. The panel descriptions are the same as in
Fig.~\ref{fig-lineage-tracing}}

\end{suppfig}%

\begin{suppfig}

\centering{

\includegraphics{figures/fig.pdf}

}

\caption{\label{suppfig-plot}An example figure (an empty plot)}

\end{suppfig}%

\FloatBarrier

\newpage{}

\renewcommand\refname{Supporting Information}

\end{document}