### 2025-05-12 nbs/development/index.qmd
docs(development): init
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Development Overview
format:
  html:
    toc: true
---

## Introduction

This section provides technical documentation for developers working on PyroVelocity. It covers the architecture, implementation details, and testing framework of the project.

## Purpose

The development documentation is intended for:

- Contributors to the PyroVelocity codebase
- Researchers extending the models or algorithms
- Developers integrating PyroVelocity into other systems

## Contents

The development documentation is organized into the following sections:

- [Architecture](architecture.qmd): Detailed information about the modular architecture of PyroVelocity, including component interfaces, implementation patterns, and design decisions.
- [Testing](testing.qmd): Information about the testing framework, validation approach, and guidelines for writing tests.

## Development Workflow

PyroVelocity follows a protocol-first architecture with all contracts defined through Protocol interfaces. The project is organized into modular components that can be composed to create different model configurations.

For more information on contributing to PyroVelocity, please see the [Contributing](../about/contributing.qmd) guide.