### 2025-05-12 nbs/development/architecture.qmd
docs(development): init architecture
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Architecture
format:
  html:
    toc: true
---

## Overview

PyroVelocity has a modular architecture that separates concerns into distinct components. This document describes the architecture of the PyroVelocity codebase, focusing on the modular implementation in `src/pyrovelocity/models/modular`.

## Protocol-First Design

PyroVelocity follows a protocol-first architecture with all contracts defined through Protocol interfaces. Components implement these interfaces directly without inheritance, promoting composition over inheritance.

## Core Components

### Model Components

The modular implementation consists of the following key components:

- **Dynamics Models**: Implement the mathematical models of RNA dynamics
  - `StandardDynamicsModel`: The standard implementation of RNA dynamics
  - `LegacyDynamicsModel`: Compatibility implementation matching the legacy model

- **Prior Models**: Define the prior distributions for model parameters
  - `LogNormalPriorModel`: Uses log-normal distributions for parameters

- **Likelihood Models**: Define the observation models
  - `PoissonLikelihoodModel`: Uses Poisson distributions for RNA counts
  - `LegacyLikelihoodModel`: Compatibility implementation matching the legacy model

- **Observation Models**: Handle the interface between data and model
  - `StandardObservationModel`: The standard implementation for observations

- **Guide Factories**: Create guide functions for variational inference
  - `AutoGuideFactory`: Creates guides for the modular implementation
  - `LegacyAutoGuideFactory`: Compatibility implementation matching the legacy guide

### Utility Modules

The architecture includes several utility modules:

- `core_utils.py`: Core utility functions used across components
- `context_utils.py`: Utilities for managing Pyro contexts
- `pyro_utils.py`: Utilities specific to Pyro functionality

## Integration with AnnData

The modular implementation integrates directly with AnnData through methods in the `PyroVelocityModel` class, eliminating the need for adapter classes.

## Validation Framework

The architecture includes a validation framework to ensure that the modular implementation can replicate the behavior of the legacy implementation.

## Future Directions

The architecture is designed to be extensible, allowing for the addition of new components and features while maintaining backward compatibility with existing workflows.