### 2025-05-12 nbs/development/testing.qmd
docs(development): init testing
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Testing Framework
format:
  html:
    toc: true
---

## Overview

PyroVelocity uses a comprehensive testing framework to ensure code quality and correctness. This document describes the testing approach, tools, and guidelines for writing tests.

## Testing Philosophy

The testing philosophy of PyroVelocity emphasizes:

- Creating simple instances of real objects over using mocking tools
- Writing tests alongside code development
- Ensuring tests provide reasonable constraints on the code
- Systematic validation of model implementations

## Test Organization

Tests are organized in the following structure:

- `src/pyrovelocity/tests/`: Root directory for all tests
  - `models/`: Tests for model components
  - `plots/`: Tests for plotting functions
  - `utils/`: Tests for utility functions
  - `data/`: Test data fixtures

## Validation Framework

PyroVelocity includes a validation framework (PRD-678-VALID) to verify that the modular model can replicate the legacy model. Key aspects include:

- Direct comparison of model outputs
- Validation using synthetic data from test fixtures
- Validation using downsampled real datasets
- Systematic addressing of implementation differences

## Writing Tests

When writing tests for PyroVelocity:

1. Use pytest fixtures for test data and common setup
2. Create simple instances of real objects rather than mocks
3. Test both normal operation and edge cases
4. Use xdoctests with proper type annotations
5. Ensure tests are self-contained with synthetic data

## Running Tests

To run tests:

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests/
```

To run specific test modules:

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests/models/test_model.py
```

## Profiling and Benchmarking

PyroVelocity uses Scalene for profiling and benchmarking:

```bash
source .venv/bin/activate
scalene --profile-all src/pyrovelocity/tests/models/test_model.py
```

## Continuous Integration

Tests are automatically run as part of the CI/CD pipeline defined in `.github/workflows/cid.yaml`.