### 2025-05-12 nbs/development/testing.qmd
docs(development): update testing docs
- note integration of bdd docs

Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: Testing Framework and Reference
format:
  html:
    toc: true
    toc-depth: 3
---

# PyroVelocity Testing

This documentation describes the testing approach and infrastructure for the `pyrovelocity` library. Our testing strategy combines traditional pytest unit/integration testing with Behavior-Driven Development (BDD) using pytest-bdd to ensure both technical correctness and alignment with user requirements.

## Directory Structure

The tests are organized to mirror the source code structure while separating unit tests from BDD feature definitions and step implementations:

```
src/pyrovelocity/tests/
├── __init__.py
├── conftest.py                # Shared pytest fixtures and configuration
├── fixtures/                  # Test data generation utilities
├── models/                    # Tests for model components
│   ├── __init__.py
│   ├── modular/               # Tests for modular implementation
│   │   ├── __init__.py
│   │   ├── conftest.py        # Modular-specific fixtures
│   │   ├── components/        # Unit tests for component implementations
│   │   │   ├── test_dynamics.py
│   │   │   ├── test_guides.py
│   │   │   ├── test_likelihoods.py
│   │   │   ├── test_observations.py
│   │   │   └── test_priors.py
│   │   ├── integration/       # BDD step implementations
│   │   │   ├── __init__.py
│   │   │   ├── conftest.py    # BDD-specific fixtures
│   │   │   ├── test_dynamics_model_steps.py
│   │   │   ├── test_model_steps.py
│   │   │   └── test_prior_model_steps.py
│   │   ├── selection/         # Tests for model selection
│   │   ├── utils/             # Tests for utility modules
│   │   ├── test_anndata_integration.py
│   │   ├── test_comparison.py
│   │   ├── test_factory.py
│   │   ├── test_interfaces.py
│   │   ├── test_model.py
│   │   ├── test_registry.py
│   │   └── test_unified_config.py
│   ├── jax/                   # Tests for JAX implementation (not yet documented)
│   └── test_*.py              # Tests for legacy implementation
├── features/                  # BDD feature specifications
│   └── models/
│       └── modular/
│           ├── dynamics_model.feature
│           ├── guide_model.feature
│           ├── likelihood_model.feature
│           ├── model.feature
│           ├── observation_model.feature
│           └── prior_model.feature
├── plots/                     # Tests for plotting functions
├── utils/                     # Tests for utility functions
├── validation/                # Tests for validation framework
└── test_*.py                  # Package-level tests
```

## Testing Philosophy

Our testing approach implements a dual strategy that addresses both implementation details and user-facing behavior:

1. **Unit and Integration Tests** (`tests/models/modular/*.py`):
   - Traditional pytest-based tests focusing on implementation details
   - Verify internal functions, edge cases, and error handling
   - Target high code coverage and technical correctness
   - Validate that individual components work as expected in isolation

2. **Behavior-Driven Tests**:
   - **Feature Definitions** (`tests/features/models/modular/*.feature`):
     - User-centric specifications written in Gherkin syntax
     - Focus on requirements and use cases from the user's perspective
     - Serve as living documentation of expected behavior
   - **Feature Implementations** (`tests/models/modular/integration/*.py`):
     - Connect Gherkin specifications to actual code
     - Test how components work together to deliver user-facing functionality
     - Validate that the system behaves according to specifications

This approach follows Acceptance Test-Driven Development (ATDD), where feature files are written before implementation to document requirements and guide development.

### Key Principles

PyroVelocity's testing philosophy emphasizes:

- **Real Objects Over Mocks**: Creating simple instances of real objects rather than using mocking tools
- **Test-Driven Development**: Writing tests alongside code development
- **Reasonable Constraints**: Ensuring tests provide reasonable constraints on the code
- **Systematic Validation**: Comprehensive validation of model implementations
- **Protocol-First Testing**: Testing against interfaces rather than concrete implementations

## Running Tests

### All Tests

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests
```

### Only Modular Implementation Tests

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests/models/modular
```

### Only BDD Tests

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests/models/modular/integration
```

### Specific Component Tests

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests/models/modular/components/test_dynamics.py
```

### Running with Coverage

```bash
source .venv/bin/activate
pytest src/pyrovelocity/tests --cov=pyrovelocity --cov-report=term-missing
```

### Profiling and Benchmarking

PyroVelocity uses Scalene for profiling and benchmarking:

```bash
source .venv/bin/activate
scalene --profile-all src/pyrovelocity/tests/models/test_model.py
```

## Test Development Workflow

Our recommended workflow follows these steps:

1. **Specify Behavior**: Create or update a feature file in `features/` for new functionality
2. **Implement Step Definitions**: Create/update step definitions in the subpackage's `integration/` directory
3. **Add Unit Tests**: Create detailed unit tests in the subpackage's test modules
4. **Implement Source Code**: Develop the actual functionality in the source code
5. **Update Documentation**: Update this document's reference tables when adding new tests

## Validation Framework

PyroVelocity includes a validation framework (PRD-678-VALID) to verify that the modular model can replicate the legacy model. Key aspects include:

- Direct comparison of model outputs
- Validation using synthetic data from test fixtures
- Validation using downsampled real datasets (e.g., `tests/data/preprocessed_pancreas_50_7.json`)
- Systematic addressing of implementation differences

The validation framework is implemented in the `src/pyrovelocity/tests/validation/` directory and includes tests for:

- Direct comparison of model outputs
- Metrics for quantifying differences
- Visualization of differences

## Feature Specifications Reference

The table below tracks all feature specifications defined in our BDD feature files for the modular implementation. This serves as a living document of the intended functionality for the library.

| Module | Feature | Status | Description |
|--------|---------|--------|-------------|
| dynamics_model | Standard dynamics model computes expected counts | ✅ Implemented | Verify that the StandardDynamicsModel correctly computes expected unspliced and spliced counts |
| dynamics_model | Standard dynamics model computes steady state | ✅ Implemented | Verify that the StandardDynamicsModel correctly computes steady state values |
| dynamics_model | Legacy dynamics model matches legacy implementation | ✅ Implemented | Verify that the LegacyDynamicsModel matches the legacy implementation |
| dynamics_model | Dynamics model handles edge cases | ✅ Implemented | Verify that the dynamics model handles edge cases gracefully |
| dynamics_model | Dynamics model with library size correction | ✅ Implemented | Verify that the dynamics model correctly applies library size correction |
| prior_model | LogNormal prior model creates stochastic nodes | ✅ Implemented | Verify that the LogNormalPriorModel creates appropriate stochastic nodes |
| likelihood_model | Poisson likelihood model creates observation nodes | ✅ Implemented | Verify that the PoissonLikelihoodModel creates appropriate observation nodes |
| likelihood_model | Legacy likelihood model matches legacy implementation | ✅ Implemented | Verify that the LegacyLikelihoodModel matches the legacy implementation |
| observation_model | Standard observation model preprocesses data | ✅ Implemented | Verify that the StandardObservationModel correctly preprocesses data |
| guide_model | Auto guide factory creates guide | ✅ Implemented | Verify that the AutoGuideFactory creates an appropriate guide |
| guide_model | Legacy auto guide factory matches legacy implementation | ✅ Implemented | Verify that the LegacyAutoGuideFactory matches the legacy implementation |
| model | PyroVelocity model integrates components | ✅ Implemented | Verify that the PyroVelocityModel correctly integrates all components |
| model | PyroVelocity model trains on AnnData | ✅ Implemented | Verify that the PyroVelocityModel can train on AnnData objects |

### Legend
- ✅ Implemented: Feature is fully implemented and tested
- ⚠️ In Progress: Implementation or testing is underway
- 📝 Specified: Feature is specified but not yet implemented
- 🔄 Planned: Feature is planned but not fully specified

## Unit Test Coverage Reference

This table tracks the unit test coverage for each module in the modular implementation, helping identify areas that need additional testing.

| Module | Test File | Test Coverage | Test Cases | Notes |
|--------|-----------|---------------|------------|-------|
| components/dynamics.py | test_dynamics.py | High | 10+ | Tests for StandardDynamicsModel and LegacyDynamicsModel |
| components/guides.py | test_guides.py | High | 5+ | Tests for AutoGuideFactory and LegacyAutoGuideFactory |
| components/likelihoods.py | test_likelihoods.py | High | 5+ | Tests for PoissonLikelihoodModel and LegacyLikelihoodModel |
| components/observations.py | test_observations.py | High | 5+ | Tests for StandardObservationModel |
| components/priors.py | test_priors.py | High | 5+ | Tests for LogNormalPriorModel |
| model.py | test_model.py | High | 10+ | Tests for PyroVelocityModel |
| factory.py | test_factory.py | High | 10+ | Tests for factory functions |
| registry.py | test_registry.py | High | 5+ | Tests for component registries |
| interfaces.py | test_interfaces.py | Medium | 5+ | Tests for Protocol interfaces |
| data/anndata.py | test_anndata_integration.py | High | 5+ | Tests for AnnData integration |
| utils/*.py | utils/test_*.py | Medium | 10+ | Tests for utility functions |

## BDD Feature Files Explained

Our feature files use Gherkin syntax to describe expected behavior. Each feature file follows this structure:

```gherkin
Feature: [Feature Name]
  As a [type of user]
  I want to [perform some action]
  So that [achieve some benefit]

  Background:
    Given [common setup steps]

  Scenario: [Specific use case name]
    Given [preconditions]
    When [actions]
    Then [expected results]
```

### Example Feature File

```gherkin
Feature: Dynamics Model
  As a computational biologist
  I want to model RNA velocity dynamics
  So that I can understand transcriptional dynamics in single-cell data

  Background:
    Given I have a dynamics model component
    And I have input data with unspliced and spliced counts

  Scenario Outline: Standard dynamics model computes expected counts
    Given I have a StandardDynamicsModel
    When I run the forward method with alpha <alpha>, beta <beta>, and gamma <gamma>
    Then the model should compute expected unspliced and spliced counts
    And the expected counts should follow RNA velocity dynamics

    Examples:
      | alpha | beta | gamma |
      | 1.0   | 0.5  | 0.2   |
```

## Step Definitions Structure

Step definition files connect feature files to actual test code. They are organized by component in the `integration/` directory:

```python
# src/pyrovelocity/tests/models/modular/integration/test_dynamics_model_steps.py
from importlib.resources import files
import pytest
import torch
from pytest_bdd import scenarios, given, when, then, parsers

# Import the feature file using importlib.resources
scenarios(str(files("pyrovelocity.tests.features") / "models" / "modular" / "dynamics_model.feature"))

# Import the components
from pyrovelocity.models.modular.components import (
    LegacyDynamicsModel,
    StandardDynamicsModel,
)

@given("I have a dynamics model component")
def dynamics_model_component():
    """Create a generic dynamics model component."""
    return StandardDynamicsModel()

@when(parsers.parse("I run the forward method with alpha {alpha}, beta {beta}, and gamma {gamma}"), target_fixture="run_forward_method_with_parameters")
def run_forward_method_with_parameters_fixture(standard_dynamics_model, input_data, alpha, beta, gamma):
    """Run the forward method with the given parameters."""
    # Convert string parameters to float
    alpha_val = float(alpha)
    beta_val = float(beta)
    gamma_val = float(gamma)

    # Create context with input data and parameters
    context = {
        "u_obs": input_data["u_obs"],
        "s_obs": input_data["s_obs"],
        "alpha": torch.tensor([alpha_val] * input_data["n_genes"]),
        "beta": torch.tensor([beta_val] * input_data["n_genes"]),
        "gamma": torch.tensor([gamma_val] * input_data["n_genes"]),
    }

    # Run the forward method
    result_context = standard_dynamics_model.forward(context)

    # Store the result for later steps
    return result_context
```

## Contributing New Tests

When contributing new tests to the project:

1. **Check Existing Coverage**: Review the feature and unit test reference tables
2. **Update Feature Files**: If implementing a new feature, add or update the relevant feature file
3. **Follow the Workflow**: Implement step definitions and unit tests as outlined above
4. **Run Tests Locally**: Ensure all tests pass before submitting a PR
5. **Update Documentation**: Update the reference tables in this document

## Fixtures and Testing Utilities

PyroVelocity provides a comprehensive set of fixtures for testing, defined in various `conftest.py` files:

### Main Fixtures (`src/pyrovelocity/tests/conftest.py`)

| Fixture Name | Purpose | Scope |
|--------------|---------|-------|
| `adata_preprocessed_pancreas_50_7` | Provides preprocessed pancreas dataset | Function |
| `adata_trained_pancreas_50_7` | Provides trained pancreas dataset | Function |
| `adata_postprocessed_pancreas_50_7` | Provides postprocessed pancreas dataset | Function |
| `default_sample_data` | Provides synthetic sample data | Function |
| `temp_file_path` | Provides a temporary file path | Function |
| `temp_compressed_pickle_path` | Provides a temporary compressed pickle path | Function |
| `save_and_load_helper` | Helper for saving and loading data | Function |

### Modular Implementation Fixtures (`src/pyrovelocity/tests/models/modular/conftest.py`)

These fixtures are specific to the modular implementation and provide test instances of various components.

### BDD-Specific Fixtures (`src/pyrovelocity/tests/models/modular/integration/conftest.py`)

| Fixture Name | Purpose | Scope |
|--------------|---------|-------|
| `bdd_simple_data` | Provides simple data for BDD testing | Function |
| `bdd_model_parameters` | Provides model parameters for BDD testing | Function |
| `bdd_standard_dynamics_model` | Provides a StandardDynamicsModel | Function |
| `bdd_legacy_dynamics_model` | Provides a LegacyDynamicsModel | Function |
| `bdd_lognormal_prior_model` | Provides a LogNormalPriorModel | Function |
| `bdd_poisson_likelihood_model` | Provides a PoissonLikelihoodModel | Function |
| `bdd_legacy_likelihood_model` | Provides a LegacyLikelihoodModel | Function |
| `bdd_standard_observation_model` | Provides a StandardObservationModel | Function |
| `bdd_auto_guide_factory` | Provides an AutoGuideFactory | Function |
| `bdd_legacy_auto_guide_factory` | Provides a LegacyAutoGuideFactory | Function |
| `bdd_pyro_velocity_model` | Provides a PyroVelocityModel | Function |
| `bdd_anndata` | Provides a simple AnnData object | Function |
| `clear_pyro_param_store` | Clears Pyro's parameter store | Function (autouse) |

## Advanced Testing Topics

### Parameterized Testing

Both unit tests and BDD scenarios can be parameterized:

```python
# Unit test parameterization
@pytest.mark.parametrize("alpha,beta,gamma,expected_u_ss,expected_s_ss", [
    (1.0, 0.5, 0.2, 2.0, 5.0),
    (2.0, 1.0, 0.5, 2.0, 4.0),
])
def test_steady_state(alpha, beta, gamma, expected_u_ss, expected_s_ss):
    model = StandardDynamicsModel()
    u_ss, s_ss = model.steady_state(
        torch.tensor([alpha]),
        torch.tensor([beta]),
        torch.tensor([gamma]),
    )
    assert torch.allclose(u_ss, torch.tensor([expected_u_ss]))
    assert torch.allclose(s_ss, torch.tensor([expected_s_ss]))
```

```gherkin
# BDD parameterization with Scenario Outline
Scenario Outline: Standard dynamics model computes steady state
  Given I have a StandardDynamicsModel
  When I compute the steady state with alpha <alpha>, beta <beta>, and gamma <gamma>
  Then the steady state unspliced should equal alpha/beta
  And the steady state spliced should equal alpha/gamma

  Examples:
    | alpha | beta | gamma |
    | 1.0   | 0.5  | 0.2   |
    | 2.0   | 1.0  | 0.5   |
```

### Testing with AnnData

PyroVelocity provides fixtures for testing with AnnData objects:

```python
def test_anndata_integration(bdd_anndata, bdd_pyro_velocity_model):
    """Test integration with AnnData."""
    # Prepare AnnData
    adata = bdd_anndata.copy()

    # Train model
    bdd_pyro_velocity_model.train(
        adata,
        max_epochs=2,
        batch_size=5,
        use_gpu=False,
    )

    # Check that results were stored in AnnData
    assert "dynamics" in adata.uns
    assert "model_type" in adata.uns["dynamics"]
```

### Testing with Pyro

PyroVelocity provides utilities for testing with Pyro:

```python
def test_pyro_model(clear_pyro_param_store):
    """Test a Pyro model."""
    # Define a simple model
    def model():
        x = pyro.sample("x", pyro.distributions.Normal(0, 1))
        return x

    # Define a guide
    def guide():
        loc = pyro.param("loc", torch.tensor(0.0))
        scale = pyro.param("scale", torch.tensor(1.0), constraint=pyro.constraints.positive)
        pyro.sample("x", pyro.distributions.Normal(loc, scale))

    # Run inference
    pyro.clear_param_store()
    svi = pyro.infer.SVI(
        model=model,
        guide=guide,
        optim=pyro.optim.Adam({"lr": 0.01}),
        loss=pyro.infer.Trace_ELBO(),
    )

    # Train for a few steps
    for _ in range(10):
        loss = svi.step()

    # Check that parameters were updated
    assert pyro.param("loc").item() != 0.0
```

## Continuous Integration

Tests are automatically run as part of the CI/CD pipeline defined in `.github/workflows/cid.yaml`. The CI pipeline includes:

1. **Unit Tests**: Run all unit tests with coverage reporting
2. **BDD Tests**: Run all BDD tests to verify behavior
3. **Coverage Enforcement**: Ensure test coverage meets minimum thresholds

## Best Practices for PyroVelocity Testing

1. **Use Protocol Interfaces**: Test against Protocol interfaces rather than concrete implementations
2. **Create Real Objects**: Prefer creating simple instances of real objects over using mocks
3. **Test Edge Cases**: Include tests for edge cases and error handling
4. **Use Fixtures**: Leverage pytest fixtures for common setup and teardown
5. **Clear Pyro State**: Always clear Pyro's parameter store before and after tests
6. **Synthetic Data**: Use synthetic data for tests to avoid external dependencies
7. **Parameterize Tests**: Use parameterization to test multiple cases with the same logic
8. **Document Tests**: Include docstrings and comments to explain test purpose and logic
9. **Follow BDD Patterns**: Use Given-When-Then pattern for BDD tests
10. **Validate Against Legacy**: Validate new implementations against legacy code