### 2025-05-15 nbs/guides/model_components.qmd
docs(guides): init model components
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Customizing Modular Components"
format:
  html:
    code-fold: false
---

# How to Customize Modular Components in PyroVelocity

This guide shows how to customize components in PyroVelocity's modular implementation. The modular architecture allows you to create custom components that implement the Protocol interfaces, enabling you to tailor the model to your specific needs.

## Available Component Types

PyroVelocity's modular implementation defines five main component types:

1. **Dynamics Model**: Defines the mathematical models of RNA dynamics
2. **Prior Model**: Defines prior distributions for model parameters
3. **Likelihood Model**: Defines observation distributions
4. **Observation Model**: Handles data preprocessing and transformation
5. **Guide Model**: Creates guide functions for variational inference

Each component type has a corresponding Protocol interface that defines the contract that implementations must follow.

## Creating a Custom Dynamics Model

Let's create a custom dynamics model that implements a nonlinear ODE system:

```python
import torch
import pyro
from typing import Dict, Any, Tuple
from pyrovelocity.models.modular.interfaces import DynamicsModel
from pyrovelocity.models.modular.registry import DynamicsModelRegistry

@DynamicsModelRegistry.register("nonlinear")
class NonlinearDynamicsModel:
    """
    Custom dynamics model implementing a nonlinear ODE system.

    This model extends the standard RNA velocity model with nonlinear terms.
    """

    def __init__(self, shared_time: bool = True, t_scale_on: bool = False):
        """
        Initialize the nonlinear dynamics model.

        Args:
            shared_time: Whether to use shared time across genes
            t_scale_on: Whether to use time scaling
        """
        self.name = "nonlinear"
        self.description = "Nonlinear dynamics model for RNA velocity"
        self.shared_time = shared_time
        self.t_scale_on = t_scale_on

    def forward(self, context: Dict[str, Any]) -> Dict[str, Any]:
        """
        Compute expected RNA counts based on nonlinear dynamics model.

        Args:
            context: Context dictionary with model state

        Returns:
            Updated context with computed values
        """
        # Extract parameters from context
        u_obs = context["u_obs"]
        s_obs = context["s_obs"]
        alpha = context["alpha"]
        beta = context["beta"]
        gamma = context["gamma"]
        t = context.get("t", torch.ones_like(u_obs))

        # Compute steady state
        u_inf, s_inf = self.steady_state(alpha, beta, gamma)

        # Compute expected counts using numerical approximation
        # For simplicity, we'll use a basic Euler method here
        dt = 0.1
        steps = int(torch.max(t) / dt)

        u_expected = u_obs.clone()
        s_expected = s_obs.clone()

        for _ in range(steps):
            # Nonlinear ODE system
            # du/dt = alpha - beta * u * (1 + s/10)
            # ds/dt = beta * u - gamma * s
            u_delta = alpha - beta * u_expected * (1 + s_expected/10)
            s_delta = beta * u_expected - gamma * s_expected

            u_expected = u_expected + dt * u_delta
            s_expected = s_expected + dt * s_delta

        # Update context with computed values
        context.update({
            "u_expected": u_expected,
            "s_expected": s_expected,
            "u_inf": u_inf,
            "s_inf": s_inf,
        })

        return context

    def steady_state(
        self, alpha: torch.Tensor, beta: torch.Tensor, gamma: torch.Tensor, **kwargs
    ) -> Tuple[torch.Tensor, torch.Tensor]:
        """
        Compute steady-state RNA counts.

        Args:
            alpha: Transcription rate
            beta: Splicing rate
            gamma: Degradation rate

        Returns:
            Tuple of steady-state unspliced and spliced counts
        """
        # For the nonlinear model, we need to solve:
        # alpha - beta * u_inf * (1 + s_inf/10) = 0
        # beta * u_inf - gamma * s_inf = 0

        # From the second equation: s_inf = beta * u_inf / gamma
        # Substituting into the first:
        # alpha - beta * u_inf * (1 + beta * u_inf / (gamma * 10)) = 0

        # For simplicity, we'll use an approximation here
        # In a real implementation, you might use a numerical solver

        # First approximation: ignore the nonlinear term
        u_inf_approx = alpha / beta
        s_inf_approx = alpha / gamma

        # Refine with a few iterations
        for _ in range(5):
            s_inf_approx = beta * u_inf_approx / gamma
            u_inf_approx = alpha / (beta * (1 + s_inf_approx/10))

        return u_inf_approx, s_inf_approx
```

## Creating a Custom Prior Model

Now, let's create a custom prior model with informative priors:

```python
import torch
import pyro
from typing import Dict, Any
from pyrovelocity.models.modular.interfaces import PriorModel
from pyrovelocity.models.modular.registry import PriorModelRegistry

@PriorModelRegistry.register("informative")
class InformativePriorModel:
    """
    Custom prior model with informative priors.

    This model uses informative priors based on prior knowledge about
    RNA dynamics parameters.
    """

    def __init__(self):
        """Initialize the informative prior model."""
        self.name = "informative"
        self.description = "Informative prior model for RNA velocity"

    def forward(self, context: Dict[str, Any]) -> Dict[str, Any]:
        """
        Sample parameters from informative prior distributions.

        Args:
            context: Context dictionary with model state

        Returns:
            Updated context with sampled parameters
        """
        # Extract dimensions from context
        u_obs = context["u_obs"]
        num_cells, num_genes = u_obs.shape

        # Create plates for batched sampling
        with pyro.plate("genes", num_genes):
            # Sample alpha (transcription rate) from gamma distribution
            # with informative shape and rate parameters
            alpha = pyro.sample(
                "alpha",
                pyro.distributions.Gamma(
                    torch.ones(num_genes) * 2.0,  # shape parameter
                    torch.ones(num_genes) * 0.5   # rate parameter
                )
            )

            # Sample beta (splicing rate) from gamma distribution
            beta = pyro.sample(
                "beta",
                pyro.distributions.Gamma(
                    torch.ones(num_genes) * 4.0,  # shape parameter
                    torch.ones(num_genes) * 2.0   # rate parameter
                )
            )

            # Sample gamma (degradation rate) from gamma distribution
            gamma = pyro.sample(
                "gamma",
                pyro.distributions.Gamma(
                    torch.ones(num_genes) * 3.0,  # shape parameter
                    torch.ones(num_genes) * 1.5   # rate parameter
                )
            )

        # Update context with sampled parameters
        context.update({
            "alpha": alpha,
            "beta": beta,
            "gamma": gamma,
        })

        return context
```

## Using Custom Components

Now, let's use our custom components to create a model:

```python
from pyrovelocity.models.modular.model import PyroVelocityModel
from pyrovelocity.models.modular.components import (
    StandardObservationModel,
    PoissonLikelihoodModel,
    AutoGuideFactory
)

# Create components
dynamics_model = NonlinearDynamicsModel(shared_time=True, t_scale_on=False)
prior_model = InformativePriorModel()
likelihood_model = PoissonLikelihoodModel()
observation_model = StandardObservationModel(correct_library_size=True)
guide_model = AutoGuideFactory()

# Create model
custom_model = PyroVelocityModel(
    dynamics_model=dynamics_model,
    prior_model=prior_model,
    likelihood_model=likelihood_model,
    observation_model=observation_model,
    guide_model=guide_model
)

# Train the model
custom_model.train(
    adata=adata,
    max_epochs=100,
    learning_rate=0.01,
    use_gpu=False
)
```

## Creating a Model with Factory Functions

You can also use factory functions to create models with custom components:

```python
from pyrovelocity.models.modular.factory import create_model_from_config
from pyrovelocity.models.modular.config import (
    ModelConfig,
    ComponentConfig
)

# Create component configurations
dynamics_config = ComponentConfig(name="nonlinear", params={"shared_time": True})
prior_config = ComponentConfig(name="informative", params={})
likelihood_config = ComponentConfig(name="poisson", params={})
observation_config = ComponentConfig(name="standard", params={"correct_library_size": True})
guide_config = ComponentConfig(name="auto", params={})

# Create model configuration
model_config = ModelConfig(
    dynamics_model=dynamics_config,
    prior_model=prior_config,
    likelihood_model=likelihood_config,
    observation_model=observation_config,
    inference_guide=guide_config
)

# Create model from configuration
model = create_model_from_config(model_config)
```

## Registering Custom Components

Custom components need to be registered with the appropriate registry to be used with factory functions. The registration is done using the `@Registry.register` decorator:

```python
from pyrovelocity.models.modular.registry import (
    DynamicsModelRegistry,
    PriorModelRegistry,
    LikelihoodModelRegistry,
    ObservationModelRegistry,
    InferenceGuideRegistry
)

# Register a custom dynamics model
@DynamicsModelRegistry.register("my_dynamics")
class MyDynamicsModel:
    # Implementation...
    pass

# Register a custom prior model
@PriorModelRegistry.register("my_prior")
class MyPriorModel:
    # Implementation...
    pass

# Register a custom likelihood model
@LikelihoodModelRegistry.register("my_likelihood")
class MyLikelihoodModel:
    # Implementation...
    pass

# Register a custom observation model
@ObservationModelRegistry.register("my_observation")
class MyObservationModel:
    # Implementation...
    pass

# Register a custom inference guide
@InferenceGuideRegistry.register("my_guide")
class MyGuideFactory:
    # Implementation...
    pass
```

## Conclusion

This guide showed how to create and use custom components in PyroVelocity's modular implementation. By implementing the Protocol interfaces and registering your components, you can create models tailored to your specific needs.

For more information, refer to the [API reference](../reference/models.PyroVelocityModel.qmd) and the [modular implementation tutorial](../tutorials/models.qmd).