### 2025-05-15 nbs/tutorials/models.qmd
docs(tutorials): rejoin models code blocks and print model structure
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Model definition"
execute:
  freeze: true
  eval: true
  warning: false
  error: false
  cache: true
toc: true
number-sections: true
highlight-style: gruvbox
csl: ../../bibstyle.csl
lightbox: auto
format:
  html:
    html-math-method: katex
#   ipynb: default
---

# Introduction to PyroVelocity Modular Models

This tutorial introduces PyroVelocity modular models, which follows a Protocol-First architecture with composable components. The modular implementation provides a flexible and extensible framework for building RNA velocity models.

## Prerequisites

Before starting, make sure you have PyroVelocity [installed](../guides/installation.qmd).

## Basic Usage

Let's start with a simple example using the modular implementation:

```python
import torch
import pyro
import anndata as ad
import numpy as np
import matplotlib.pyplot as plt
import scanpy as sc
from pyrovelocity.utils import print_anndata, anndata_string, print_string_diff
from pyrovelocity.models.modular import PyroVelocityModel
from pyrovelocity.models.modular.factory import create_legacy_model1

# Set random seed for reproducibility
torch.manual_seed(42)
np.random.seed(42)
pyro.set_rng_seed(42)

# Create synthetic data
n_cells, n_genes = 100, 10
u_data = np.random.poisson(5, size=(n_cells, n_genes))
s_data = np.random.poisson(5, size=(n_cells, n_genes))

# Create AnnData object
adata = ad.AnnData(X=s_data)
adata.layers["spliced"] = s_data
adata.layers["unspliced"] = u_data
adata.obs_names = [f"cell_{i}" for i in range(n_cells)]
adata.var_names = [f"gene_{i}" for i in range(n_genes)]
print_anndata(adata)
initial_data_state_representation = anndata_string(adata)

# Set up AnnData for PyroVelocity
adata = PyroVelocityModel.setup_anndata(adata)
post_setup_anndata_representation = anndata_string(adata)
print_string_diff(
    initial_data_state_representation,
    post_setup_anndata_representation,
    diff_title="Setup AnnData diff",
)
print_anndata(adata)

# Compute PCA and UMAP embeddings
sc.pp.pca(adata)
sc.pp.neighbors(adata)
sc.tl.umap(adata)
post_embedding_representation = anndata_string(adata)
print_string_diff(
    post_setup_anndata_representation,
    post_embedding_representation,
    diff_title="Embedding diff",
)
print_anndata(adata)

# Create a standard model with default components
model = create_legacy_model1()
print(model)

# Train the model
model.train(
    adata=adata,
    max_epochs=20,  # Use a small number for demonstration
    learning_rate=0.01,
    use_gpu=False  # Set to True if GPU is available
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=30  # Use a small number for demonstration
)

# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)

# Compute velocity and store it in AnnData
velocity = model.get_velocity(
    adata=adata,
    num_samples=30,
    basis='umap'  # This will be used for embedding
)

# Print available keys in posterior samples
print("Available keys in posterior samples:")
for key in posterior_samples.keys():
    print(f"- {key}: {posterior_samples[key].shape}")
print(model)
```

## Understanding the Modular Architecture

The modular implementation is based on a Protocol-First architecture with five main component types:

1. **Dynamics Model**: Defines the mathematical models of RNA dynamics
2. **Prior Model**: Defines prior distributions for model parameters
3. **Likelihood Model**: Defines observation distributions
4. **Observation Model**: Handles data preprocessing and transformation
5. **Guide Model**: Creates guide functions for variational inference

Each component implements a Protocol interface, allowing for flexible composition and extension.

## Creating Custom Models

You can create custom models by combining different components:

```python
from pyrovelocity.models.modular.components import (
    StandardDynamicsModel,
    LegacyDynamicsModel,
    LogNormalPriorModel,
    PoissonLikelihoodModel,
    LegacyLikelihoodModel,
    StandardObservationModel,
    AutoGuideFactory
)
from pyrovelocity.models.modular.model import PyroVelocityModel

# Create components
dynamics_model = LegacyDynamicsModel(shared_time=True, t_scale_on=False)
prior_model = LogNormalPriorModel()
likelihood_model = LegacyLikelihoodModel()
observation_model = StandardObservationModel(correct_library_size=True)
guide_model = AutoGuideFactory()

# Create model
custom_model = PyroVelocityModel(
    dynamics_model=dynamics_model,
    prior_model=prior_model,
    likelihood_model=likelihood_model,
    observation_model=observation_model,
    guide_model=guide_model
)

# Train the model
custom_model.train(
    adata=adata,
    max_epochs=20,
    learning_rate=0.01,
    use_gpu=False
)
```

## Using Factory Functions

The modular implementation provides factory functions for creating models with different configurations:

```python
from pyrovelocity.models.modular.factory import (
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig

# Create a model that replicates the legacy implementation
legacy_model = create_legacy_model1()

# Create a model from a configuration
config = ModelConfig.standard()
config_model = create_model_from_config(config)
```

## Visualizing Results

After training the model and generating posterior samples, you can visualize the results:

```python
import scanpy as sc
import scvelo as scv

# Compute PCA and UMAP embeddings
sc.pp.pca(adata)
sc.pp.neighbors(adata)
sc.tl.umap(adata)

# Visualize velocity
scv.pl.velocity_embedding_grid(
    adata,
    basis='umap',
    color='gene_0',
    vkey='velocity_pyro'  # Use the velocity key added by get_velocity
    color="gray",
    s=3.5,
    show=False,
    alpha=0.25,
    min_mass=3.5,
    scale=0.35,
    frameon=False,
    density=0.4,
    arrow_size=3,
    linewidth=1,
)
```

## Conclusion

This tutorial introduced the basics of PyroVelocity's modular implementation. The modular architecture provides a flexible and extensible framework for building RNA velocity models, allowing you to customize components and create models tailored to your specific needs.

For more advanced usage, refer to the [API reference](../reference/models.PyroVelocityModel.qmd) and [how-to guides](../guides/model_components.qmd).