### 2025-05-27 nbs/guides/inference.qmd
docs: preprocess data in inference guide
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Inference with PyroVelocity"
format:
  html:
    code-fold: false
---

# Inference with PyroVelocity

This guide explains how to perform inference with PyroVelocity, covering both the legacy and modular implementations.

## Legacy Implementation

The legacy implementation provides a single `PyroVelocity` class that handles all aspects of the model:

```python
import anndata as ad
import numpy as np
from pyrovelocity.models import PyroVelocity
from pyrovelocity.tasks.preprocess import preprocess_dataset
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata, _, _ = preprocess_dataset("bifurcation_14", adata)

# Set up AnnData for PyroVelocity
PyroVelocity.setup_anndata(adata)

# Create model
model = PyroVelocity(
    adata=adata,
    likelihood="Poisson",
    model_type="auto",
    guide_type="auto",
    correct_library_size=True,
    add_offset=True,
)

# Train model
model.train_faster(
    max_epochs=1000,
    lr=0.01,
    use_gpu="auto"
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=5,
)

# Compute statistics from posterior samples
model.compute_statistics_from_posterior_samples(
    adata=adata,
    posterior_samples=posterior_samples
)

# Save model
model.save_model("models/bifurcation_14_model2/model")

# Load model
model = PyroVelocity.load_model("models/bifurcation_14_model2/model", adata)
```

## Modular Implementation

The modular implementation provides a more flexible approach with composable components:

```python
import anndata as ad
import numpy as np
from pyrovelocity.models.modular import (
    PyroVelocityModel
)
from pyrovelocity.models.modular.factory import create_legacy_model2
from pyrovelocity.tasks.preprocess import preprocess_dataset
from pyrovelocity.io.datasets import bifurcation_14
from pyrovelocity.utils import print_anndata

# Load data
adata = bifurcation_14()
adata, _, _ = preprocess_dataset("bifurcation_14", adata)

# Set up AnnData for PyroVelocity
adata = PyroVelocityModel.setup_anndata(adata)

# Create model using factory function
model = create_legacy_model2()

# Train model
model.train(
    adata=adata,
    max_epochs=1000,
    learning_rate=0.01,
    use_gpu=False  # Set to True if GPU is available
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=5,
)

# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)
print_anndata(adata)
```

### Creating Custom Models

The modular implementation allows you to create custom models by combining different components:

```python
from pyrovelocity.models.modular.factory import ComponentFactory
from pyrovelocity.models.modular.config import ComponentConfig, ComponentType
from pyrovelocity.models.modular.model import PyroVelocityModel
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)

# Create components consistent with create_legacy_model2
dynamics_model = ComponentFactory.create_component(
    ComponentConfig(
        name="legacy",
        params={
            "shared_time": True,
            "t_scale_on": False,
            "correct_library_size": True,
            "add_offset": True,
        }
    ),
    ComponentType.DYNAMICS_MODEL
)

prior_model = ComponentFactory.create_component(
    ComponentConfig(name="lognormal", params={}),
    ComponentType.PRIOR_MODEL
)

likelihood_model = ComponentFactory.create_component(
    ComponentConfig(name="legacy", params={}),
    ComponentType.LIKELIHOOD_MODEL
)

observation_model = ComponentFactory.create_component(
    ComponentConfig(name="standard", params={}),
    ComponentType.OBSERVATION_MODEL
)

guide_model = ComponentFactory.create_component(
    ComponentConfig(
        name="legacy_auto",
        params={
            "init_scale": 0.1,
            "add_offset": True,
        }
    ),
    ComponentType.INFERENCE_GUIDE
)

# Create model
custom_model = PyroVelocityModel(
    dynamics_model=dynamics_model,
    prior_model=prior_model,
    likelihood_model=likelihood_model,
    observation_model=observation_model,
    guide_model=guide_model
)

# Train the model
custom_model.train(
    adata=adata,
    max_epochs=1000,
    learning_rate=0.01,
    use_gpu=False
)
```

### Using Factory Functions

The modular implementation provides factory functions for creating models with different configurations:

```python
from pyrovelocity.models.modular.factory import (
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig
from pyrovelocity.models.modular.model import PyroVelocityModel
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)

# Create a model that replicates the legacy implementation (add_offset=True)
legacy_model2 = create_legacy_model2()

# Create an alternative legacy model (add_offset=False)
legacy_model1 = create_legacy_model1()

# Create a model from a configuration
config = ModelConfig.standard()
config_model = create_model_from_config(config)
```

## Inference Parameters

Both implementations support the following inference parameters:

| Parameter | Description | Default |
|-----------|-------------|---------|
| max_epochs | Maximum number of training epochs | 1000 |
| learning_rate | Learning rate for the optimizer | 0.01 |
| batch_size | Batch size for mini-batch training | None (full dataset) |
| train_size | Fraction of data to use for training | 0.8 |
| valid_size | Fraction of data to use for validation | 0.2 |
| early_stopping | Whether to use early stopping | True |
| early_stopping_patience | Number of epochs to wait for improvement | 10 |
| use_gpu | Whether to use GPU for training | "auto" |

## Posterior Sampling

After training, you can generate posterior samples to quantify uncertainty:

```python
from pyrovelocity.io.datasets import bifurcation_14
from pyrovelocity.models.modular.factory import create_legacy_model2
from pyrovelocity.models.modular.model import PyroVelocityModel

# Load data and set up model
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)
model = create_legacy_model2()

# Train the model first
model.train(adata=adata, max_epochs=50, learning_rate=0.01, use_gpu=False)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=100
)
```

The posterior samples contain the following keys:

| Key | Description |
|-----|-------------|
| alpha | Transcription rate |
| beta | Splicing rate |
| gamma | Degradation rate |
| u_inf | Steady-state unspliced RNA |
| s_inf | Steady-state spliced RNA |
| t | Latent time |
| switching | Switching time |
| ut | Latent unspliced RNA |
| st | Latent spliced RNA |

## Conclusion

This guide covered the basics of inference with PyroVelocity, including both the legacy and modular implementations. For more advanced usage, refer to the [API reference](../reference/models.PyroVelocityModel.qmd) and the [modular components guide](model_components.qmd).